### 2018-09-12 0_abstract.tex
Initial commit (will not compile)
Brought in LaTeX from dissertation, plus some stuff from last submission.

\section{Abstract}
\label{sec:mvtraits-abstract}

We investigated whether global leaf economic relationships across plant functional types (PFTs) are also present within-PFTs, and the extent to which this hierarchical structure can be used to constrain trait estimates.
We developed a hierarchical multivariate Bayesian model that assumes separate means and covariance structures within and across PFTs and fit this model to seven leaf traits from the TRY database: longevity, specific leaf area, nitrogen content, phosphorus content, dark respiration rate, maximum photosynthetic carboxylation rate ($V_{c,max}$), and maximum photosynthetic electron transport rate ($J_{max}$). 
We evaluated the ability of this model to estimate trait means and reproduce leaf economic trade-offs.
Trait correlations were generally consistent in direction within and across PFTs, and consistent with predictions of the leaf economic spectrum.
However, correlation strength varied substantially across PFTs indicating that leaf economic relationships within PFTs are often confounded by the unique physiology of certain plant types or environmental conditions in certain biomes.
Leveraging covariance in multivariate models reduced uncertainties in mean trait estimates, particularly for undersampled trait-PFT combinations.
However, additional constraint from the across-PFT hierarchy was limited compared to the simple multivariate models.

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% End:

### 2018-09-12 1_introduction.tex
Initial commit (will not compile)
Brought in LaTeX from dissertation, plus some stuff from last submission.

\section{Introduction}

The diversity and dimensionality of the terrestrial biosphere is vast and complex, and therefore there has been a recurring debate in ecology about the utility of reductionist approaches for capturing this variability.
In particular, the use of functional groups with common characteristics has been widely applied in biodiversity studies \cite{naeem_disentangling_2003} and is essential to the structure of many ecosystem models \cite{lavorel_plant_1997,wullschleger_plant_2014}.
However, ecologists have long recognized the importance of individual variability and stochasticity in shaping ecosystems
\cite{gleason_individualistic_1926,bolnick_why_2011,rosindell_unified_2011,clark_why_2016},
and the benefits of more finely-resolved representation of functional diversity for predictive ecology are supported by an increasing body of trait ecology literature
\cite{mayfield_diversity_2006,mcmahon_improving_2011,van_bodegom_going_2012, reichstein_linking_2014,violle_emergence_2014,medlyn_using_2015,moran_intraspecific_2016}.

Plant functional traits can be used to link directly measurable features of individuals to their fitness within an ecosystem and, by extension, ecosystem performance as a whole \cite{violle_let_2007}.
Recent syntheses of global trait databases have revealed that although the functional diversity across plant species is immense, this diversity is constrained by allometries and trade-offs between plant strategies \cite{wright_worldwide_2004,kattge_try_2011,kleyer_why_2015,diaz_global_2016}.
One axis of trait covariation currently receiving attention is the ‘leaf economic spectrum’, which defines a trade-off between plant investment in productive but short-lived leaves versus less productive but sturdy and long-lived leaves \cite{wright_worldwide_2004,shipley_fundamental_2006,reich_world-wide_2014,diaz_global_2016}
Leaf economic traits are well-correlated with
individual plant productivity \cite{shipley_functional_2005,niinemets_within-canopy_2016,wu_convergence_2016},
litter decomposition rates \cite{bakker_leaf_2011,hobbie_plant_2015},
community composition \cite{burns_patterns_2004,cavender-bares_multiple_2004},
and ecosystem function \cite{diaz_plant_2004,musavi_imprint_2015}.
The relative position of plant species along the leaf economic spectrum has been shown to be influenced by climate and soil conditions
\cite{wright_worldwide_2004,wright_modulation_2005,cornwell_community_2009,ordonez_global_2009,wigley_leaf_2016}.
As a result, relationships between leaf economic traits and climate have been incorporated into ecosystem models to allow for continuous variation in plant function and environmental responses
[@sakschewski_leaf_2015; @verheijen_inclusion_2015].

However, the use of among-trait and trait-environment correlations at the global scale, for both ecological inference and land surface modeling, has several important caveats.
First, observed correlations at the global scale do not always hold at smaller scales (such as sites, species, and individuals).
For example, some studies suggested consistent correlations across scales [@wright_worldwide_2004; @albert_multi-trait_2010; @asner_amazonian_2014]
whereas others showed no or even opposite correlations [@albert_intraspecific_2010; @messier_how_2010; @wright_does_2012; @feng_scale_2013; @grubb_relationships_2015; @wigley_leaf_2016; @messier_traitnetwork; @kichenin_2013_contrasting].
Many mechanisms have been suggested for deviation from global trait relationships at smaller scales.
Trade-offs between strategies may only be applicable when multiple competing strategies co-occur, and in cases where strong environmental filters allow only a narrow range of strategies, alternative processes can drive community assembly [@rosado_trait_dominance; @pierce_csr_cooccur; @grim_pierce_book].
Different selective pressures dominate at different scales, particularly within versus across species [@albert_intraspecific_2010; @messier_how_2010; @kichenin_2013_contrasting],
and the large heterogeneity in the variance structures of traits suggests that different traits have contrasting sensitivity to these different pressures [@messier_trait_2016].
Experimental evidence shows that species can alter different aspects of their leaf economy in a relatively uncoordinated fashion, even when the direction of univariate trait responses to environmental change is consistent [@wright_does_2012].
Meanwhile, across different plant functional types, resource allocation patterns, for instance of nutrients to photosynthesis versus structure and defense, differ substantially, suggesting different investment strategies and varying relationships among traits [@ghimire_2017_Nallocation].
Second, among-trait correlations at any scale do not provide causal evidence for functional trade-offs or even similarity in response to external stimuli [@messier_trait_2016].
Therefore, ascribing too much leverage to trait correlations can lead to an underestimation of plant functional diversity [@grubb_trade-offs_2015].
Third, plants maintain their fitness in a given environment through multiple independent strategies (corresponding to multiple mutually orthogonal axes of trait variability).
As a consequence, changes in key leaf economic traits such as leaf nitrogen content and specific leaf area area may not affect other aspects of plant function, such as
hydraulics [@li_leaf_2015],
overall plant carbon budget [@edwards_leaf_2014],
and dispersal [@westoby_plant_2002].
Finally, modeling ecosystem function based on trait correlations is sampling from the hypothetical space of potential species and communities that could have evolved, rather than constraining models to forecast the actual vegetation we have today as the result of spatial separation and constraints on convergent evolution.
Among other problems, this approach fails to account for the timescales required for adaptation as well as actual limitations of the physiology of different species and community assembly.

An alternative approach is to preserve existing PFT classifications
[though potentially with finer taxonomic, functional, or spatial resolution, e.g. @boulangeat_improving_2012]
while using statistical analyses to account for uncertainty and variability in the aggregated trait values.
For example, the Predictive Ecosystem Analyzer (PEcAn, pecanproject.org), an ecosystem model-data informatics system, parameterizes PFTs using trait probability distributions from a Bayesian meta-analysis of plant trait data across many studies
[@dietze_improving_2013; @lebauer_facilitating_2013].
This approach explicitly separates the processes driving PFT-level differentiation from processes that drive finer-scale functional variability,
and is useful for guiding future data collection and model refinement [@dietze_quantitative_2014].
However, a univariate meta-analysis, like the one currently in PEcAn, is limited by its failure to account for trait correlations, therefore neglecting useful knowledge about relationships across PFTs and between traits.
At the other extreme, existing regional and global scale analyses [e.g. @van_bodegom_going_2012; @sakschewski_leaf_2015] ignore variability within PFTs, often resulting in macroecological, evolutionary, and competitive trade-offs across PFTs being used to drive both acclimation and instantaneous responses within PFTs.

While the leaf economic spectrum has been investigated at the global scale, where it is robust, and at the site or plot scale, where deviations from it are common, it has received less attention at the intermediate scale of PFTs.
Thus, this paper seeks to answer the following questions:
First, to what extent does the leaf economic spectrum hold within vs. across plant functional types?
Second, to what extent can the leaf economic spectrum and other patterns of trait covariance be leveraged to constrain trait estimates, particularly under data limitation?
The answer to these question has implications for both functional ecology and ecosystem modelling.
To evaluate these questions, we develop a hierarchical multivariate Bayesian model that explicitly accounts for across- and within-PFT variability in trait correlations.
We then fit this model to a global database of foliar traits to estimate mean trait values and variance-covariance matrices for PFTs as defined in a major earth system model [Community Land Model, CLM, @clm45_note].
We evaluate the ability of this model to reduce uncertainties in trait estimates and reproduce observed patterns of global trait variation compared to non-hierarchical multivariate and univariate models.
Finally, we assess the generality and scale dependence of trait trade-offs by comparing covariance estimates globally and within each PFT.

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% End:

### 2018-09-12 2_methods.tex
Initial commit (will not compile)
Brought in LaTeX from dissertation, plus some stuff from last submission.

\section{Materials and methods}\label{sec:mvtraits-methods}

\subsection{Trait data}

Foliar trait data for this analysis comes from the TRY global traits database (Kattge et al.~2011). \nocite{kattge_try_2011}
We focused our research on seven foliar traits:
Leaf longevity (months),
specific leaf area (SLA, m$^2$ kg$^{-1}$),
leaf nitrogen content ($N_{\mass}$, mg N g$^{-1}$ or $N_{\area}$, g N m$^{-2}$),
leaf phosphorus content ($P_{\mass}$, mg P g$^{-1}$ or $P_{\area}$, g P m$^{-2}$),
leaf dark respiration at \SI{25}{\degreeCelsius} ($R_{d,\mass}$, $\mu$mol g$^{-1}$ s$^{-1}$, or $R_{d,\area}$, $\mu$mol m$^{-2}$ s$^{-1}$),
maximum Rubisco carboxylation rate at \SI{25}{\degreeCelsius} ($V_{c,\max,\mass}$, $\mu$mol g$^{-1}$ s$^{-1}$, or $V_{c,\max,\area}$, $\mu$mol m$^{-2}$ s$^{-1}$),
and maximum electron transport rate at \SI{25}{\degreeCelsius} ($J_{\max,\mass}$, $\mu$mol g$^{-1}$ s$^{-1}$, or $J_{\max,\area}$, $\mu$mol m$^{-2}$ s$^{-1}$.
For $V_{c,\max}$, we only used values already reported in TRY as being at \SI{25}{\degreeCelsius}.
For $R_{d}$, we normalized the values to \SI{25}{\degreeCelsius} based on reported leaf temperature values following the same methods as Atkin et al.~(2015). \nocite{atkin_global_2015}
For $J_{\max}$, we normalized the values to \SI{25}{\degreeCelsius} based on reported leaf temperature values using the temperature response function described in Kattge \& Knorr (2007, Equation 1 therein). \nocite{kattge_2007_temperature}
To avoid potential artifacts caused by different trait normalization, we performed analyses separately for both mass- and area-normalized traits~\cite{osnas_global_2013,lloyd_les}.
We restricted our analysis to TRY data that have been quality-controlled and for which adequate species information was provided for functional type classification (see Kattge et al.~2011).\nocite{kattge_try_2011}

Although the light- and CO2-saturated photosynthetic rate ($A_{\max}$) was an important trait in previous studies, we did not include it in our study for two reasons.
First of all, data on raw photosynthetic rates are highly sensitive to measurement methodology and environmental conditions, which were generally inconsistent or unavailable in TRY\@.
Second, $A_{\max}$ is not a good measure of photosynthetic capacity because it integrates over variability in many physiologically independent traits such as $V_{c,\max}$, $J_{\max}$, and stomatal conductance, and is therefore not used in vegetation models as a photosynthetic parameter~\cite{Ali_2015}.

Following past studies~\cite{wright_worldwide_2004,wright_assessing_2005,onoda_2011_global,diaz_global_2016}, we log-transformed all trait values to correct for their strong right-skewness.

\subsection{Plant functional types}

\begin{longtable}[]{@{}ccc@{}}
\caption{\label{tab:pfts}Names, labels, and species counts for plant
functional types (PFTs) used in this analysis.}\tabularnewline
\toprule
Label & PFT & Number of species\tabularnewline
\midrule
\endfirsthead
\toprule
Label & PFT & Number of species\tabularnewline
\midrule
\endhead
BlETr & Broadleaf Evergreen Tropical & 1229\tabularnewline
BlETe & Broadleaf Evergreen Temperate & 363\tabularnewline
BlDTr & Broadleaf Deciduous Tropical & 286\tabularnewline
BlDTe & Broadleaf Deciduous Temperate & 345\tabularnewline
BlDBo & Broadleaf Deciduous Boreal & 62\tabularnewline
NlETe & Needleleaf Evergreen Temperate & 130\tabularnewline
NlEBo & Needleleaf Evergreen Boreal & 30\tabularnewline
NlD & Needleleaf Deciduous & 19\tabularnewline
ShE & Shrub Evergreen & 1120\tabularnewline
ShDTe & Shrub Deciduous Temperate & 330\tabularnewline
ShDBo & Shrub Deciduous Boreal & 94\tabularnewline
C3GAr & C3 Grass Arctic & 157\tabularnewline
C3GTe & C3 Grass Temperate & 624\tabularnewline
C4G & C4 Grass & 255\tabularnewline
\bottomrule
\end{longtable}

We assigned each species to a unique plant functional type (PFT) following the scheme in the Community Land Model~\cite[CLM4.5,][]{clm45_note}; Table~\ref{tab:pfts}, Fig.~\ref{fig:mvtraits-fig1}. %TODO: Table
We obtained categorical data on growth form, leaf type, phenology, and photosynthetic pathway from the TRY database.
Where species attributes disagreed between datasets (e.g., categorized as a shrub in one dataset but a tree in another), we assigned the attribute that was observed most frequently between the datasets (e.g., if five datasets say ``shrub'' but only one says ``tree'', we would classify it as a shrub).
Where species lacked data on certain attributes, we assigned those attributes based on higher order phylogeny where appropriate (e.g., *Poaceae* family are usually grasses, *Larix spp.* are deciduous needleleaved trees) or otherwise omitted the species from our analyses.
For biome specification, we collected all latitude and longitude data for each species,
matched these data to 30 second ($\sim1$ km$^{2}$) mean annual temperature ($AMT$, averaged 1970--2000) data from WorldClim-2~\cite{worldclim},
calculated the mean AMT for all sites at which that species was observed,
and then binned these species based on the following cutoffs: boreal/arctic ($AMT \leq 5^\circ C$), temperate ($AMT \leq 20^\circ C$), and tropical ($AMT > 20^\circ C$).

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{1_mvtraits/figures/sample_size.pdf}
  \caption{%
    Sample sizes for each trait-PFT pair. $y$ axis is scaled logarithmically.
  }\label{fig:mvtraits-fig1}
\end{figure}

\subsection{Multivariate analysis}

\subsubsection{Basic model description}

In this study, we compared three different models representing different levels of complexity.

The simplest model was the `univariate' model, in which each trait was modeled independently.
For an observation $x_{i,t}$ of trait $t$ and sample $i$:

\[x_{i,t} \sim N(\mu_t, \sigma_t)\]

where $N$ is the univariate normal (Gaussian) distribution with mean $\mu_t$ and standard deviation $\sigma_t$ for trait $t$.

The second-simplest model was the `multivariate' model, in which traits were modeled as samples from a multivariate distribution with a single mean vector and covariance matrix.
For the observed vector of traits ${\mathbf{x_i}}$ for sample $i$:

\[\mathbf{x_i} \sim mvN(\mathbf{\mu}, \mathbf{\Sigma})\]

where $mvN$ is the multivariate normal (Gaussian) distribution with mean vector $\mathbf{\mu}$ and variance-covariance matrix $\mathbf{\Sigma}$.
We ran both of these models independently for each PFT as well as for the entire dataset (as if every observation belonged to the same PFT).

The most complex model was the `hierarchical' model, in which observed trait values were drawn from a PFT-specific multivariate normal distribution describing within-PFT variation and whose parameters were themselves sampled from a global multivariate distribution describing the variation across PFTs.
For the observed vector of traits $\mathbf{x}_{i,p}$ for sample $i$ belonging to PFT $p$:

\[\mathbf{x}_{i,p} \sim mvN(\mathbf{\mu}_p, \mathbf{\Sigma}_p)\]
\[\mathbf{\mu}_p \sim mvN(\mathbf{\mu}_g, \mathbf{\Sigma}_g)\]

where $\mathbf{\mu}_p$ and $\mathbf{\Sigma}_p$ are the mean vector and variance-covariance matrix describing variation within PFT $p$, and $\mathbf{\mu}_g$ and $\mathbf{\Sigma}_g$ are the mean vector and variance-covariance matrix describing across-PFT (global) variation.

\subsubsection{Model implementation}

We fit the above models using a Gibbs sampling algorithm that leveraged known conjugate prior relationships for efficient exploration of the sampling space.
For priors on all multivariate mean vectors ($\mathbf{\mu}$), we used normal distributions:

\[P(\mathbf{\mu}) \sim mvN(\mathbf{\mu}_0, {\mathbf{\Sigma}}_0)\]

This gives rise to the following expression for the posterior:

\[P(\mathbf{\mu} \mid 
    \mathbf{x}, \mathbf{\Sigma}, 
    \mathbf{\mu}_0, \mathbf{\Sigma}_0)
  \sim
  mvN(\mathbf{\mu^*}, \mathbf{\Sigma^*})\]

\[\mathbf{\Sigma^*} = {(\mathbf{\Sigma}_0^{-1} + n \mathbf{\Sigma}^{-1})}^{-1}\]
\[\mathbf{\mu^*} = \mathbf{\mu}_0 \mathbf{\Sigma}_0^{-1} + \bar{\mathbf{x}} n \mathbf{\Sigma}^{-1}\]

where ${\bar{{\mathbf{x}}}}$ are the sample means of the data and $n$ is the number of rows in the data.

For priors on all multivariate variance-covariance matrices, we used the Wishart distribution ($W$):

\[P(\mathbf{\Sigma}) \sim W(\nu_0, \mathbf{S}_0)\]

This gives rise to the following expression for the posterior:

\[P(\mathbf{\Sigma} \mid
  \mathbf{x}, \mathbf{\mu},
  \nu_0, \mathbf{\Sigma}_0)
  \sim
  {(W(\nu^*, S^*))}^{-1}\]

\[\nu^* = 1 + \nu_0 + n + m\]
\[\mathbf{x^*} = \mathbf{x} - \bar{\mu}\]
\[\mathbf{SS} = \mathbf{x^*}^{T} \mathbf{x^*}\]
\[\mathbf{S^*} = {(\mathbf{S}_0 + \mathbf{SS})}^{-1}\]

where $n$ is the number of rows and $m$ is the number of columns in data matrix $x$~\cite[full derivation in][]{gelman_bayesian}.

The fundamentally multivariate nature of the sampling procedure described above makes it incapable of accommodating partially missing observations.
Therefore, our algorithm also included imputation of partially missing data, which proceeded as follows:
For a block of data $\mathbf{x}$ containing missing observations in columns $\mathbf{m}$ and present observations in columns $\mathbf{p}$,
the missing values $\mathbf{x}[m]$ are drawn randomly from a conditional multivariate normal distribution at each iteration of the sampling algorithm:

\[\mathbf{x'}[m|p] \sim mvN(\mathbf{\mu}', \mathbf{\Sigma}')\]

\[\mathbf{\mu'} = 
  (\mathbf{x'}[p] - \mathbf{\mu'}[p]) 
  ({\mathbf{\Sigma}[p,p]}^{-1} \mathbf{\Sigma}[p,m])\]
\[\mathbf{\Sigma'} = \mathbf{\Sigma}[m,m] - 
  \mathbf{\Sigma}[m,p]
  ({\mathbf{\Sigma}[p,p]}^{-1} \mathbf{\Sigma}[p,m])\]

For each model fit, we ran five parallel MCMC chains, continuing the sampling until the final result achieved convergence as determined by a Gelman-Rubin potential scale reduction statistic less than 1.1~\cite{gelman_1992_inference}.
We implemented this sampling algorithm in an open source, publicly available R~\cite[version 3.4.3,][]{rstats} package (\texttt{http://github.com/ashiklom/mvtraits}).

\subsubsection{Analysis of results}

To assess the impact of multivariate and hierarchical constraint on trait estimates,
we compared the mean and 95\% confidence intervals of trait estimates for each PFT from each model (Fig.~\ref{fig:mvtraits-fig2}).
For reference, we also added the default parameter values of CLM 4.5~\cite[Table 8.1 in][]{clm45_note} for SLA, $N_{\mass}$, $N_{\area}$, $V_{c,\max,\mass}$, and $V_{c,\max,\area}$ to Fig.~\ref{fig:mvtraits-fig2}.
To convert CLM's reported C:N ratio to $N_{\mass}$, we assumed a uniform leaf C fraction of 0.46.
We then divided this calculated $N_{\mass}$ by the reported SLA to obtain $N_{\area}$.
We calculated $V_{c,\max,\mass}$ by multiplying the reported $V_{c,\max,\area}$ by the reported SLA\@.

To test the hypothesis that the multivariate and hierarchical models offer more value in terms of uncertainty constraint at smaller sample sizes, we calculated the relative uncertainty ($\alpha$) as a function of the mean ($\mu$) and upper ($q_{0.975}$) and lower ($q_{0.025}$) confidence limits of trait estimates.

\[ \alpha = \frac{q_{0.975} - q_{0.025}}{\mu} \]

We then fit a generalized linear model relating relative uncertainty to sample size ($n$) for each of the model types (univariate, multivariate, and hierarchical; Fig.~\ref{fig:mvtraits-fig3}).

\[ \log{\alpha} = b_0 + b_1 \log{n} \]

If all three models performed equally well at all sample sizes, their respective slope and intercept coefficients would be statistically indistinguishable.
On the other hand, models that perform better should have
lower intercept ($b_0$) coefficients, indicating generally lower uncertainty,
and
lower slope ($b_1$) coefficients, indicating a reduced sensitivity of uncertainty ($\alpha$) to sample size ($n$).

To assess the consistency of within- and across-PFT trait trade-offs, we looked at covariance estimates for each trait pair and, where these values were significantly different from zero ($p < 0.05$),
we calculated the eigenvalues from the pairwise variance-covariance matrix for that trait pair and plotted the corresponding dominant eigenvectors centered on the mean estimates (Fig.~\ref{fig:mvtraits-fig4}).
This figure provides a visual representation of relative positions of PFTs in trait space and both the direction and extent of within-PFT trait covariance, and is directly analogous to conceptual figures describing hierarchical trait variability across environmental gradients as presented in, for instance,~\cite{cornwell_community_2009} and~\cite{albert_intraspecific_2010}.
Due to the small number of points used to estimate across-PFT covariance in the hierarchical model, none of the across-PFT covariances estimated in the hierarchical model were significantly different from zero ($p < 0.05$).
For this reason, we compared within-PFT covariances as estimated by the hierarchical model with the covariances estimated by fitting a multivariate model to all of the data.

Besides the consistency in the direction of trait covariance globally and between different PFTs, we also investigated the strength and predictive power of these covariances, which is represented by correlation coefficients (i.e.\ the pairwise covariance normalized to the variances of the component variables).
To do this, we plotted the mean and 95\% confidence interval of the pairwise trait correlation coefficients for the global estimate from the pooled multivariate model and PFT-level estimates from the hierarchical model (Fig.~\ref{fig:mvtraits-fig5}).

The R code and ancillary data for running these analyses is publicly available online via the Open Science Framework (OSF) at \texttt{https://osf.io/w8y73/}.
The TRY data used for this analysis can be requested at \texttt{http://try-db.org}.

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% End:

### 2018-09-12 3_results.tex
Initial commit (will not compile)
Brought in LaTeX from dissertation, plus some stuff from last submission.

\section{Results}\label{sec:mvtraits-results}

\subsection{Estimates of PFT-level means}

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{1_mvtraits/figures/mean_comparison.pdf}
  \caption{%
    Mean and 95\% confidence interval on best estimates of traits for each plant functional type from the univariate, multivariate, and hierarchical models.
    For leaf lifespan and SLA, results were not significantly different between the mass- and area-based models, so only results from the mass-based model are shown.
    For some PFT-trait combinations, where large error bars resulting from the relatively uninformative priors are substantially larger than the variability among means, the $y$ axes are constrained to facilitate comparison.
  }\label{fig:mvtraits-fig2}
\end{figure}

In general, leaf trait estimates from the univariate, multivariate, and hierarchical models were similar (Fig.~\ref{fig:mvtraits-fig2}).
Where we observed differences between models, the largest were between the univariate and multivariate models, while the additional constraint from the hierarchical model tended to have a minimal effect on trait estimates.
Significant differences in trait estimates between univariate and multivariate models occurred even for traits with relatively large sample sizes, such as leaf nitrogen content.

Evergreen PFTs had by far the largest leaf lifespan, with the longest lifespan observed for temperate and boreal needleleaf evergreen species.
Meanwhile, all of deciduous species had lifespans shorter than 7 months.
Among deciduous species, lifespan was generally longer in warmer biomes than colder ones.

Across-PFT patterns in SLA and $N_{\mass}$, $P_{\mass}$, and $R_{d,\mass}$ were similar.
Temperate broadleaved deciduous trees and shrubs generally had among the highest values of these traits, while temperate evergreen trees and shrubs had generally among the lowest.
However, none of these patterns were universal to all four traits.
For example, tropical evergreen trees had relatively high $N_{\mass}$ and mean SLA and $R_{d,\mass}$, but among the lowest $P_{\mass}$.
Similarly, temperate and boreal shrubs had higher $N_{\mass}$ and $P_{\mass}$ than any of the grasses, but comparable SLA\@.

Across-PFT patterns in $N_{\area}$, $P_{\area}$, and $R_{d,\area}$ were different from their mass-normalized counterparts.
For example, tropical broadleaved evergreen and needleleaf evergreen trees had among the lowest $N_{\mass}$ and $P_{\mass}$ basis but among the highest $N_{\area}$ and $P_{\area}$, while the opposite was true of deciduous temperate trees and shrubs.
Species with N contents near the middle of the observed range did not shift as dramatically depending on type of normalization.

C3 grasses had both the highest $V_{c,\max,\mass}$ and $V_{c,\max,\area}$.
Compared to broadleaved trees, temperate needleleaved evergreen trees had lower $V_{c,\max,\mass}$ but higher $V_{c,\max,\area}$.
Among broadleaved trees, deciduous trees had higher $V_{c,\max,\mass}$ and slightly higher $V_{c,\max,\area}$ than evergreen trees.
Between the deciduous and evergreen tree PFTs, we observed no significant trend by climate zone.

C3 grasses and temperate needleleaved evergreen trees had the highest $J_{\max,\area}$, but temperate broadleaved deciduous trees had the highest $J_{\max,\mass}$.
All of the shrub PFTs had the lowest $J_{\max,\mass}$ but average or above-average $J_{\max,\area}$, while the opposite was true of broadleaved tropical PFTs.
Of the tree PFTs, needleleaved evergreen trees had the highest $J_{\max,\area}$ but the lowest $J_{\max,\mass}$.

A key application of this study was to provide data-driven parameter estimates for Earth System Models.
To this end, we compared our mean parameter estimates with corresponding default parameters in CLM 4.5~\cite{clm45_note} (Fig.~\ref{fig:mvtraits-fig2}).
Our mean estimates of SLA agreed with CLM's defaults~\cite[Table 8.1 in]{clm45_note} only for tropical broadleaved evergreen trees, and for all other PFTs, our estimates are significantly lower.
For $N_{\mass}$, our estimates agreed reasonably well with CLM for evergreen temperate trees, needleleaved trees, and C3 arctic grasses, and were substantially different for all other PFTs.
Our $N_{\mass}$ estimates also varied much more across PFTs than CLM's parameters.
For $N_{\area}$, our estimates were significantly higher than CLM's for all PFTs, likely due to CLM's overestimates of SLA\@.
Our estimates of $V_{c,\max_\mass}$ were lower across all PFTs, with particularly large differences for tropical and temeprate broadleaf deciduous trees and evergreen shrubs, and temperate C3 grasses.
Our estimates of $V_{c,\max,\area}$ showed better agreement, though our values were still significantly lower for many PFTs.
Like us, Kattge et al.~(2009) \nocite{kattge_2009_vcmax} also found that $V_{c,\max,\area}$ was overestimated by Earth System models,
but their estimates of $V_{c,\max,\area}$ and $N_{\area}$ are generally slightly higher than ours.

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{1_mvtraits/figures/relative_ci_model.pdf}
  \caption{%
    Relative uncertainty in PFT-level trait estimates as a function of sample size for each model type.
    Lines represent linear models ($\log(y) = b_0 + b_1 \log(x)$) fit independently for each model type.
    In general, differences in estimate uncertainty between the univariate and multivariate models were minimal at large sample sizes but increasingly important at low sample sizes.
    However, differences in estimate uncertainty between the multivariate and hierarchical models were consistently negligible.
  }\label{fig:mvtraits-fig3}
\end{figure}

We observed clear differences in the relative uncertainties of mean estimates with respect to sample size.
All of the high-latitude PFTs consistently had among the largest error bars around their mean estimates relative to other PFTs, while the traits with the largest uncertainties were dark respiration, $V_{c,\max}$, and $J_{\max}$.
For many of these trait-PFT combinations, the additional constraint from trait covariance provided by the multivariate and hierarchical models substantially reduced error bars, making it possible to compare estimates against those of other PFTs.
Our analysis of the relationship between model type, sample size, and estimate relative uncertainty found that this covariance-based constraint from the multivariate model both reduced uncertainty overall (lower intercept) and reduced the sensitivity of estimate uncertainty to sample size (lower slope) compared to the univariate model (Fig.~\ref{fig:mvtraits-fig3}).
However, this analysis revealed no consistent significant benefit from the hierarchical model.

\subsection{Trait correlation patterns across- and within-PFTs}

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{1_mvtraits/figures/stick_pairs.pdf}
  \caption{%
    Pairwise trait mean and covariance estimates for all data pooled globally (black) and for each PFT (colored).
    Covariance estimates not significantly different from zero ($p < 0.05$) are indicated by x symbols at the mean estimate.
    $x$ and $y$ axes vary on a log scale, reflecting the fact that the model was fit using the base 10 log of all traits.
    With the exception of leaf lifespan, pairwise covariances are consistent in direction but vary somewhat in magnitude between PFTs, and when comparing PFT-level and global estimates.
    However, many pairwise covariances are not statistically significant, particularly (but not always) for undersampled traits and PFTs.
  }\label{fig:mvtraits-fig4}
\end{figure}

For all traits except leaf lifespan, pairwise trait correlations were generally consistent in direction both globally and within each PFT (Fig.~\ref{fig:mvtraits-fig4}).
In particular, mass- and area-normalized traits were all positively correlated with each other and, respectively, positively and negatively correlated with SLA, both globally and within each PFT\@.
The same was generally true of correlations of mass-based traits with leaf lifespan, but correlations of leaf lifespan with area-normalized traits were more variable.
The correlation between $N_{\area}$ and leaf lifespan was positive globally and for evergreen shrubs, tropical broadleaved deciduous trees, temperate needleleaved evergreen trees but negative for temperate and boreal broadleaved deciduous trees and not significant for any other PFTs.
Similarly, the correlation between $P_{\area}$ and leaf lifespan was positive globally but negative for evergreen shrubs and not significant for any other PFTs.
The correlation between leaf lifespan and $R_{d,\area}$ was significant and negative globally, but was not significant within any PFTs.
The only significant correlations of leaf lifespan with $V_{c,\max,\area}$ and $J_{\max,\area}$ were negative for temperate broadleaved deciduous trees.

A large number of pairwise trait correlations were not significant.
In some cases, this was driven by sample size (Fig.~\ref{fig:mvtraits-fig1}).
For instance, needleleaved deciduous trees, the most undersampled PFT in our analysis, were often the only PFT for which a correlation was not statistically significant.
In other cases, though, PFTs with smaller sample sizes had significant pairwise correlations while PFTs with much larger sample sizes had none.
For example, tropical broadleaved evergreen trees were relatively well-sampled for all traits, but none of their area-normalized traits were significantly correlated with leaf lifespan.
In general, we observed fewer significant trait correlations among area-normalized traits than mass-normalized traits.

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{1_mvtraits/figures/correlation_boxplot.pdf}
  \caption{%
    Mean and 95\% CI on estimates of pairwise correlation coefficients
    for all data pooled globally (dark grey) and for each PFT (colored).
    For most PFT-trait pairs, correlations are mutually consistent in magnitude but vary in strength.
  }\label{fig:mvtraits-fig5}
\end{figure}

The strength of pairwise trait correlations varied substantially depending on scale, PFT, and trait (Fig.~\ref{fig:mvtraits-fig5}).
The two pairwise trait correlations that exhibited the most consistent strength globally and within each PFT were the correlation between SLA and $N_{\area}$, and between $N_{\mass}$ and $P_{\mass}$.
Correlation strength was often, but not always, related to sample size, with well-sampled PFTs exhibiting stronger correlations and undersampled PFTs exhibiting weaker correlations.

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% End:

### 2018-09-12 4_discussion.tex
Initial commit (will not compile)
Brought in LaTeX from dissertation, plus some stuff from last submission.

\section{Discussion}

\subsection{Scale dependence of the leaf economic spectrum}

The canonical leaf economic spectrum is defined by a negative correlation of between SLA and leaf lifespan, and a positive correlation of SLA with $N_{\mass}$, $P_{\mass}$, and photosynthesis and respiration rates~\cite{wright_worldwide_2004}.
The first objective of this paper was to investigate the extent to which these relationships hold within and across PFTs.
Our results indicate that the leaf economic spectrum generally still holds within PFTs, at least at the functional and phylogenetic scale of the current generation of Earth System Models.
Within PFTs, correlations between SLA, $N_{\mass}$, and $P_{\mass}$ were consistently positive, and correlations of these traits with leaf lifespan were generally negative (though, for many PFTs, correlations were not significantly different from zero).
Although we did not include maximum photosynthesis rate ($A_{\max}$), $V_{c,\max,\mass}$ and $J_{\max,\mass}$ generally exhibited the expected positive correlations with SLA and negative correlations with leaf lifespan, as did $R_{d,\mass}$, though many correlations were not significant, primarily due to sample size.

While the direction of relationships we observed was largely consistent across different PFTs, the strength of these correlations was more variable.
For example, the correlations of SLA with $N_{\mass}$ and $P_{\mass}$ were generally weaker in needleleaved PFTs than in broadleaved PFTs.
On the other hand, relationships of SLA and $N_{\area}$ were strongly negative for all PFTs (except the data-limited needleleaved deciduous trees), and temperate needleleaved species in particular had among the strongest SLA-$N_{\area}$ and SLA-$P_{\area}$ correlations of any PFT\@.
Taken together with the finding that evergreen conifers are relatively unique in their consistently proportional allocation of N to cell walls and Rubisco~\cite{onoda_physiological_2017}, our results support the idea that needleleaved species primarily adapt to environmental changes through structural changes (i.e.\ increasing or decreasing SLA) rather than through modulating foliar biochemistry~\cite{robakowski_2004_needle}.

We also found that the strength of the correlations of $N$ with photosynthetic parameters $V_{c,\max}$ and $J_{\max}$--often used in ecosystem modeling to parameterize photosynthesis~\cite{clm45_note,rogers_roadmap_2017}--varied by plant functional type, whether the traits were normalized by mass or area.
Although trait correlations are not necessarily indicative of allocation strategies, this result generally supports the findings of Ghimire et al.~(2017) \nocite{ghimire_2017_Nallocation} that N allocation strategies to photosynthesis vary widely by plant functional type.
In particular, the correlations of $N_{\mass}$ with traits related to photosynthesis and respiration ($R_{d,\mass}$, $V_{c,\max,\mass}$, and $J_{\max,\mass}$) were weaker for tropical evergreen broadleaved trees than for most other PFTs, but the correlation of these traits with $P_{\mass}$ was comparable or even slightly higher.
This suggests that the productivity of tropical species is generally P-limited~\cite{reich_global_2004,ghimire_2017_Nallocation}, that N allocation strategies are more variable under N-poor than N-rich conditions~\cite{ghimire_2017_Nallocation}, or more generally that photosynthetic parameters are affected more strongly by environmental covariates than leaf nitrogen contents~\cite{Ali_2015}.
Meanwhile, our result that the $N_{\area}$---$V_{c,\max,\area}$ correlation is generally weaker for needleleaved than broadleaved species was also found by Kattge et al.~(2009), and supports the lower allocation of N to photosynthesis found by Ghimire et al.~(2017). \nocite{ghimire_2017_Nallocation,kattge_2009_vcmax}
In light of this fact and considering that boreal forests, largely populated by needleleaf species, have the largest influence on global climate of any biome~\cite{Snyder_2004,bonan_forests_2008}, we suggest that parameterization of needleleaf tree productivity based on foliar nitrogen content in Earth System Models be treated with caution.

Correlations of all traits with leaf lifespan were weaker (often to the point of becoming insignificant) within most PFTs than globally.
This suggests that leaf economic relationships related to leaf lifespan are dominated by the fundamental differences between deciduous and evergreen PFTs, while the factors driving variability in leaf lifespan within PFTs are more complex and idiosyncratic~\cite{Reich_2014,wu_leaf_2016}.
That being said, much of this within-PFT variability is known to be driven by variations in shade responses, and a key limitation of our study is that we have no information about the relative canopy positions at which traits were collected~\cite{lusk_why_2008,keenan_global_2016}.

Across PFTs, the interaction between growth form and biome in PFT definitions (Table 1) confounds the interpretation of our results with respect to well established ecological and biogeographic patterns.
For example, Poorter et al.~(2009) \nocite{poorter_2009_causes} found variation in SLA by growth form was, from highest to lowest, grasses, deciduous woody plants, evergreen woody plants, and succulents, while the pattern by biome was, again from highest to lowest, grassland, tundra, tropical forest, temperate forest, woodland, shrubland, and desert.
We observed as expected that arctic grasses had lower mean SLA than temperate grasses, and that evergreen trees had lower SLA than their deciduous counterparts.
However, we observed by far the highest mean SLA values in temperate deciduous broadleaf trees, rather than in grass PFTs as expected.
Similarly to Onoda et al.~(2011), we found no consistent patterns in SLA with temperature: \nocite{onoda_2011_global}
Among broadleaved evergreen PFTs, temperate species had lower SLA than tropical, but among broadleaved deciduous PFTs, temperate species had higher SLA than both tropical and boreal species.
With respect to the expected decline in foliar N:P ratios with latitude~\cite{reich_global_2004}, we found that $N_{\mass}$ was generally higher in PFTs associated with colder biomes compared to warmer ones, but observed biome-related differences in $P_{\mass}$ only among deciduous broadleaved and needleleaved evergreen PFTs.
Contrary to Atkin et al.~(2015) \nocite{atkin_global_2015}, our results for both $R_{d,\mass}$ and $R_{d,\area}$ failed to show a trend with respect to biome.
However, this comparison may not be entirely fair because our study design inherently averages over the extensive climatic variability within PFTs.

Finally, there has been some debate in the trait ecology community about the use of mass- or area-normalized traits in analyses of the leaf economic spectrum.
Two studies~\cite{osnas_global_2013,lloyd_les} independently concluded that leaf economic relationships among mass-based traits emerge inevitably out of variation in SLA and are therefore not ecologically meaningful.
Responses to these criticisms have suggested that both mass- and area-based normalization have merit, as mass-based traits have a natural interpretation in terms of investment costs and allocation while area-based traits are important due to the fundamentally area-based nature of energy and gas fluxes through leaf surfaces~\cite{westoby_lloyd_response,poorter_les_response}.
Our study suggests the latter, that investigation of trait correlations on both a mass- and area-basis can yield biologically and ecologically meaningful conclusions.
For one, our discussion of differences in ecological strategies between broadleaved and needleaved species fundamentally depends on comparative analysis of mass- and area-normalized nutrient contents.
Meanwhile, our discussion of tropical tree productivity with respect to foliar nutrient contents is generally supported by both mass- and area-normalized traits (i.e.\ the correlation between $P$ and $V_{c,\max}$ is stronger than the correlation of $N$ and $V_{c,\max}$ for tropical species whether we use the mass- or area-normalized versions of these traits).

\subsection{Covariance as constraint}

The second objective of this paper was to investigate the ability of trait covariance to provide additional information that could be used to reduce uncertainties in trait estimates.
In agreement with our expectations, our results show that accounting for trait covariance constrained the uncertainty around PFT-level trait means, particularly for trait-PFT combinations with low sample sizes (Fig.~\ref{fig:mvtraits-fig2} and~\ref{fig:mvtraits-fig3}).
Moreover, accounting for trait covariance occasionally also changed the position of trait mean estimates, even for PFT-trait combinations with relatively large sample sizes (e.g. $N_{\mass}$ for temperate broadleaved deciduous trees, Fig.~\ref{fig:mvtraits-fig2}).
This result echoes Diaz et al.~(2016) \nocite{diaz_global_2016} in demonstrating the importance of studying the multivariate trait space rather than individual traits. 
Such shifts suggest that the sampling of these traits in the TRY database is not representative, which is evident from looking at the relative sample sizes of different traits (Fig.~\ref{fig:mvtraits-fig1}; see also Kattge et al.~2011). \nocite{kattge_try_2011} 
These shifts also indicate that parameter estimates based on univariate trait data~\cite[e.g.,][]{lebauer_facilitating_2013,dietze_quantitative_2014,butler_2017_mapping} may not only be overestimating uncertainty, but may also be systematically biased.
Although our results clearly show that many time- and labor-intensive traits, such as $R_{d}$, $V_{c,\max}$, and $J_{\max}$, still lack the observations to estimate covariance with other traits for certain PFTs,
our results also show that the effective sample size of all traits is enhanced when covariance is taken into account.
For example, field and remote sensing studies that estimate only certain traits like SLA and $N_{\mass}$ but not others (such as $P_{\mass}$ and $R_{d,\mass}$) can leverage trait correlations to provide indirect constraint~\cite{serbin_spectroscopic_2014,singh_imaging_2015,musavi_imprint_2015,lepine_2016_examining}.
We therefore suggest that future observational campaigns consider trait covariance when deciding which traits to measure.

The additional benefit of hierarchical multivariate modeling in our study was limited, largely due to the relatively low number of points used to estimate that across-PFT covariance structure.
Therefore, for parameterizing the current generation of ecosystem models using well-sampled traits, we suggest that simple multivariate models fit independently to each PFT are sufficient and that the additional conceptual challenges and computational overhead of hierarchical modeling are not required.
However, for modeling work that requires larger numbers of PFTs~\cite{boulangeat_improving_2012}, and especially for models that are parameterized on the basis of individual species~\cite[e.g. Linkages,][]{linkages}, the benefits of hierarchical modeling may accumulate~\cite{clark_why_2016,dietze_capturing_2008,cressie_accounting_2009,webb_structured_2010}.

More generally, we foresee tremendous potential of multivariate and hierarchical modeling approaches to elucidating the role of functional traits in organismal and ecosystem function. 
Besides the many important foliar traits that we did not include (e.g.\ pigment contents, leaf hydraulic traits), our approach could readily be applied to other plant traits whose relationship to the leaf economic spectrum is less clear. 
One example is hydraulic traits:
While stem and leaf hydraulic traits are correlated~\cite{bartlett_hydraulic}, a potential scaling between hydraulic and leaf economic traits is poorly understood~\cite{reich_world-wide_2014,li_leaf_2015}. 
Similarly, reexamining the relationships defining wood~\cite{chave_woodeconomics,baraloto_decoupled_2010,fortunel_leaf_2012} and root~\cite{kramer-walter_root_2016,valverde-barrantes_root_2016} economic spectra, as well as their relationship to the foliar traits, would provide useful information on possible scale-dependence of plant growth and allocation strategies.
We emphasize that the relative difficulty of measuring hydraulic and other non-foliar traits~\cite[e.g.][]{jansen_hydraulic_traits} further increases the value of any technique that can fully leverage the information they provide. 
More generally, multivariate and hierarchical modeling has the potential to reveal functional trade-offs that are mutually confounding at different scales, thereby enhancing our understanding of processes driving functional diversity.

\subsection{Conclusions}

The tremendous functional diversity of plants continues to be a major challenge for functional ecology and ecosystem modeling.
Functional diversity research fundamentally depends on dimensionality reduction through a search for meaningful pattern that can be exploited to take reasonable guesses at average behavior. 
The trait trade-offs comprising the leaf economic spectrum are one such pattern. 
In this paper, we reaffirm the existence of the leaf economic spectrum both globally and, with some caveats, within plant functional types typically used in the current generation of Earth System Models.
We also highlight how the strength of leaf economic relationships can be influenced by biotic and abiotic factors specific to certain PFTs.
Finally, we show how patterns of trait covariance like the leaf economic spectrum can be leveraged to inform trait estimates, particularly at small sample sizes.

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% End:

### 2018-09-12 main.tex
Initial commit (will not compile)
Brought in LaTeX from dissertation, plus some stuff from last submission.

\documentclass{article}

\title{My title}
\author{Alexey N. Shiklomanov\(^1\), Elizabeth M. Cowdery\(^1\), Michael
  Bahn\(^2\), Chaeho Byun\(^3\), Joseph Craine\(^4\), Andrés
  Gonzalez-Melo\(^5\), Steven Jansen\(^6\), Nathan Kraft\(^7\), Koen
  Kramer\(^{8}\), Vanessa Minden\(^{9,10}\), Ülo Niinemets\(^{11}\),
  Yusuke Onoda\(^{12}\), Enio Egon Sosinski\(^{13}\), Nadejda A.
  Soudzilovskaia\(^{14}\), Michael C. Dietze\(^1\)}

\begin{document}

% Affiliations
\(^1\)Department of Earth \& Environment, Boston University, Boston, MA
02215, USA;
\(^2\)Institute of Ecology, University of Innsbruck, 6020
Innsbruck, Austria;
\(^3\)School of Biological Sciences, Seoul National
University, Gwanak-ro 1, 08826, Seoul, Seoul, South Korea;
\(^4\)Jonah
Ventures, Manhattan, KS 66502, USA;
\(^5\)Universidad del Rosario.
Facultad de Ciencias Naturales y Matemáticas. Carrera 26 \#63B-48.
Bogotá, Colombia;
\(^6\)Institute of Systematic Botany and Ecology, Ulm
University, Ulm, Germany;
\(^7\)Department of Ecology and Evolutionary
Biology, University of California Los Angeles, Los Angeles, CA 90095,
USA;
\(^{8}\)Vegetation, Forest, and Landscape Ecology, Wageningen
Environmental Research and Wageningen University, Wageningen,
Gelderland, The Netherlands;
\(^{9}\)Institute for Biology and
Environmental Sciences, Carl von Ossietzky-University of Oldenburg, Carl
von Ossietzky Str. 9-11, 26129 Oldenburg, Germany;
\(^{10}\)Department
of Biology, Ecology and Evolution, Vrije Universiteit Brussel, Pleinlaan
2, 1050 Brussels;
\(^{11}\)Institute of Agricultural and Environmental
Sciences, Estonian University of Life Sciences, Kreutzwaldi 1, 51014
Tartu, Estonia;
\(^{12}\)Graduate School of Agriculture, Kyoto
University, Kyoto, 606-8502, Japan;
\(^{13}\)Embrapa Temperate
Agriculture, Rodovia BR 392, Km 78, CEP 96010-971 Pelotas, RS, Brazil;
\(^{14}\)Conservation Biology Department, Institute of Environmental
Sciences, Leiden University, Leiden, The Netherlands

Corresponding author:

Alexey Shiklomanov,

telephone: +1 (302) 897-3348,

email: ashiklom@bu.edu

Key words: Functional trade-off; hierarchical modeling; trait variation;
ecological modeling; leaf morphology; leaf biochemistry

\begin{longtable}[]{@{}lll@{}}
\toprule
Section & Word count &\tabularnewline
\midrule
\endhead
Introduction & 1060 & Figures: 5 (all colored)\tabularnewline
Materials and Methods & 1695 & Tables: 2\tabularnewline
Results & 1108 & Supplementary figures: 2\tabularnewline
Discussion & 1907 & Supplementary tables: 4\tabularnewline
Acknowledgements & 88 & Supplementary methods: 1\tabularnewline
Total & 5858 &\tabularnewline
\bottomrule
\end{longtable}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:

### 2019-03-20 ELEshiklomanovSA2.Rmd
Fix Supplementary Information Rmd file

---
title: Supporting information
author:
  - >
    Alexey N. Shiklomanov,
    Elizabeth M. Cowdery,
    Michael Bahn,
    Chaeho Byun,
    Joseph Craine,
    Andrés Gonzalez-Melo,
    Steven Jansen,
    Koen Kramer,
    Vanessa Minden,
    Ülo Niinemets,
    Yusuke Onoda,
    Enio Egon Sosinski,
    Nadejda A. Soudzilovskaia,
    Michael C. Dietze
output: pdf_document
header-includes:
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage{wrapfig}
  - \usepackage{float}
  - \usepackage{colortbl}
  - \usepackage{pdflscape}
  - \usepackage{tabu}
  - \usepackage{threeparttable}
  - \usepackage{threeparttablex}
  - \usepackage[normalem]{ulem}
  - \usepackage{makecell}
  - \renewcommand{\thetable}{S\arabic{table}}
  - \renewcommand{\thefigure}{S\arabic{figure}}
---

```{r libraries, include = FALSE}
library(knitr)
library(kableExtra)
library(drake)
library(ggplot2)
library(GGally)

correlation_scatter_cap <- paste(
  "Relationship between squared pairwise correlation and sample size of pairwise observations for each PFT.",
  "Grey shading is the 95% confidence interval around a local regression (LOESS).",
  "Upper triangle is area-normalized values, and lower triangle is mass-normalized values.",
  "Note that the x-axis uses a $\\log_{10}$ scale."
)

correlation_fig_cap <- paste(
  "Fraction of statistically significant ($p < 0.05$) pairwise trait correlation estimates",
  "as a function of sample size for each plant functional type."
)

correlation_range_cap <- paste(
  "Relationship between pairwise correlation strength and data range for each PFT.",
  "In each pair of sub-panels, the left panel is relationship between pairwise correlation",
  "and the range of the variable on the x axis",
  "while the right panel is the relationship between pairwise correlation",
  "and the y variable",
  "(For instance, in the two-panel figure in row 1 column 2,",
  "the left panel is the SLA-leaf lifespan correlation as a function of the range of SLA",
  "and the right panel is the same correlation as a function of the range of leaf lifespan).",
  "Data range is expressed as the difference between the 2.5% and 97.5% quantiles of each normalized trait (i.e. $\\frac{X}{E[X]}$).",
  "Grey shading is the 95% confidence interval around a local regression (LOESS).",
  "Upper triangle is area-normalized values, and lower triangle is mass-normalized values."
)
```

```{r mean_value_table, echo = FALSE, results = "asis"}
table_cap_mass <- paste(
  "Mean and 95\\% confidence interval of trait estimates",
  "for mass-normalized traits from the hierarchical model.",
  "Note that some traits for some PFTs have effectively no constraint relative to the mostly uninformative prior (as evidenced by their very large confidence intervals),",
  "so their values should be used with caution."
)
table_cap_area <- gsub("mass", "area", table_cap_mass)

readd(table_dat) %>%
  dplyr::select(-dplyr::matches("area")) %>%
  kable(
    caption = table_cap_mass,
    format = "latex",
    col.names = linebreak(c(
      "PFT",
      "Leaf lifespan\n(months)",
      "SLA\n(kg m$^{-2}$)",
      "$N_\\textrm{mass}$\n(mg g$^{-1}$)",
      "$P_\\textrm{mass}$\n(mg g$^{-1}$)",
      "$R_\\textrm{d, mass}$\n($\\mu$mol g$^{-1}$ s$^{-1}$)",
      "$V_\\textrm{c, max, mass}$\n($\\mu$mol g$^{-1}$ s$^{-1}$)",
      "$J_\\textrm{max, mass}$\n($\\mu$mol g$^{-1}$ s$^{-1}$)"
    )),
    escape = FALSE,
    booktabs = TRUE
  ) %>%
  kableExtra::kable_styling(latex_options = c("scale_down", "hold_position"))

readd(table_dat) %>%
  dplyr::select(-dplyr::matches("mass")) %>%
  kable(
	caption = table_cap_area,
	format = "latex",
    col.names = linebreak(c(
      "PFT",
      "Leaf lifespan\n(months)",
      "SLA\n(kg m$^{-2}$)",
      "$N_\\textrm{area}$\n(g m$^{-2}$)",
      "$P_\\textrm{area}$\n(g m$^{-2}$)",
      "$R_\\textrm{d, area}$\n($\\mu$mol m$^{-2}$ s$^{-1}$)",
      "$V_\\textrm{c, max, area}$\n($\\mu$mol m$^{-2}$ s$^{-1}$)",
      "$J_\\textrm{max, area}$\n($\\mu$mol m$^{-2}$ s$^{-1}$)"
    )),
	escape = FALSE,
	booktabs = TRUE
) %>%
  kableExtra::kable_styling(latex_options = c("scale_down", "hold_position"))
```

```{r sample_size_table, echo = FALSE, results = "asis"}
ssw <- readd(sample_size_wide)
kable(
  ssw %>%
    dplyr::select(pft, leaf_lifespan, SLA,
                  Nmass, Narea, Pmass, Parea, Rdmass, Rdarea,
                  Vcmax_mass, Vcmax_area, Jmax_mass, Jmax_area),
  col.names = c(
    "PFT", "Leaf lifespan", "SLA",
    "$N_\\textrm{mass}$", "$N_\\textrm{area}$",
    "$P_\\textrm{mass}$", "$P_\\textrm{area}$",
    "$R_\\textrm{d, mass}$", "$R_\\textrm{d, area}$",
    "$V_\\textrm{c, max, mass}$", "$V_\\textrm{c, max, area}$",
    "$J_\\textrm{max, mass}$", "$J_\\textrm{max, area}$"
  ),
  format = "latex",
  booktabs = TRUE,
  escape = FALSE,
  caption = "The number of non-missing (outside parentheses) and imputed (inside parentheses) observations of each trait for each plant functional type used in this analysis."
) %>%
  kable_styling(latex_options = c("scale_down", "hold_position")) %>%
  group_rows("Broadleaf", 2, 6) %>%
  group_rows("Needleleaf", 7, 9) %>%
  group_rows("Shrub", 10, 12) %>%
  group_rows("Grass", 13, 15) %>%
  landscape()
```

```{r pairwise_correlation_table, echo = FALSE, results = "asis"}
corr_results <- readd(corr_processed)
kable(
  corr_results %>%
    dplyr::select(pft, yvar, xvar, corr_value, present, missing),
  "latex",
  booktabs = TRUE, longtable = TRUE, escape = TRUE,
  col.names = c("PFT", "Trait 1", "Trait 2", "Correlation (95% CI)", "present", "missing"),
  caption = "Pairwise trait correlation values for each plant functional type. Values with asterisks indicate correlations significantly different from zero ($p < 0.05$)."
)
```

```{r correlation_sample_size_pairs, echo = FALSE, results = "hide", warning = FALSE, fig.cap = correlation_scatter_cap, fig.dim = c(12, 12)}
readd(corr_ss_plot_gg)
```

```{r correlation_range_analysis, echo = FALSE, results = "hide", warning = FALSE, fig.cap = correlation_range_cap, fig.dim = c(12, 12)}
readd(corr_range_plot_gg)
```

\clearpage

# Method S1: Estimating trait means and covariances through multiple imputation {-}

## Introduction {-}

Under _single_ imputation, the imputation step occurs once outside of the MCMC loop.
In _multiple_ imputation, as implemented in this paper, the imputation step is part of the MCMC loop,
such that imputed data values are conditioned on the current state of the parameters and vice-versa.
In the figure below,
$Y$ is the original data (with missing values),
$Y^*$ is the original data with missing values imputed,
$\mu_p$ and $\Sigma_p$ are values of the mean vector and covariance matrix (respectively) calculated only from the original data,
and $\mu^*_t$ and $\Sigma^*_t$ are the draws of the mean vector and covariance matrix (respectively) at MCMC iteration $t$.

```{tikz diagram, echo = FALSE}
\usetikzlibrary{positioning, calc, fit}
\begin{tikzpicture}[
	every node/.style = {shape = rectangle, draw = black},
	scale = 1.1
]
\node (Y) at (-1, 3) {Data $Y$};
\node (data params) at (-1, 2) {Calculate $\mu_p, \Sigma_p$};
\node (Ystar) at (-1, 1) {Impute $Y^* \mid \mu_p, \Sigma_p$};
\node (guess params) [align = center, right = of Ystar] {Initial guess\\$\mu^*_0$ $\Sigma^*_0$};
\node (mu star) at (0, 0) {Draw $\mu^*_t \mid Y^*, \Sigma^*_{t-1}$};
\node (sigma star) at (0, -1) {Draw $\Sigma^*_t \mid Y^*, \mu^*_t$};
\path [->] (Y) edge (data params);
\path [->] (data params) edge (Ystar);
\path [->] (Ystar) edge (mu star);
\path [->] (guess params) edge (mu star);
\path [->] (mu star) edge (sigma star);
\path [->, out = 0, in = 0] (sigma star.east) edge
	node (t plus one) [right, draw = none, font = {\footnotesize}] {$t = t + 1$}
	(mu star.east);
\node (mcmc loop) [fit = (mu star)(sigma star)(t plus one), shape = rectangle, draw = black!50] {};
\node (mcmc loop text) [align = center, anchor = west, draw = none] at (mcmc loop.east) {MCMC\\loop};
\node (title) [anchor = north, draw = none] at (mcmc loop.south) {Single imputation};
\end{tikzpicture}
\hspace{1cm}
\begin{tikzpicture}[
	every node/.style = {shape = rectangle, draw = black},
	scale = 1.1
]
\node (Y) at (-1, 1) {Data $Y$};
\node (guess params) [align = center] at (1, 1) {Initial guess\\$\mu^*_0$ $\Sigma^*_0$};
\node (Ystar) at (0, 0) {Impute $Y^*_t \mid Y, \mu^*_{t-1}, \Sigma^*_{t-1}$};
\node (mu star) at (0, -1) {Draw $\mu^*_t \mid Y^*_t, \Sigma^*_{t-1}$};
\node (sigma star) at (0, -2) {Draw $\Sigma^*_t \mid Y^*_t, \mu^*_t$};
\path [->] (Y) edge (Ystar);
\path [->] (guess params) edge (Ystar);
\path [->] (Ystar) edge (mu star);
\path [->] (mu star) edge (sigma star);
\path [->, out = 0, in = 0] (sigma star.east) edge
	node (t plus one) [right, draw = none, font = {\footnotesize}] {$t = t + 1$}
	(Ystar.east);
\node (mcmc loop) [fit = (Ystar)(mu star)(sigma star)(t plus one), shape = rectangle, draw = black!50] {};
\node (mcmc loop text) [align = center, anchor = west, draw = none] at (mcmc loop.east) {MCMC\\loop};
\node (title) [anchor = north, draw = none] at (mcmc loop.south) {Multiple imputation};
\end{tikzpicture}
```

## Demonstration {-}

Consider two positively correlated traits, A and B.

```{r create_traits}
set.seed(12345) # For reproducibility
library(mvtraits)

true_mu <- c(0, 0)
true_cov <- matrix(c(1, 0.8, 0.8, 1), nrow = 2)
Y_all <- random_mvnorm(1000, true_mu, true_cov)
colnames(Y_all) <- c("A", "B")
plot(Y_all, pch = 19)
```

Simulate missingness by randomly removing half of the data.

```{r remove_na}
Y <- Y_all
Y[sample.int(2000, 1000)] <- NA
plot(Y_all, pch = 19, col = "grey")
points(Y, pch = 19, col = "black")
legend(
  "topleft",
  legend = c("True missing", "Present"),
  col = c("grey", "black"),
  pch = 19,
  bg = "white"
)
```

Set an uninformative multivariate normal prior on $\mu$...
```{r prior_mu}
mu_0 <- c(0, 0)
Sigma_0 <- diag(10, 2)
```

...and an uninformative Wishart prior on $\Sigma$.
```{r prior_sigma}
v0 <- 2
S0 <- diag(10, 2)
```

Take an initial guess at the mean and covariance by drawing from their priors.

```{r initial}
mu_star <- random_mvnorm(1, mu_0, Sigma_0)[1,]
mu_star
Sigma_star <- rWishart(1, v0, S0)[,,1]
Sigma_star
```

Impute values based on this initial guess.

```{r impute}
Ystar <- mvnorm_fill_missing(Y, mu_star, Sigma_star)

mylegend <- function() {
  legend(
    "topleft",
    legend = c("True missing", "Present", "Imputed"),
    col = c("grey", "black", "red"),
    pch = 19,
	bg = "white"
  )
}
plot(Y_all, pch = 19, col = "grey")
points(Ystar, pch = 19, col = "red")
points(Y, pch = 19, col = "black")
mylegend()
```

Clearly, these initial values are a bad fit to the data.
Nevertheless, draw $\mu$ and $\Sigma$ conditioned on this set of imputed values.

```{r draw}
mu_star <- draw_mu(Ystar, Sigma_star, mu_0, Sigma_0)
mu_star
Sigma_star <- draw_Sigma(Ystar, mu_star, v0, S0)
Sigma_star
```

Draw a new set of imputed values, conditioned on the current $\mu^*$ and $\Sigma^*$.

```{r impute_2}
Ystar <- mvnorm_fill_missing(Y, mu_star, Sigma_star)
plot(Y_all, pch = 19, col = "grey")
points(Ystar, pch = 19, col = "red")
points(Y, pch = 19, col = "black")
mylegend()
```

This is still a bad fit to the data, but, though it is hard to tell, it has improved slightly.
Doing this repeatedly (in a loop) gradually improves the estimates of the covariance.

```{r loop}
par(mfrow = c(5, 4), mar = c(3, 3, 2, 0.1), cex = 0.4)
for (i in 1:20) {
  mu_star <- draw_mu(Ystar, Sigma_star, mu_0, Sigma_0)[1,]
  Sigma_star <- draw_Sigma(Ystar, mu_star, v0, S0)
  Ystar <- mvnorm_fill_missing(Y, mu_star, Sigma_star)
  title <- sprintf("iteration: %d, covariance: %.2f", i, Sigma_star[2, 1])
  plot(Y_all, pch = 19, col = "grey", xlab = "", ylab = "", main = title)
  points(Ystar, pch = 19, col = "red")
  points(Y, pch = 19, col = "black")
}
```

After a lot of MCMC samples, we can generate a distribution of covariance values, which provides an uncertainty estimate on the covariance.

```{r cov_estimate}
cov_ab <- numeric(1000)
for (i in seq_along(cov_ab)) {
  mu_star <- draw_mu(Ystar, Sigma_star, mu_0, Sigma_0)[1,]
  Sigma_star <- draw_Sigma(Ystar, mu_star, v0, S0)
  Ystar <- mvnorm_fill_missing(Y, mu_star, Sigma_star)
  cov_ab[i] <- Sigma_star[2, 1]
}
hist(cov_ab, xlab = "Covariance estimate", ylab = "Count", main = "Covariance estimate")
abline(v = true_cov[2, 1], col = "red", lwd = 2)
legend("topright", legend = "True cov.", lty = "solid", lwd = 2, col = "red")
```

### 2019-04-01 EAshiklomanovSA2.Rmd
Fix author list in supplement 2
Formatting fixes for submission

--- 2019-03-20 ELEshiklomanovSA2.Rmd
@@
     Chaeho Byun,
-    Joseph Craine,
-    Andrés Gonzalez-Melo,
     Steven Jansen,
@@
     Yusuke Onoda,
-    Enio Egon Sosinski,
     Nadejda A. Soudzilovskaia,
@@
   "Relationship between pairwise correlation strength and data range for each PFT.",
-  "In each pair of sub-panels, the left panel is relationship between pairwise correlation",
+  "In each pair of sub-panels, the top panel is relationship between pairwise correlation",
   "and the range of the variable on the x axis",
-  "while the right panel is the relationship between pairwise correlation",
+  "while the bottom panel is the relationship between pairwise correlation",
   "and the y variable",
   "(For instance, in the two-panel figure in row 1 column 2,",
-  "the left panel is the SLA-leaf lifespan correlation as a function of the range of SLA",
-  "and the right panel is the same correlation as a function of the range of leaf lifespan).",
+  "the top panel is the SLA-leaf lifespan correlation as a function of the range of SLA",
+  "and the bottom panel is the same correlation as a function of the range of leaf lifespan).",
   "Data range is expressed as the difference between the 2.5% and 97.5% quantiles of each normalized trait (i.e. $\\frac{X}{E[X]}$).",

### 2019-08-23 EAshiklomanovSA2.Rmd
Add bootstrap to SA2

@@
 library(mvtraits)
-
 true_mu <- c(0, 0)
@@
 Y <- Y_all
-Y[sample.int(2000, 1000)] <- NA
+ymiss <- sample.int(2000, 1000)
+# Save true values for validation
+ymiss_true <- Y[ymiss]
+Y[ymiss] <- NA
 plot(Y_all, pch = 19, col = "grey")
@@
 ```
+
+The `mvtraits` package provides the convenience function `fit_mvnorm` for performing this entire procedure automatically.
+
+```{r}
+results <- fit_mvnorm(Y)
+results$stats$mu$Mean
+true_mu
+results$stats$Sigma$Mean
+true_cov
+```
+
+It also provides a function for bootstrapping the results of a model fit this way:
+
+```{r}
+data_boot <- bootstrap_missing(results, Y)
+boot_means <- apply(data_boot, c(1, 2), mean)[ymiss]
+plot(ymiss_true ~ boot_means, pch = 19,
+     xlab = "Imputed values (bootstrapped mean)",
+     ylab = "True missing values")
+abline(0, 1, lty = "dashed")
+```
+
+We can see that for all data points where at least one observation was available as constraint, the algorithm imputed values that were close to the true values.
+However, where values for both traits were missing, the algorithm falls back to the global mean, which has the same statistical properties as the missing data (i.e.
+central tendency and spread) but doesn't match the true values.

### 2019-08-23 EAshiklomanovSA2.Rmd
Add more detailed prior information

@@
 library(GGally)
+stopifnot(
+  requireNamespace("shiklomanov2017np", quietly = TRUE),
+  requireNamespace("dplyr", quietly = TRUE),
+  requireNamespace("tibble", quietly = TRUE)
+)
 
@@
 )
+```
+
+```{r echo = FALSE, results = "asis"}
+prior_df <- tibble::tribble(
+  ~param, ~mean, ~stdev,
+  "Jmax_area", 2, 0.19,
+  "Jmax_mass", 0, 0.26,
+  "leaf_lifespan", 0.8, 0.4,
+  "Narea", 0.25, 0.3,
+  "Nmass", 1.25, 0.3,
+  "Parea", -1, 0.3,
+  "Pmass", 0, 0.3,
+  "Rdarea", log10(0.0011), 0.15,
+  "Rdmass", log10(0.8), 0.20,
+  "SLA", 1, 0.6,
+  "Vcmax_mass", log10(0.45), 0.2,
+  "Vcmax_area", log10(50), 0.18
+) %>%
+  dplyr::mutate(variance = stdev ^ 2)
+
+prior_df %>%
+  dplyr::mutate(param = factor(param, shiklomanov2017np::both_params)) %>%
+  dplyr::arrange(param) %>%
+  dplyr::select(Trait = param, "Lognormal Mean" = mean, "Lognormal SD" = stdev) %>%
+  knitr::kable(
+    format = "latex",
+    booktabs = TRUE,
+    caption = paste(
+      "Parameters for trait-specific prior distributions.",
+      "For the multivariate and hierarchical distributions,",
+      "the variance-covariance matrix prior was a diagonal matrix containing the",
+      "square of the lognormal standard deviation values (i.e. the variance)",
+      "shown here and all off-diagnoal terms set to zero." 
+    )
+  ) %>%
+  kableExtra::kable_styling(latex_options = c("scale_down", "hold_position"))
 ```

### 2019-08-23 EAshiklomanovSA2.Rmd
Address some more minor comments

@@
 
-## Demonstration {-}
+## Demonstration: Simulated bivariate data {-}
 
@@
 ```{r remove_na}
-Y <- Y_all
-ymiss <- sample.int(2000, 1000)
+Ym <- Y_all
+
+# Remove half of the data
+Ym[sample.int(2000, 1000)] <- NA
+
+# Remove completely missing rows -- they are useless
+ikeep <- apply(Ym, 1, function(x) any(!is.na(x)))
+Y <- Ym[ikeep, ]
+
 # Save true values for validation
-ymiss_true <- Y[ymiss]
-Y[ymiss] <- NA
+ymiss <- which(is.na(Y))
+ymiss_true <- Y_all[ikeep,][ymiss]
+
 plot(Y_all, pch = 19, col = "grey")
@@
 boot_means <- apply(data_boot, c(1, 2), mean)[ymiss]
-plot(ymiss_true ~ boot_means, pch = 19,
+# 95% CI for imputed values
+boot_lo <- apply(data_boot, c(1, 2), quantile, probs = 0.025)[ymiss]
+boot_hi <- apply(data_boot, c(1, 2), quantile, probs = 0.975)[ymiss]
+plot(
+  x = 0, y = 0,
+  xlim = range(c(boot_lo, boot_hi)),
+  ylim = range(ymiss_true),
+  type = "n",
   xlab = "Imputed values (bootstrapped mean)",
-     ylab = "True missing values")
+  ylab = "True missing values"
+)
+arrows(boot_lo, ymiss_true, boot_hi, ymiss_true, length = 0.01, code = 3,
+       col = "gray70")
+points(boot_means, ymiss_true, pch = 19)
 abline(0, 1, lty = "dashed")
+abline(lm(ymiss_true ~ boot_means), col = "blue")
+legend("bottomright", c("1:1", "Regression"),
+       lty = c("dashed", "solid"),
+       col = c("black", "blue"))
 ```
 
-We can see that for all data points where at least one observation was available as constraint, the algorithm imputed values that were close to the true values.
-However, where values for both traits were missing, the algorithm falls back to the global mean, which has the same statistical properties as the missing data (i.e.
-central tendency and spread) but doesn't match the true values.
+We can see that mean estimates of the imputed values are close to the true values, and that the imputed values have non-trivial (but reasonable) uncertainty around them.
+
+## Demonstration: Hierarchical modeling of the Iris dataset {-}
+
+For a real-world application of this, we can use the Iris dataset.
+Because this data is grouped by species, we can also demonstrate the hierarchical model
+(though with only 3 species, the statistical power for estimating across-species covariance is minimal).
+
+First, an overview of the Iris dataset:
+
+```{r}
+summary(iris)
+pairs(iris[,1:4], col = iris[,5], pch = 19)
+```
+
+Let's also look at the true values of the correlations, which we can compare against the model estimates.
+
+```{r}
+```
+
+First, we isolate the groups from the actual data.
+
+```{r}
+grp <- iris[, 5]
+head(grp)
+dat <- as.matrix(iris[, 1:4])
+head(dat)
+```
+
+As in the previous example, let's introduce 50% random missingness into the data.
+
+```{r}
+dat_m <- dat
+ndata <- prod(dim(dat_m))
+dat_m[sample(ndata, floor(ndata * 0.5))] <- NA
+
+# Again, we only keep rows that have at least 1 value
+ikeep <- apply(dat_m, 1, function(x) any(!is.na(x)))
+Y <- dat_m[ikeep, ]
+grp_m <- grp[ikeep]
+
+# Save true values for validation
+ymiss_true <- dat[ikeep, ]
+ymiss_true[!is.na(Y)] <- NA
+```
+
+Now, fit the hierarchical model.
+
+```{r}
+result <- fit_mvnorm_hier(Y, grp_m)
+```
+
+We can compare our correlation matrix estimates to the true correlations in the data.
+
+```{r}
+result_corr <- lapply(result$stats$Corr_group, "[[", "Mean")
+corr_cols <- colnames(result_corr[[1]])
+iris_split <- split(iris[, corr_cols], iris[, 5])
+true_corr <- lapply(iris_split, cor)
+
+# Take the ratio of our correlation estimate and the true value
+true_corr
+result_corr
+Map("/", result_corr, true_corr)
+```
+
+In all cases, we underestimate the true correlations, but the extent varies by trait and species.
+The largest underestimates (in relative terms) occur for correlations that were the weakest in the observational data, which makes sense since statistical power for correlation estimates scales super-linearly with sample size.
+
+If our imputation algorithm works correctly, then we expect traits with stronger correlations to be imputed more accurately and precisely than traits with weak correlations.
+Let's estimate the missing values to see if this is the case.
+
+```{r}
+boot_data <- bootstrap_missing_hier(result, Y, grp_m)
+boot_means <- apply(boot_data, c(1, 2), mean)
+boot_lo <- apply(boot_data, c(1, 2), quantile, probs = 0.025)
+boot_hi <- apply(boot_data, c(1, 2), quantile, probs = 0.975)
+boot_means[!is.na(Y)] <- NA
+boot_lo[!is.na(Y)] <- NA
+boot_hi[!is.na(Y)] <- NA
+
+par(mfrow = c(2, 2))
+for (i in 1:4) {
+  plot(
+    0, 0,
+    type = "n",
+    xlim = range(c(boot_lo[, i], boot_hi[, i]), na.rm = TRUE),
+    ylim = range(ymiss_true[, i], na.rm = TRUE),
+    xlab = "Estimate",
+    ylab = "True value",
+    main = colnames(iris)[i]
+  )
+  arrows(boot_lo[, i], ymiss_true[, i], boot_hi[, i],
+         length = 0.01, col = "gray70", code = 3)
+  points(boot_means[, i], ymiss_true[, i],
+         pch = 19, col = grp_m)
+  abline(0, 1, lty = "dashed")
+  legend("bottomright", levels(grp_m),
+         pch = 19, col = 1:3)
+}
+```

### 2019-08-26 EAshiklomanovSA2.Rmd
Respond to sampling bias comment

@@
 ```{r}
-boot_data <- bootstrap_missing_hier(result, Y, grp_m)
+boot_data <- bootstrap_missing_hier(result, Y, grp_m, progress = FALSE)
 boot_means <- apply(boot_data, c(1, 2), mean)
@@
 ```
+
+# Method S2: Biased mean estimates in non-representatively sampled bivariate data
+
+One important result of this manuscript is that accounting for trait covariance can reduce biases in trait sampling.
+We can demonstrate this with an idealized case where we know the correct answer.
+Consider two traits "A" and "B", with mean zero, unit variance, and strong covariance ($R = 0.8$):
+
+```{r}
+library(mvtraits)
+mu <- c(0, 0)
+sigma <- matrix(c(1, 0.8, 0.8, 1), nrow = 2, ncol = 2)
+n <- 5000
+sim <- random_mvnorm(n, mu, sigma)
+colnames(sim) <- c("A", "B")
+plot(sim[, "A"], sim[, "B"], xlab = "A", ylab = "B", pch = 19, cex = 0.3)
+```
+
+Now, suppose that trait B is sampled with a strong preference for low values.
+We can simulate this by removing a random sample of B weighted by its value along the normal distribution cumulative distribution function.
+
+```{r}
+sim2 <- sim
+sim2[sample(n, n / 2, prob = pnorm(sim2[, "B"])), "B"] <- NA
+
+colMeans(sim2, na.rm = TRUE)
+
+plot(density(sim[, "B"]), col = "black",
+     main = "Kernel density of trait B",
+     xlab = "Trait B")
+lines(density(sim2[, "B"], na.rm = TRUE), col = "red")
+abline(v = mean(sim[, "B"]), col = "black", lty = "dashed")
+abline(v = mean(sim2[, "B"], na.rm = TRUE), col = "red", lty = "dashed")
+legend("topright", c("True", "Observed"), lty = 1, col = c("black", "red"))
+```
+
+We can see that the univariate estimate of trait B, -0.43, is systematically biased.
+Now, let's fit these data using our multivariate model.
+
+```{r}
+fit <- fit_mvnorm(sim2)
+subset(fit$summary_table,
+       variable == "mu",
+       select = c("variable", "index", "Mean", "2.5%", "97.5%"))
+```
+
+We can see that the mean estimate for B is still negatively biased, but much less so than in the univariate case.

### 2019-09-03 EAshiklomanovSA2.Rmd
Update SI2 with changes from np repo

@@
 library(kableExtra)
-library(drake)
-library(ggplot2)
-library(GGally)
+# library(drake)
+library(dplyr, exclude = "group_rows")
+library(tidyverse)
+## library(GGally, exclude = "nasa")
 stopifnot(
-  requireNamespace("shiklomanov2017np", quietly = TRUE),
-  requireNamespace("dplyr", quietly = TRUE),
-  requireNamespace("tibble", quietly = TRUE)
+  requireNamespace("shiklomanov2017np", quietly = TRUE)
 )
-
-correlation_scatter_cap <- paste(
-  "Relationship between squared pairwise correlation and sample size of pairwise observations for each PFT.",
-  "Grey shading is the 95% confidence interval around a local regression (LOESS).",
-  "Upper triangle is area-normalized values, and lower triangle is mass-normalized values.",
-  "Note that the x-axis uses a $\\log_{10}$ scale."
-)
-
-correlation_fig_cap <- paste(
-  "Fraction of statistically significant ($p < 0.05$) pairwise trait correlation estimates",
-  "as a function of sample size for each plant functional type."
-)
-
-correlation_range_cap <- paste(
-  "Relationship between pairwise correlation strength and data range for each PFT.",
-  "In each pair of sub-panels, the top panel is relationship between pairwise correlation",
-  "and the range of the variable on the x axis",
-  "while the bottom panel is the relationship between pairwise correlation",
-  "and the y variable",
-  "(For instance, in the two-panel figure in row 1 column 2,",
-  "the top panel is the SLA-leaf lifespan correlation as a function of the range of SLA",
-  "and the bottom panel is the same correlation as a function of the range of leaf lifespan).",
-  "Data range is expressed as the difference between the 2.5% and 97.5% quantiles of each normalized trait (i.e. $\\frac{X}{E[X]}$).",
-  "Grey shading is the 95% confidence interval around a local regression (LOESS).",
-  "Upper triangle is area-normalized values, and lower triangle is mass-normalized values."
-)
+npdir <- here::here("..", "np-trait-analysis")
 ```
@@
 prior_df %>%
-  dplyr::mutate(param = factor(param, shiklomanov2017np::both_params)) %>%
+  dplyr::mutate(param = factor(param, shiklomanov2017np::both_params),
+                mean = sprintf("%.3f", mean)) %>%
   dplyr::arrange(param) %>%
@@
   ) %>%
-  kableExtra::kable_styling(latex_options = c("scale_down", "hold_position"))
+  kableExtra::kable_styling(latex_options = c("hold_position"))
 ```
@@
 )
-table_cap_area <- gsub("mass", "area", table_cap_mass)
 
-readd(table_dat) %>%
-  dplyr::select(-dplyr::matches("area")) %>%
+table_cap_area <- paste(
+  "Mean and 95\\% confidence interval of trait estimates",
+  "for area-normalized traits from the hierarchical model.",
+  "Note that some traits for some PFTs have effectively no constraint relative to the mostly uninformative prior (as evidenced by their very large confidence intervals),",
+  "so their values should be used with caution.",
+  "See mass-normalized table for SLA and leaf lifespan estimates."
+)
+
+with_clm <- readRDS(file.path(npdir, "extdata", "tidy-means.rds"))
+
+# Mass-normalized table
+with_clm %>%
+  filter(mass_area == "mass",
+         model_type == "hierarchical") %>%
+  select(pft, param, mu_hi, mu_lo, mu_mean) %>%
+  # Fix units
+  mutate_at(
+    vars(starts_with("mu")),
+    ~if_else(param %in% c("Nmass", "Pmass"), . * 1000, .)
+  ) %>%
+  mutate(string = sprintf("%.2f (%.2f, %.2f)",
+                          mu_mean, mu_lo, mu_hi)) %>%
+  select(pft, param, string) %>%
+  spread(param, string) %>%
+  arrange(pft) %>%
   kable(
@@
       "Leaf lifespan\n(months)",
-      "SLA\n(kg m$^{-2}$)",
+      "SLA\n(m$^2$ kg$^{-1}$)",
       "$N_\\textrm{mass}$\n(mg g$^{-1}$)",
@@
 
-readd(table_dat) %>%
-  dplyr::select(-dplyr::matches("mass")) %>%
+with_clm %>%
+  filter(mass_area == "area",
+         model_type == "hierarchical") %>%
+  select(pft, param, mu_hi, mu_lo, mu_mean) %>%
+  mutate(string = sprintf("%.2f (%.2f, %.2f)",
+                          mu_mean, mu_lo, mu_hi)) %>%
+  select(pft, param, string) %>%
+  spread(param, string) %>%
+  arrange(pft) %>%
   kable(
@@
       "PFT",
-      "Leaf lifespan\n(months)",
-      "SLA\n(kg m$^{-2}$)",
+      ## "Leaf lifespan\n(months)",
+      ## "SLA\n(m$^2$ kg$^{-1}$)",
       "$N_\\textrm{area}$\n(g m$^{-2}$)",
@@
 
-```{r sample_size_table, echo = FALSE, results = "asis"}
-ssw <- readd(sample_size_wide)
+```{r pairwise_correlation_table, echo = FALSE, results = "asis"}
+rma_display <- readRDS(file.path(npdir, "extdata", "rma-display.rds"))
 kable(
-  ssw %>%
-    dplyr::select(pft, leaf_lifespan, SLA,
-                  Nmass, Narea, Pmass, Parea, Rdmass, Rdarea,
-                  Vcmax_mass, Vcmax_area, Jmax_mass, Jmax_area),
-  col.names = c(
-    "PFT", "Leaf lifespan", "SLA",
-    "$N_\\textrm{mass}$", "$N_\\textrm{area}$",
-    "$P_\\textrm{mass}$", "$P_\\textrm{area}$",
-    "$R_\\textrm{d, mass}$", "$R_\\textrm{d, area}$",
-    "$V_\\textrm{c, max, mass}$", "$V_\\textrm{c, max, area}$",
-    "$J_\\textrm{max, mass}$", "$J_\\textrm{max, area}$"
-  ),
-  format = "latex",
-  booktabs = TRUE,
-  escape = FALSE,
-  caption = "The number of non-missing (outside parentheses) and imputed (inside parentheses) observations of each trait for each plant functional type used in this analysis."
-) %>%
-  kable_styling(latex_options = c("scale_down", "hold_position")) %>%
-  group_rows("Broadleaf", 2, 6) %>%
-  group_rows("Needleleaf", 7, 9) %>%
-  group_rows("Shrub", 10, 12) %>%
-  group_rows("Grass", 13, 15) %>%
-  landscape()
-```
+  rma_display,
+  "latex",
+  booktabs = TRUE, longtable = TRUE,
+  caption = paste("Pairwise reduced major axis (RMA) slope values",
+                  "(Mean and 95pct credible interval)",
+                  "for each plant functional type.",
+                  "Values with asterisks indicate slopes",
+                  "with credible intervals that don't intersect zero.")
+)
 
-```{r pairwise_correlation_table, echo = FALSE, results = "asis"}
-corr_results <- readd(corr_processed)
-kable(
-  corr_results %>%
-    dplyr::select(pft, yvar, xvar, corr_value, present, missing),
-  "latex",
-  booktabs = TRUE, longtable = TRUE, escape = TRUE,
-  col.names = c("PFT", "Trait 1", "Trait 2", "Correlation (95% CI)", "present", "missing"),
-  caption = "Pairwise trait correlation values for each plant functional type. Values with asterisks indicate correlations significantly different from zero ($p < 0.05$)."
-)
-```
-
-```{r correlation_sample_size_pairs, echo = FALSE, results = "hide", warning = FALSE, fig.cap = correlation_scatter_cap, fig.dim = c(12, 12)}
-readd(corr_ss_plot_gg)
-```
-
-```{r correlation_range_analysis, echo = FALSE, results = "hide", warning = FALSE, fig.cap = correlation_range_cap, fig.dim = c(12, 12)}
-readd(corr_range_plot_gg)
+## corr_results <- readd(corr_processed)
+## kable(
+##   corr_results %>%
+##     dplyr::select(pft, yvar, xvar, corr_value, present, missing),
+##   "latex",
+##   booktabs = TRUE, longtable = TRUE, escape = TRUE,
+##   col.names = c("PFT", "Trait 1", "Trait 2", "Correlation (95% CI)", "present", "missing"),
+##   caption = "Pairwise trait correlation values for each plant functional type. Values with asterisks indicate correlations significantly different from zero ($p < 0.05$)."
+## )
 ```

### 2019-09-03 EAshiklomanovSA2.Rmd
Add Figure S1 (comparison of all means)

@@
 npdir <- here::here("..", "np-trait-analysis")
+```
+
+```{r mean-compare-all, echo = FALSE, fig.cap = mca_cap}
+mca_cap <- paste(
+  "Mean and 95% credible intervals on estimates of trait means for each plant functional type",
+  "from the univariate, multivariate, and hierarchical models.",
+  "For leaf lifespan and SLA, results were similar for mass- and area-based models,",
+  "so only results from the mass-based model area shown.",
+  "For some PFT-trait combination, where error bars are large due to the relatively",
+  "uniformative prior, $y$ axes are constrained to facilitate comparison."
+)
+knitr::include_graphics(file.path(npdir, "figures", "manuscript", "mean_comparison_all.pdf"))
 ```

### 2019-09-03 EAshiklomanovSA2.Rmd
Fix SA2 formatting

--- 2019-09-03 EAshiklomanovSA2.Rmd
@@
 ```
+
+(Page intentionally left blank)
+
+\clearpage
+
+```{r mean-compare-all, echo = FALSE, fig.cap = mca_cap, fig.position = "H"}
+mca_cap <- paste(
+  "Mean and 95% credible intervals on estimates of trait means for each plant functional type",
+  "from the univariate, multivariate, and hierarchical models.",
+  "For leaf lifespan and SLA, results were similar for mass- and area-based models,",
+  "so only results from the mass-based model area shown.",
+  "For some PFT-trait combination, where error bars are large due to the relatively",
+  "uniformative prior, $y$ axes are constrained to facilitate comparison."
+)
+knitr::include_graphics(file.path(npdir, "figures", "manuscript", "mean_comparison_all.pdf"))
+```
+
+\clearpage
 
@@
 
+\clearpage
+
 ```{r pairwise_correlation_table, echo = FALSE, results = "asis"}
@@
                   "with credible intervals that don't intersect zero.")
-)
+) %>%
+  kableExtra::kable_styling(latex_options = "hold_position")
 
@@
 
+# TODO: Fix these figure margins
 par(mfrow = c(2, 2))

### 2019-09-04 EAshiklomanovSA2.Rmd
Add LL vs. climate supporting figure

@@
 knitr::include_graphics(file.path(npdir, "figures", "manuscript", "mean_comparison_all.pdf"))
+```
+
+\clearpage
+
+```{r ll-clim, echo = FALSE, fig.cap = ll_clim_cap, warning = FALSE}
+ll_clim_cap <- paste(
+  "Relationships of leaf lifespan to mean annual temperature ",
+  "within a subset of PFTs,",
+  "with least-squares linear regression and 95% confidence interval."
+)
+traits_clim <- file.path(npdir, "extdata", "traits_with_climate.rds") %>%
+  readRDS() %>%
+  as_tibble()
+clim_dat <- traits_clim %>%
+  distinct(ObservationID, Latitude, Longitude, AMT)
+trait_analysis <- file.path(npdir, "extdata", "traits_analysis.rds") %>%
+  readRDS() %>%
+  as_tibble()
+dat <- inner_join(trait_analysis, clim_dat, "ObservationID") %>%
+  mutate(pft = forcats::fct_recode(clm45, !!!shiklomanov2017np::abbr2pft))
+ppp <- c("BlDTe", "BlDBo", "ShDTe", "ShDBo", "NlETe", "NlEBo")
+
+dat %>%
+  filter(pft %in% ppp, !is.na(leaf_lifespan), !is.na(AMT)) %>%
+  mutate(pft = factor(pft, ppp)) %>%
+  ggplot() +
+  aes(x = AMT, y = leaf_lifespan) +
+  geom_point() +
+  geom_smooth(method = "lm") +
+  facet_wrap(vars(pft), scales = "free_y", dir = "v", ncol = 3) +
+  labs(x = expression("Annual mean temperature" ~ (degree * C)),
+       y = "Leaf lifespan (months)") +
+  cowplot::theme_cowplot()
+
+## dat %>%
+##   filter(pft %in% ppp, !is.na(SLA), !is.na(AMT)) %>%
+##   mutate(pft = factor(pft, ppp)) %>%
+##   ggplot() +
+##   aes(x = AMT, y = SLA) +
+##   geom_point() +
+##   geom_smooth(method = "lm") +
+##   facet_wrap(vars(pft), scales = "free_y", dir = "v", ncol = 3) +
+##   labs(x = expression("Annual mean temperature" ~ (degree * C)),
+##        y = expression("Specific leaf area" ~ (kg ~ m^{-2}))) +
+##   cowplot::theme_cowplot()
+
 ```

### 2019-09-05 track-changes.sh
Add track-changes script

@@ -1,0 +1,5 @@
+#!/usr/bin/env bash
+
+rm -rf shiklomanov_ea_diff.{aux,log,out,tex,pdf}
+latexdiff shiklomanov_ea_old.tex shiklomanov_ea.tex > shiklomanov_ea_diff.tex
+pdflatex shiklomanov_ea_diff.tex

### 2019-09-06 ecoapp-comments/response.txt
Add plain-text version of response letter

Dear Ecological Applications editorial staff and reviewers,

We appreciate your thorough review of our manuscript, "Does the leaf economic spectrum hold within plant functional types? A Bayesian multivariate trait meta-analysis." We are pleased to present a significantly revised version of the manuscript, which we feel has improved significantly since the last version, in large part thanks to the reviewers' thoughtful and helpful comments and suggestions. A summary of the major changes in this revision is as follows:

First, and most importantly, we have re-written virtually all of the results and discussion concerning the scale dependence of leaf economic relationships. We now organize both of these sections around three groups of traits that show varying degrees of consistency with the global leaf economic spectrum: Morphological and biochemical traits (SLA, N, and P), for which these relationships are most robust; metabolic traits (R_d, V_{c,max}, J_{max}), for which they are somewhat more variable; and leaf lifespan, for which they are even more variable. We have also replaced the 2 figures related to this topic---the previous Figures 2 and 3---with a new Figure 1 that should be easier to read and which better shows the major features of our results.

Second, we have re-organized the introduction to better set up the need and ecological context for our analysis. We now begin with a discussion of trait covariance in general and of the global leaf economic spectrum specifically. We then provide a brief overview of evidence for deviations from the leaf economic spectrum at smaller scales. We then briefly discuss the challenges with working with individual species from a modeling perspective, which leads to a paragraph introducing plant functional types. This leads us to our guiding question about whether the leaf economic spectrum holds at the PFT scale.

Finally, following the request of Reviewer 1, we have performed a cross-validation of our approach to trait imputation. This analysis provides a more robust comparison of the three models we tried, and serves as a validation of our imputation approach, which is important given the massive volume of imputation performed in this analysis.

In addition to these changes, we have addressed many concerns and suggestions of the reviewers about specific lines of text, aesthetic choices, etc.

Thank you in advance for your consideration of this revised manuscript!

(NOTE: All line numbers below refer to the document with changes highlighted.)

1 Reviewer 1
============

        a) at several crucial places in the ms, the authors interpret a reduction of posterior uncertainty essentially as an improvement of model fit. Although I would agree that this pattern is indicative of a better fit, it is not a direct proof, and it is possible to have counterexamples. To resolve this issue, the authors should explicitly assess model fit (e.g. calculate RMSE or similar either on the fitted data, or a hold-out), and replace this in the discussion appropriately.

We have added a 20-fold cross-validation exercise comparing the ability of the grouped univariate, multivariate, and hierarchical models to impute trait values. This analysis is described on lines 289-292, and the results are shown in the new Fig. 3, described on lines 410-417, and briefly discussed in lines 554-555. We also revised phrasing throughout the manuscript to the more precise "reduced predictive uncertainty".

        b) The paper itself is written comparatively well, but motivation and ecological interpretation could be extended at some points. See my detailed remarks.

  We have re-written large parts of the introduction to address this concern, with particular emphasis on the need for PFTs in the context of trait ecology and vegetation modeling. We have also heavily revised the results (section 3.1 throughout) and discussion (section 4.1 throughout) concerning within- vs. across-PFT covariance patterns to better highlight overarching ecological patterns. Specifically, we divided the seven traits in this analysis into three groups with varying degrees of consistency for leaf economic relationships: Morphological and biochemical traits (SLA, N, and P) show the most robust relationships; metabolic traits (R_d, V_{c,max}, J_max) show slightly less robust relationships; and leaf lifespan shows the least consistent relationships.

        c) The choice to structure species by PFTs seems obvious from the perspective of a person working with dynamic vegetation models, but less obvious from a general ecology / data analysis perspective. If we forget the models for a moment, I am quite sure that there are other grouping options that might make sense, maybe more sense, other than a PFT grouping. Some things that come to mind is

        - Group species according to taxonomy
        - Group species according to traits (have to make sure
          that this doesn’t become circular)
        - Group species according to space / habitat / ecosystem
        - Fit species grouping with the model (could also consider
          unsupervised classification according to some objective)

        I think the authors should at least consider and discuss these options in the paper as a possibility for further research. Personally, I think adding some of these analyses would make a much stronger and more general paper, but I hesitate to make this a requirement, as this would shift the focus of the paper and would probably require significant rewriting of intro and discussion as well.

  We agree that exploring other species groupings, including those suggested here, would be an interesting future direction for this work, and we have added text to this effect to the discussion (lines 584-587). However, we have chosen to keep this paper focused on groupings by plant functional type, primarily because these rules are relevant, and can be used as inputs, to dynamic vegetation models.

        2 Abstract goes directly into the topic. A bit of motivation / relevance would be nice.

  We have added two sentences introducing the leaf economic spectrum and its study at different scales to the abstract (lines 2-5).

        8 indicating … -> Spontaneously, I don’t see how this can be concluded, but I’ll read on and wait to be educated

  We have removed this sentence from the revised introduction, so this comment no longer applies.

        13 I was again missing a conclusion about the relevance / applicability of the findings

  We have revised the abstract accordingly to better highlight the ecological implications of our results (lines 13-19, 21-24, 27-29).

        15 Personally, I would have found it more logical to start this paper with trait correlations and the LES, because this is what you examine in the end, and then follow up with the PFTs

  We have re-organized the introduction in accordance with this comment. We now start with a discussion of the leaf economic spectrum, then move to empirical evidence of and mechanisms for local-scale deviations from it, and eventually arrive at why PFTs are a useful intermediate scale for this study.

        17 Why "however". There doesn’t seem to be a contradiction. It’s more: at the same time

        19 Benefits … most papers cited here just demonstrate the fact that there is variability and speculate about the benefits, but OK …

  We have rewritten this part of the introduction, so this comment no longer applies.

        23 could make a paragraph break here

  In the original draft, we agree that this would have made sense. However, in the revised version, this entire paragraph is shorter, and this sentence is a natural continuation of the previous point, so we have not added a break.

        24 It’s a bit weird to say the allometry constrains the
        traits, as the allometry can be said to be a trait itself.
        I guess what you mean is that there are relationships
        between traits, so that trait-space can be reduced.

  We have replaced "allometries" with "biophysical limitations", and clarified the first half of this sentence to refer specifically to "variability in plant functional traits" (lines 42-43).

        29 erase well

  We have revised this accordingly (line 50).

        31 causality probably goes both ways – in particular climates, you find particular parts of the LES, although there may also be plasticity

  We agree, and have replaced "influenced by" with "related to" (line 54).

        35 could also cite Scheiter, Simon, Liam Langan, and Steven I. Higgins. "Next‐generation dynamic global vegetation models: learning from community ecology." New Phytologist 198, no. 3 (2013): 957-969.

  We have added this reference (line 57).

        51 Here and for the other points above: I agree with the conclusions summarized here, but it might be helpful for the reader to provide a very short sketch of the premises / arguments for the respective conclusions in the cited papers. E.g. what’s the argument in Messier 2016 for the causality problem?

  In our re-organization of the introduction, we ended up dropping the specific sentence to which this is referring.

        51 Also, the Messier paper is 2017, not 2016

  We have revised this reference everywhere it occurs.

        53 OK, but why is that a problem for using global trait correlations?

  In our re-organization of the introduction, we ended up dropping the specific sentence to which this is referring (line 84).

        59 I get the assembly point, this is basically about transient dynamics through dispersal limitation, but isn’t the larger concern if the potential trait space / trade-off is actually much larger / different than what we see as being “filtered by evolution”.

  In our re-organization of the introduction, we ended up dropping the specific sentence to which this is referring (lines 88-91).

        60 An issue with this entire section that ends here is that it doesn’t become clear which of the issues you discuss here (if any) can be solved by an approach where you fit trait correlations in hierarchical groups. The arguments in favor of the PFT analysis in the following paragraph make sense to me, but they relate only vaguely to the specific issues in the previous section. Please be more clear about which issues are solved by working with PFTs (and why), and which remain unsolved.

  We have re-organized and re-written the introduction to address this and related concerns. In the revised version, there is a more logical argument flow: (1) Global trait correlations may not hold at local scales (lines 60-72). (2) But, trying to do ecosystem ecology from the bottom-up starting with individual species is also problematic (lines 96-104). (3) Therefore, PFTs are a useful intermediate scale (lines 105-125).

        78 What do you mean by constrain? Predict? See also my major comment: what’s the performance criterion under which you want to evaluate an “improvement”.

  We have clarified this by replacing "constrain" with "reduce uncertainties in" (line 130). We address the broader concern about model fit in the main response.

        109 Seems obvious that one could also try this analysis on taxonomic groups

  We agree that such an analysis would be worthwhile, but it is outside the scope of this paper. See main response.

        127 of course, both the univariate and the multivariate models could be made hierarchical. Not 100% clear why you only choose to make the multivariate model hierarchical.

  We have clarified here that the "hierarchical" model means "hierarchical multivariate" (line 180). More generally, our hypothesis is that the benefit of hierarchy to trait estimation comes from leveraging the across-PFT covariance structure, which is nonexistent in the univariate model.

        134 Could you comment on whether Gibbs sampling is the most efficient / preferred method for the estimation of covariance matrices in such a setting?

  We have added a sentence about the advantages of Gibbs sampling in terms of its 100% proposal acceptance rate (compared to the 10-65% target acceptance rates of Metropolis Hastings, Differential Evolution, and Hamiltonian Monte-Carlo) (line 191-195). We recognize that this is not a complete picture of the efficiency, since other algorithms could potentially compensate for lower acceptance by achieving faster convergence. A thorough comparison of algorithm efficiency would be useful and interesting, but is outside the scope of this work.

        136 You should specify the parameters, i.e. were all priors uninformative?

        137 Same for the Wishart – was this a more or less uninformative Wishart?

  We added a table containing values for the weakly-informative priors on the multivariate normal distribution to the supplementary information (Tab. S1), and a description of the other priors to this section (lines 210-213).

        138 Estimating covariances is notoriously problematic, in particular in small data situations. It would seem useful to me to spend a few words on the degrees of freedom, scaling, and whether you see an overfitting concern here, plus about possible ideas for Bayesian regularization of the means or covariances (in case of low data situations). If you think that the data is usually so strong that this is not an issue, that’s also fine, but would be good to be explicit about this. Also, I wonder: have you tested (via simulation) that your current setup can faithfully retrieve the correlation parameters?

  We have added a few sentences to the methods describing how our approach avoids overfitting by tending towards zero covariance estimates with wide credible intervals in cases where data are limiting (lines 241-246). Beyond that, given Reviewer 2's comments that this manuscript is "too focused on the technical methods at the expense of the motivating questions", we are reluctant to spend too much time discussing alternative approaches, including Bayesian regularization.

        139 Not sure if this equation is useful to the reader. Is there anything we learn from that? Seems just a restatement of what you said

  We have removed this equation as well as the preceding one (which was similarly uninformative).

        139 Equations should have numbers

  We have added numbers to all equations.

        141 Which page? This is a large book.

  We have added the section number, section title, and page number (line 207-208).

        142 I don’t understand why this would be the case. Can you explain why you couldn’t model missing observations as latent variables?

  This was unclear wording on our part. What we meant to say here was that the most popular R packages for Bayesian analysis using graphical models (namely, BUGS/JAGS and Stan) do not have convenient support for multivariate distributions with partial missingness. Our model actually does treat the partially missing observations as latent variables -- it just samples them in blocks conditioned on the partially present data in each row and the current sampled mean vector and variance-covariance matrix. We have revised this section to make this clearer (lines 214-237).

        147 The explanation is not 100% clear to me: do I understand correctly: in a loop, you a) impute b) fit the hierarchical model? What I don’t get is: i) how does the fitted model then feed back to the imputation step? Also, when do you stop? I think it would be important to explain this procedure in more detail.

  When we (more correctly) describe our approach as modeling missing values as latent variables (see previous comment), this section should make more sense. However, for clarity, we have added a more detailed step-by-step explanation of the sampling procedure here (lines 227-237). We also refer readers to the detailed (and now, revised and extended) demonstration of the algorithm in Supporting Information Method S1.

        147 Moreover, if this a new method, I would ask you to confirm that this actually works with randomly removed data (it’s clear that imputation will always run in trouble if you remove data purposefully, but at least let’s make sure it works under ideal conditions). Ideal validation would be via <https://arxiv.org/pdf/1804.06788.pdf> , in this case you check the entire analysis chain.

  We have added a section to Supplementary Method S1 that provides a detailed validation of our approach for imputing missing values on simulated bivariate data and the classic Iris dataset.

        150 How many missing values are we talking about here in the real data? 10%, 90%? See comment above, if you check the methods, make sure that the scenario corresponds to what you find in the real data

  We have added the sample sizes for each trait and PFT to Table 1. Information on pairwise missingness and the corresponding estimates of the reduced major axis slope coefficients are shown in Table S4.

        153 Is this the univariate psrf, i.e. per parameter? If so, write: for all parameters

  We have revised this accordingly, clarifying that this is the univariate PSRF for every parameter (line 248).

        160 I realize that it’s inherently difficult to summarize a 7-dim correlation, but just as a comment: would it be possible that, if the first and second largest EV are very close to each other, one could get a large switch of the direction of the dominant EV by a small change in the data, similar to what often happens in PCAs? As a more robust alternative, one could consider similarity measures of the MVN, e.g. KL distance, which would be independent of these axis problems.

  We have replaced this analysis and corresponding figure with one showing the reduced major axis slopes (Fig. 1). This calculation (including the uncertainty propagation) is conceptually much simpler, so we feel this comment no longer applies.

        Fig.3 What you provide here is basically the raw data for the covariance matrix. It’s really hard to see any trends here. Consider if you can somehow compress this to something more meaningful. Also, axis should be properly labeled.

  As mentioned above, we have replaced this figure with a tile plot of reduced major axis slopes (Fig. 1) that, we hope, more clearly shows patterns within- and across PFTs.

        Major comment analysis: all measures you report here are based on the fitted models only (i.e. don’t consider the fit to the data / residuals). I find that somewhat dangerous. What I would ask you to add is some measure if predictive error for all three models (univariate, multivariate, hierarchical multivariate). I think in this case it’s unlikely that you’re grossly overfitting, so it would be OK for me if you would calculate this on the same data that you used for fitting, but more ideal would of course be to have a hold-out. Consider also if there is a possibility to analyze the error, e.g. by plotting / regressing it against suitable predictors such as PFT, taxonomy, location

  See main response.

        200 Here and later: it’s unclear what model you refer to when you say “global” etc. … you introduced three models: univariate, multivariate, and multivariate hierarchical. I assume that you can’t mean the univariate, but with global you could mean the global multivariate, the between PFT correlation in the hierarchical, or the global (within + across PFT) correlation in the hierarchical model.

        226 Maybe I’m missing it, but is there any discussion of the correlation between PFTs? In general, I think you could leverage the hierarchical setup much better. Essentially, what we have is a global correlation. With the hierarchical model, we can now partition this global LES into a within and between PFT correlation. What I understand from this section so far is that the within PFT correlation is broadly identical to the global correlation, except for some PFTs. But what about the between PFT correlation, does this also follow the LES, or is this completely different? Apart from the verbal discussion, it would be great if the partitioning could be visualized in some way, so that one could see where the different PFTs lie in the overall trait spectrum, and how the traits correlate within them. If this could be done, I think it could be an ecologically very valuable figure.

        234 OK, I see this addresses in some sense my point about the correlation across PFTs above, but would be nice to discuss this in terms of correlation of the means, not means only.

  We have completely re-written this section of the results around the new Figure 1 and with a heavier and explicit focus on the across-PFT patterns (section 3.1, throughout).

        233 and which had lower error?

  In the revised results section, this comment no longer applies.

        As a side note: given the small number of PFTs, one does wonder if the trait correlation across PFTs can even properly be estimated (I think you make some comments to this effect later, but if it can’t estimate, then why fit this model?).

  We have added a simple power analysis of correlation coefficients (lines 376-380). The result is that, with 14 PFTs, we can estimate correlations of at least 0.74 with 95% power and confidence. This means that our "non-significant" correlations still allow us to place an upper limit on the strength of across-PFT correlations, albeit a higher one than we would have liked.

        244 significantly? This is a Bayesian analysis.

        268 You use again “significant”, but I don’t think you have run significant tests here. Clarify what you mean

  We have replaced "siginficantly" with more precise text about intersection with the 95% credible interval throughout the text.

        244 how were CLM parameters derived?

  This is described in the "Methods: Analysis of results section" (lines 285-288).

        251 OK, this is the uncertainty of the mean, but more important for me would be the predictive error, which could be very different (also depending on how much IV is in these traits).

  See main response.

        279 It’s a bit unfortunate that this analysis doesn’t allow separating spread and direction of the trade-off. What I mean is that a lower correlation could result from a smaller slope in a regression between two traits, or from more spread around the regression line. Would be nice to see which of the two possibilities is responsible here, a change of the direction of the trade-off, or an increase in variability. Please discuss if applicable.

        282 Again, for building this into models, one would probably be interested in the slope and not R2

        296 Again, I wonder if “weak” means that slope or R2, plus, again: significance. Moreover, note that significance is not effect size.

  As mentioned above, we have replaced the analysis of correlations with analysis of reduced major axis slopes, which make this distinction possible.

        316 The logical conclusions seems to examine other groups than the standard model PFTs, e.g. divide species taxonomically, morphologically, or via biomes

  We have added a sentence about this as a future direction later in the discussion (line 584-587).

        331 This is a weird wording, at least you should use predictive uncertainty, but I’m not even sure if you showed this. What you seem to discuss here is a reduction of posterior predictive uncertainty, but that is just an estimate. I can reduce posterior predictive uncertainty just by adjusting the prior, but that does’t mean that the model will better fit the data. I think you urgently need to establish a measure of model fit / predictive error for this discussion.

        335 but again, a change doesn’t imply an improvement. You have to show that the multivariate model is better

  First of all, we have revised this wording to be more precise ("posterior predictive uncertainty", line 555). Second, while it is technically true that the posterior predictive uncertainty can be reduced by tweaking the prior, that is not the case in our analysis because all three models used the same priors (on corresponding terms; e.g. variance priors in univariate model were the same as diagonals of covariance matrix priors in multivariate models). Finally, and most importantly, we have added the RMSE analysis requested here -- see main response.

        337 I don’t see how you arrive at this conclusion. If the data-generating process was multivariate normal, univariate and multivariate means should be identical. To me, this result rather suggests that the correlation is not multivariate, which seems to question the assumptions of your model, or that the PFT means are not in line with the global mean (so that you get a mixture distribution, but this has nothing to do with multivariate). It may be a good idea to provide at least a few visual checks of model adequacy, e.g. by posterior predictive model checks

        339 OK, I think if the traits are correlated, the multivariate model is clearly better, because we want to know this correlation. I find this so obvious that it wouldn’t even need discussing. However, with the comments above in mind, I think you should specify what you mean by “bias” – as said, I don’t think there should be a bias in the mean if the data-generating process is multivariate normal, and you fit univariate normal distributions.

  The univariate and multivariate means are only identical if partially missing data are dropped, in which case additional information about missing parameters based on covariance with existing parameters is neglected. However, by leveraging among-trait covariance, we can make more educated guesses about parts of the trait space that have data for other traits, and we can therefore at least partially overcome sampling biases. We have added a demonstration of this effect using simulated data as Method S2 in the Supplementary Information, and reference it here (line 563).

        341 What do you mean by “reliable”?

  We changed this to say "too few observations to estimate covariance patterns for some PFTs with much statistical power" (line 567).

        349 That makes sense to me, but I was missing the relating info in the results

  See our above responses (comment "As a side note...", between "233..." and "244...") about the power analysis and pairwise sample size tables in the Supplementary Information.

        355 It also seems obvious that the procedure could make use of phylogenetic or morphological similarity, as it is commonly done in PGLS type models

  We have mentioned this as an additional future direction of this work (line 584-585).

        356 It seems what you discuss here is independent of PFTs, right? But then I would say people are already working on this, and it doesn’t really relate to the results of this study

  We respectfully disagree with this comment for several reasons. A key result of this work is the additional constraint on trait estimates provided through multivariate approaches, whether hierarchical or not, and we expect this constraint to be particularly valuable for the hard-to-measure traits discussed in this section. Moreover, if, as we show (more clearly in this revision), even well-established leaf economic relationships have some hierarchical character, then examining the extent to which this is the case for other "economic spectra" (namely, wood and roots, as mentioned here) is a very relevant future application.

        368 I think the mutually confounding issue is actually the point that is most interesting from an ecological perspective. The relevant question is if trade-offs such as the LES scale through all levels (individual, PFTs, all plants), or if they could be different at some scales. When taking together your results and other literature that looks at LES within species, it seems to me that current evidence suggestions that trade-offs tend to be universal? Could you extend on this discussion, possibly also highlighting which trade-offs you suspect to be universal, and which might be different at different ecological scales?

  As mentioned in the main comments, our revised version of the manuscript focuses much more on the differing extent to which some trait relationships are universal or actually scale-dependent.

        372 I wouldn’t sign this statement, but OK, if you think so

        385 See my concern: uncertainty != predictive error

  We have removed this section, so these comments are no longer relevant.

        Fig. 1 not a big fan of this lattice plot layouts, you are losing a lot of space with the plot headings, and the PFTs would better be placed below the plots, under the bars, imo. Consider if the visualization can be improved.

        Fig. 2 as said, not sure what you mean by significant. Moreover, why would you not plot the n.s. too? A correlation of zero is also an information, isn’t it?

        Fig. 3 as said in the previous comments, I wonder if this could be better condensed. Difficult to see any patterns because of the high amount of info.

  We have replaced all three of these figures in the revised version.

        Fig. 4 Same point about lattice layout. Also, not sure why you suddenly introduce the units to the traits, while they are absent in all other plots.

  The units in this figure make more sense now, since it is the first figure to actually show trait values (rather than their slopes). We have decided to keep the "lattice" layout because it is the most effective way we have to show results for all 8 traits. However, we have moved the results from the univariate and multivariate models to the supplementary information to reduce visual clutter and improve readability.

        Fig. 5 see my main comment: CI is not fit.

  See main response.

2 Reviewer 2
============

        This manuscript starts from the relatively simple question of whether leaf economic spectrum (LES) patterns (correlations between SLA and leaf N, etc) hold when they are applied to a subset of the global distribution of plants (divided into PFTs). This is an interesting question because we use the LES a lot in trait-driven land surface models, yet these relationships tend to break down using smaller subsets (by area, etc) of plants. Reaching beyond the traits traditionally considered in the LES, this paper also considers metabolic 'traits'.

        While I think this is an interesting question, I found myself wondering at the end of this paper what the actual point was. It reads a bit like a bunch of exploratory analyses in search of a conclusion, as evidenced by the many nearly unreadable multi-panel and matrix plots. Given that we know quite a bit about the LES, it would have been interesting to start this manuscript with some hypotheses instead of an exploration -> looking at all of these plots it is difficult to untangle what is unexpected versus what is normal. Similarly, why did the authors think the hierarchical model was necessary, and what does it mean that it didn't improve estimates? Overall this manuscript feels a bit too focused on the technical methods at the expense of the motivating questions.

  We appreciate this point and agree that the introduction did not do a good enough job of laying out the motivation for this study, and that its driving hypotheses were implicit, not explicit, making things more difficult for the reader. We have revised the introduction to be clearer and more logically organized, with an explicit motivating question at its end. (See main response.)

        Writing-wise, this manuscript is well written and clear, however, it feels like it is written to target people already in the small niche of ecological land surface modelers. I think it would be more appealing to a wider audience if the introduction was edited to explain more of the underlying concepts and why we care about them. For example, the manuscript refers to LES working at the global scale and not at local scales, then refers to PFTs as an intermediate scale (lines 74-76) but this is using "scale" in a very abstract way that deserves more explanation. Global and local scales are both spatial, but PFTs are really a conceptual scale between lumping everything together ('global') and modeling each species or individual separately.

  We have re-written large parts of the introduction to address this concern, with particular emphasis on the need for PFTs in the context of trait ecology and vegetation modeling.

        At the very end of the manuscript PFTs are then semi-equated with biomes (line 383-384) which isn't right.

  We have removed this paragraph.

        Throughout the introduction it would help if the manuscript spent a bit more time explaining to a novice/outside reader why PFTs, traits, and land surface models are all worth reading and learning about.

  See earlier comment about revisions to the introduction.

        Table 1: Make sure it's clear that the top 8 PFTs in the list are trees.

  We have merged this table with the sample size table (see next comment), and have replaced this accordingly.

        Figure 1: Seems like this would be better as a table in an appendix? Dedicating a whole figure to show sample sizes seems like a lot.

  We have replaced this figure with a table.

        Figure 2: This figure needs more explanation in the caption of what is being shown. What does the slope and length of the lines represent? Also, it's almost impossible to tell this many colors apart in a figure (colorbrewer doesn't even have 14 classes as an option) - it might be worth trying to come up with a pared down version of this that focuses on some key points.

        Figure 3: Is this showing something substantially different from figure 2? This one is even harder to interpret just given the sizes of the plots and the numbers of bars.

  We have replaced both of these figures with a single, very different figure (Fig. 1) that, we hope, does a better job of highlighting the key results.

        Figure 4: Again, this is a lot of information making it hard to see what is important and what isn't, especially given that in most cases the three model types produce almost identical results. Also, the caption states that leaf lifespan and SLA only show mass based results, but is there a mass vs. area difference for leaf lifespan? and SLA is per mass by definition, no? Finally on this one, it looks like CLM values have confidence interval bars on them, but they are just single values, I think?

  We have moved the original version of this figure to the supplementary information and replaced it with a version that only has the results from the hierarchical model and the CLM 4.5 parameters. We have also changed the symbol for CLM 4.5 values from "+" to "x" to make it clearer that those values do not have error bars. Finally, we have clarified in the caption that the normalization refers to traits /other/ than SLA or leaf lifespan.

        Tables S1 and S2: SLA is listed as kg m-2 in both of these - should be m2 kg-1 (as in text, line 90), and/or flipped to LMA for area-normalized?

  Thanks; yes, these units should have been m2 kg-1. We have fixed this in the revision.

        Table S4: What do 'present' and 'missing' mean here? or how are there correlation values when 'present' = 0?

  We have replaced this table, so this comment no longer applies.

### 2019-09-09 AppendixS1.tex
Fix Appendix 1. Refactor authors file.

% Created 2019-09-09 Mon 08:53
% Intended LaTeX compiler: pdflatex
\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{grffile}
\usepackage{longtable}
\usepackage{wrapfig}
\usepackage{rotating}
\usepackage[normalem]{ulem}
\usepackage{amsmath}
\usepackage{textcomp}
\usepackage{amssymb}
\usepackage{capt-of}
\usepackage{hyperref}
\usepackage[left=1in,right=1in,top=1in,bottom=1in]{geometry}
\usepackage[backend=biber,style=authoryear,date=year]{biblatex}
\addbibresource{/Users/shik544/Dropbox/references/library.bib}
\input{authors.tex}
\usepackage{hanging}
\date{}
\title{Does the leaf economic spectrum hold within plant functional types? A Bayesian multivariate trait meta-analysis}
\hypersetup{
 pdfauthor={},
 pdftitle={Does the leaf economic spectrum hold within plant functional types? A Bayesian multivariate trait meta-analysis},
 pdfkeywords={},
 pdfsubject={},
 pdfcreator={Emacs 26.3 (Org mode 9.3)}, 
 pdflang={English}}
\begin{document}

\maketitle
\noindent \textbf{Journal:} Ecological Applications

\clearpage

\section*{Appendix S1: References for the original sources of the TRY trait data used in our analysis.}

\begin{hangparas}{1.5em}{1}\hypertarget{citeproc_bib_item_1}{Adler, P. B., D. G. Milchunas, W. K. Lauenroth, O. E. Sala, and I. C. Burke. 2004. Functional Traits of Graminoids in Semi-Arid Steppes: a Test of Grazing Histories. Journal of Applied Ecology 41:653–663.}

\hypertarget{citeproc_bib_item_2}{Adriaenssens, S. 2012. Dry deposition and canopy exchange for temperate tree species under high nitrogen deposition.}

\hypertarget{citeproc_bib_item_3}{Akhmetzhanova, A. A., N. A. Soudzilovskaia, V. G. Onipchenko, W. K. Cornwell, V. A. Agafonov, I. A. Selivanov, and J. H. C. Cornelissen. 2012. A Rediscovered Treasure: Mycorrhizal Intensity Database for 3000 Vascular Plant Species Across the Former Soviet Union. Ecology 93:689–690.}

\hypertarget{citeproc_bib_item_4}{Araujo, A. de, J. Ometto, A. Dolman, B. Kruijt, M. Waterloo, and J. Ehleringer. 2012. LBA-ECO CD-02 C and N Isotopes in Leaves and Atmospheric CO2, Amazonas, Brazil. ORNL Distributed Active Archive Center.}

\hypertarget{citeproc_bib_item_5}{Atkin, O. K., M. Schortemeyer, N. McFarlane, and J. R. Evans. 1999. The Response of Fast- and Slow-Growing Acacia Species To Elevated Atmospheric CO2: An Analysis of the Underlying Components of Relative Growth Rate. Oecologia 120:544–554.}

\hypertarget{citeproc_bib_item_6}{Auger, S., and B. Shipley. 2012. Inter-Specific and Intra-Specific Trait Variation Along Short Environmental Gradients in an Old-Growth Temperate Forest. Journal of Vegetation Science 24:419–428.}

\hypertarget{citeproc_bib_item_7}{Bakker, C., J. Rodenburg, and P. M. van Bodegom. 2005. Effects of Ca- and Fe-Rich Seepage on P Availability and Plant Performance in Calcareous Dune Soils. Plant and Soil 275:111–122.}

\hypertarget{citeproc_bib_item_8}{Bakker, C., P. M. Van Bodegom, H. J. M. Nelissen, W. H. O. Ernst, and R. Aerts. 2006. Plant Responses To Rising Water Tables and Nutrient Management in Calcareous Dune Slacks. Plant Ecology 185:19–28.}

\hypertarget{citeproc_bib_item_9}{Baraloto, C., C. E. T. Paine, L. Poorter, J. Beauchene, D. Bonal, A.-M. Domenach, B. Hérault, S. Patiño, J.-C. Roggy, and J. Chave. 2010. Decoupled Leaf and Stem Economics in Rain Forest Trees. Ecology Letters 13:1338–1347.}

\hypertarget{citeproc_bib_item_10}{Beckmann, M., M. Hock, H. Bruelheide, and A. Erfmeier. 2012. The Role of UV-B Radiation in the Invasion of Hieracium pilosella-A Comparison of German and New Zealand Plants. Environmental and Experimental Botany 75:173–180.}

\hypertarget{citeproc_bib_item_11}{Blonder, B., V. Buzzard, I. Simova, L. Sloat, B. Boyle, R. Lipson, B. Aguilar-Beaucage, A. Andrade, B. Barber, C. Barnes, and et al. 2012. The Leaf-Area Shrinkage Effect Can Bias Paleoclimate and Ecology Research. American Journal of Botany 99:1756–1763.}

\hypertarget{citeproc_bib_item_12}{Blonder, B., C. Violle, L. P. Bentley, and B. J. Enquist. 2010. Venation Networks and the Origin of the Leaf Economics Spectrum. Ecology Letters 14:91–100.}

\hypertarget{citeproc_bib_item_13}{Blonder, B., C. Violle, and B. J. Enquist. 2013. Assessing the Causes and Scales of the Leaf Economics Spectrum Using Venation Networks Inpopulus Tremuloides. Journal of Ecology 101:981–989.}

\hypertarget{citeproc_bib_item_14}{Bocanegra, K., F. Fernández, and J. Galvis. 2015. Grupos Funcionales De Árboles En Bosques Secundarios De La Región Bajo Calima (Buenaventura, Colombia). Boletín Científico. Centro de Museos. Museo de Historia Natural 19:17–40.}

\hypertarget{citeproc_bib_item_15}{Bodegom, P. M. van, B. K. Sorrell, A. Oosthoek, C. Bakker, and R. Aerts. 2008. Separating the Effects of Partial Submergence and Soil Oxygen Demand on Plant Physiology. Ecology 89:193–204.}

\hypertarget{citeproc_bib_item_16}{Bond-Lamberty, B., S. T. Gower, C. Wang, P. Cyr, and H. Veldhuis. 2006. Nitrogen Dynamics of a Boreal Black Spruce Wildfire Chronosequence. Biogeochemistry 81:1–16.}

\hypertarget{citeproc_bib_item_17}{Bond-Lamberty, B., C. Wang, S. T. Gower, and J. Norman. 2002a. Leaf Area Dynamics of a Boreal Black Spruce Fire Chronosequence. Tree Physiology 22:993–1001.}

\hypertarget{citeproc_bib_item_18}{Bond-Lamberty, B., C. Wang, and S. T. Gower. 2002b. Aboveground and Belowground Biomass and Sapwood Area Allometric Equations for Six Boreal Tree Species of Northern Manitoba. Canadian Journal of Forest Research 32:1441–1450.}

\hypertarget{citeproc_bib_item_19}{Bond-Lamberty, B., C. Wang, and S. T. Gower. 2004. Net Primary Production and Net Ecosystem Production of a Boreal Black Spruce Wildfire Chronosequence. Global Change Biology 10:473–487.}

\hypertarget{citeproc_bib_item_20}{Brown, K. A., D. F. B. Flynn, N. K. Abram, J. C. Ingram, S. E. Johnson, and P. Wright. 2011. Assessing Natural Resource Use By Forest-Reliant Communities in Madagascar Using Functional Diversity and Functional Redundancy Metrics. PLoS ONE 6:e24107.}

\hypertarget{citeproc_bib_item_21}{Burrascano, S., R. Copiz, E. Del Vico, S. Fagiani, E. Giarrizzo, M. Mei, A. Mortelliti, F. M. Sabatini, and C. Blasi. 2015. Wild Boar Rooting Intensity Determines Shifts in Understorey Composition and Functional Traits. Community Ecology 16:244–253.}

\hypertarget{citeproc_bib_item_22}{Butterfield, B. J., and J. M. Briggs. 2010. Regeneration Niche Differentiates Functional Strategies of Desert Woody Plant Species. Oecologia 165:477–487.}

\hypertarget{citeproc_bib_item_23}{Byun, C., S. de Blois, and J. Brisson. 2012. Plant Functional Group Identity and Diversity Determine Biotic Resistance To Invasion By an Exotic Grass. Journal of Ecology 101:128–139.}

\hypertarget{citeproc_bib_item_24}{Campbell, C., L. Atkinson, J. Zaragoza-Castells, M. Lundmark, O. Atkin, and V. Hurry. 2007. Acclimation of Photosynthesis and Respiration Is Asynchronous in Response To Changes in Temperature Regardless of Plant Functional Group. New Phytologist 176:375–389.}

\hypertarget{citeproc_bib_item_25}{Campetella, G., Z. Botta-Dukát, C. Wellstein, R. Canullo, S. Gatto, S. Chelli, L. Mucina, and S. Bartha. 2011. Patterns of Plant Trait-Environment Relationships Along a Forest Succession Chronosequence. Agriculture, Ecosystems \& Environment 145:38–48.}

\hypertarget{citeproc_bib_item_26}{Carswell, F. E., P. Meir, E. V. Wandelli, L. C. M. Bonates, B. Kruijt, E. M. Barbosa, A. D. Nobre, J. Grace, and P. G. Jarvis. 2000. Photosynthetic Capacity in a Central Amazonian Rain Forest. Tree Physiology 20:179–186.}

\hypertarget{citeproc_bib_item_27}{Cavender-Bares, J., A. Keen, and B. Miles. 2006. Phylogenetic Structure of Floridian Plant Communities Depends on Taxonomic and Spatial Scale. Ecology 87:S109–S122.}

\hypertarget{citeproc_bib_item_28}{Cerabolini, B. E. L., G. Brusa, R. M. Ceriani, R. De Andreis, A. Luzzaro, and S. Pierce. 2010. Can CSR Classification Be Generally Applied Outside Britain? Plant Ecology 210:253–261.}

\hypertarget{citeproc_bib_item_29}{Chambers, J. Q., E. S. Tribuzy, L. C. Toledo, B. F. Crispim, N. Higuchi, J. d. Santos, A. C. Araújo, B. Kruijt, A. D. Nobre, and S. E. Trumbore. 2004. Respiration From a Tropical Forest Ecosystem: Partitioning of Sources and Low Carbon Use Efficiency. Ecological Applications 14:72–88.}

\hypertarget{citeproc_bib_item_30}{Chen, Y., W. Han, L. Tang, Z. Tang, and J. Fang. 2011. Leaf Nitrogen and Phosphorus Concentrations of Woody Plants Differ in Responses To Climate, Soil and Plant Growth Form. Ecography 36:178–184.}

\hypertarget{citeproc_bib_item_31}{Choat, B., S. Jansen, T. J. Brodribb, H. Cochard, S. Delzon, R. Bhaskar, S. J. Bucci, T. S. Feild, S. M. Gleason, U. G. Hacke, and et al. 2012. Global Convergence in the Vulnerability of Forests To Drought. Nature 491:752–755.}

\hypertarget{citeproc_bib_item_32}{Cornelissen, J. H. C. 1996. An Experimental Comparison of Leaf Decomposition Rates in a Wide Range of Temperate Plant Species and Types. The Journal of Ecology 84:573.}

\hypertarget{citeproc_bib_item_33}{Cornelissen, J. H. C., P. C. Diez, and R. Hunt. 1996. Seedling Growth, Allocation and Leaf Attributes in a Wide Range of Woody Plant Species and Types. The Journal of Ecology 84:755.}

\hypertarget{citeproc_bib_item_34}{Cornelissen, J. H. C., H. M. Quested, D. Gwynn-jones, R. S. P. van Logtestijn, M. A. H. De Beus, A. Kondratchuk, T. V. Callaghan, and R. Aerts. 2004. Leaf Digestibility and Litter Decomposability Are Related in a Wide Range of Subarctic Plant Species and Types. Functional Ecology 18:779–786.}

\hypertarget{citeproc_bib_item_35}{Cornelissen, J., B. Cerabolini, P. Castro-Díez, P. Villar-Salvador, G. Montserrat-Martí, J. Puyravaud, M. Maestro, M. Werger, and R. Aerts. 2003. Functional Traits of Woody Plants: Correspondence of Species Rankings Between Field Adults and Laboratory-Grown Seedlings? Journal of Vegetation Science 14:311–322.}

\hypertarget{citeproc_bib_item_36}{Cornwell, W. K., J. H. C. Cornelissen, K. Amatangelo, E. Dorrepaal, V. T. Eviner, O. Godoy, S. E. Hobbie, B. Hoorens, H. Kurokawa, N. Pérez-Harguindeguy, and et al. 2008. Plant Species Traits Are the Predominant Control on Litter Decomposition Rates Within Biomes Worldwide. Ecology Letters 11:1065–1071.}

\hypertarget{citeproc_bib_item_37}{Craine, J. M., A. J. Elmore, M. P. M. Aidar, M. Bustamante, T. E. Dawson, E. A. Hobbie, A. Kahmen, M. C. Mack, K. K. McLauchlan, A. Michelsen, and et al. 2009. Global Patterns of Foliar Nitrogen Isotopes and Their Relationships With Climate, Mycorrhizal Fungi, Foliar Nutrient Concentrations, and Nitrogen Availability. New Phytologist 183:980–992.}

\hypertarget{citeproc_bib_item_38}{Craine, J. M., W. G. Lee, W. J. Bond, R. J. Williams, and L. C. Johnson. 2005. Environmental Constraints on a Global Relationship Among Leaf and Root Traits of Grasses. Ecology 86:12–19.}

\hypertarget{citeproc_bib_item_39}{Craine, J. M., J. B. Nippert, E. G. Towne, S. Tucker, S. W. Kembel, A. Skibbe, and K. K. McLauchlan. 2011. Functional Consequences of Climate Change-Induced Plant Species Loss in a Tallgrass Prairie. Oecologia 165:1109–1117.}

\hypertarget{citeproc_bib_item_40}{Craine, J. M., E. G. Towne, T. W. Ocheltree, and J. B. Nippert. 2012. Community Traitscape of Foliar Nitrogen Isotopes Reveals N Availability Patterns in a Tallgrass Prairie. Plant and Soil 356:395–403.}

\hypertarget{citeproc_bib_item_41}{Craven, D., D. Braden, M. Ashton, G. Berlyn, M. Wishnie, and D. Dent. 2007. Between and Within-Site Comparisons of Structural and Physiological Characteristics and Foliar Nutrient Content of 14 Tree Species At a Wet, Fertile Site and a Dry, Infertile Site in Panama. Forest Ecology and Management 238:335–346.}

\hypertarget{citeproc_bib_item_42}{Demey, A., J. Staelens, L. Baeten, P. Boeckx, M. Hermy, J. Kattge, and K. Verheyen. 2013. Nutrient Input From Hemiparasitic Litter Favors Plant Species With a Fast-Growth Strategy. Plant and Soil 371:53–66.}

\hypertarget{citeproc_bib_item_43}{Diaz, S., J. Hodgson, K. Thompson, M. Cabido, J. Cornelissen, A. Jalili, G. Montserrat-Martí, J. Grime, F. Zarrinkamar, Y. Asri, and et al. 2004. The Plant Traits That Drive Ecosystems: Evidence From Three Continents. Journal of Vegetation Science 15:295–304.}

\hypertarget{citeproc_bib_item_44}{Domingues, T. F., L. A. Martinelli, and J. R. Ehleringer. 2013. Seasonal Patterns of Leaf-Level Photosynthetic Gas Exchange in an Eastern Amazonian Rain Forest. Plant Ecology \& Diversity 7:189–203.}

\hypertarget{citeproc_bib_item_45}{Domingues, T. F., P. Meir, T. R. Feldpausch, G. Saiz, E. M. Veenendaal, F. Schrodt, M. Bird, G. Djagbletey, F. Hien, H. Compaore, and et al. 2010. Co-Limitation of Photosynthetic Capacity By Nitrogen and Phosphorus in West Africa Woodlands. Plant, Cell \& Environment 33:959–980.}

\hypertarget{citeproc_bib_item_46}{Fitter, A. H., and H. J. Peat. 1994. The Ecological Flora Database. The Journal of Ecology 82:415.}

\hypertarget{citeproc_bib_item_47}{Fonseca, C. R., J. M. Overton, B. Collins, and M. Westoby. 2000. Shifts in Trait-Combinations Along Rainfall and Phosphorus Gradients. Journal of Ecology 88:964–977.}

\hypertarget{citeproc_bib_item_48}{Frenette-Dussault, C., B. Shipley, J.-F. Léger, D. Meziane, and Y. Hingrat. 2011. Functional Structure of an Arid Steppe Plant Community Reveals Similarities With Grime’s C-S-R Theory. Journal of Vegetation Science 23:208–222.}

\hypertarget{citeproc_bib_item_49}{Fyllas, N. M., S. Patiño, T. R. Baker, G. Bielefeld Nardoto, L. A. Martinelli, C. A. Quesada, R. Paiva, M. Schwarz, V. Horna, L. M. Mercado, and et al. 2009. Basin-Wide Variations in Foliar Properties of Amazonian Forest: Phylogeny, Soils and Climate. Biogeosciences 6:2677–2708.}

\hypertarget{citeproc_bib_item_50}{Gallagher, R. V., and M. R. Leishman. 2012. A Global Analysis of Trait Variation and Evolution in Climbing Plants. Journal of Biogeography 39:1757–1771.}

\hypertarget{citeproc_bib_item_51}{Garnier, E., S. Lavorel, P. Ansquer, H. Castro, P. Cruz, J. Dolezal, O. Eriksson, C. Fortunel, H. Freitas, C. Golodets, and et al. 2007. Assessing the Effects of Land-Use Change on Plant Traits, Communities and Ecosystem Functioning in Grasslands: a Standardized Methodology and Lessons From an Application To 11 European Sites. Annals of Botany 99:967–985.}

\hypertarget{citeproc_bib_item_52}{Givnish, T. J., R. A. Montgomery, and G. Goldstein. 2004. Adaptive Radiation of Photosynthetic Physiology in the Hawaiian Lobeliads: Light Regimes, Static Light Responses, and Whole-Plant Compensation Points. American Journal of Botany 91:228–246.}

\hypertarget{citeproc_bib_item_53}{Guerin, G. R., H. Wen, and A. J. Lowe. 2012. Leaf Morphology Shift Linked To Climate Change. Biology Letters 8:882–886.}

\hypertarget{citeproc_bib_item_54}{Gutiérrez, A. G., and A. Huth. 2012. Successional Stages of Primary Temperate Rainforests of Chiloé Island, Chile. Perspectives in Plant Ecology, Evolution and Systematics 14:243–256.}

\hypertarget{citeproc_bib_item_55}{Guy, A. L., J. M. Mischkolz, and E. G. Lamb. 2013. Limited Effects of Simulated Acidic Deposition on Seedling Survivorship and Root Morphology of Endemic Plant Taxa of the Athabasca Sand Dunes in Well-Watered Greenhouse Trials. Botany 91:176–181.}

\hypertarget{citeproc_bib_item_56}{Han, W., J. Fang, D. Guo, and Y. Zhang. 2005. Leaf Nitrogen and Phosphorus Stoichiometry Across 753 Terrestrial Plant Species in China. New Phytologist 168:377–385.}

\hypertarget{citeproc_bib_item_57}{Hickler, T. 1999. Plant Functional Types and community characteristics along environmental gradients on Öland’s Great Alvar. University of Lund, Lund, Sweden.}

\hypertarget{citeproc_bib_item_58}{Kattge, J., W. Knorr, T. Raddatz, and C. Wirth. 2009. Quantifying Photosynthetic Capacity and Its Relationship To Leaf Nitrogen Content for Global-Scale Terrestrial Biosphere Models. Global Change Biology 15:976–991.}

\hypertarget{citeproc_bib_item_59}{Kazakou, E., D. Vile, B. Shipley, C. Gallet, and E. Garnier. 2006. Co-Variations in Litter Decomposition, Leaf Traits and Plant Growth in Species From a Mediterranean Old-Field Succession. Functional Ecology 20:21–30.}

\hypertarget{citeproc_bib_item_60}{Kerkhoff, A. J., W. F. Fagan, J. J. Elser, and B. J. Enquist. 2006. Phylogenetic and Growth Form Variation in the Scaling of Nitrogen and Phosphorus in the Seed Plants. The American Naturalist 168:E103–E122.}

\hypertarget{citeproc_bib_item_61}{Kichenin, E., D. A. Wardle, D. A. Peltzer, C. W. Morse, and G. T. Freschet. 2013. Contrasting Effects of Plant Inter- and Intraspecific Variation on Community-Level Trait Measures Along an Environmental Gradient. Functional Ecology 27:1254–1261.}

\hypertarget{citeproc_bib_item_62}{Kisel, Y., A. C. Moreno-Letelier, D. Bogarín, M. P. Powell, M. W. Chase, and T. G. Barraclough. 2012. Testing the Link Between Population Genetic Differentiation and Clade Diversification in Costa Rican Orchids. Evolution 66:3035–3052.}

\hypertarget{citeproc_bib_item_63}{Kleyer, M., R. Bekker, I. Knevel, J. Bakker, K. Thompson, M. Sonnenschein, P. Poschlod, J. van Groenendael, L. Klimevs, J. Klimevsová, and et al. 2008. The LEDA Traitbase: a Database of Life-History Traits of the Northwest European Flora. Journal of Ecology 96:1266–1274.}

\hypertarget{citeproc_bib_item_64}{Kraft, N. J. B., R. Valencia, and D. D. Ackerly. 2008. Functional Traits and Niche-Based Tree Community Assembly in an Amazonian Forest. Science 322:580–582.}

\hypertarget{citeproc_bib_item_65}{Laughlin, D. C., P. Z. Fulé, D. W. Huffman, J. Crouse, and E. Laliberté. 2011. Climatic Constraints on Trait-Based Forest Assembly. Journal of Ecology 99:1489–1499.}

\hypertarget{citeproc_bib_item_66}{Loranger, J., S. T. Meyer, B. Shipley, J. Kattge, H. Loranger, C. Roscher, and W. W. Weisser. 2012. Predicting Invertebrate Herbivory From Plant Traits: Evidence From 51 Grassland Species in Experimental Monocultures. Ecology 93:2674–2682.}

\hypertarget{citeproc_bib_item_67}{Louault, F., V. Pillar, J. Aufrère, E. Garnier, and J.-F. Soussana. 2005. Plant Traits and Functional Types in Response To Reduced Disturbance in a Semi-Natural Grassland. Journal of Vegetation Science 16:151–160.}

\hypertarget{citeproc_bib_item_68}{Loveys, B. R., L. J. Atkinson, D. J. Sherlock, R. L. Roberts, A. H. Fitter, and O. K. Atkin. 2003. Thermal Acclimation of Leaf and Root Respiration: an Investigation Comparing Inherently Fast- and Slow-Growing Plant Species. Global Change Biology 9:895–910.}

\hypertarget{citeproc_bib_item_69}{Manzoni, S., G. Vico, S. Palmroth, A. Porporato, and G. Katul. 2013. Optimization of Stomatal Conductance for Maximum Carbon Gain Under Dynamic Soil Moisture. Advances in Water Resources 62:90–105.}

\hypertarget{citeproc_bib_item_70}{Medlyn, B. E., F. .-W. Badeck, D. G. G. De Pury, C. V. M. Barton, M. Broadmeadow, R. Ceulemans, P. De Angelis, M. Forstreuter, M. E. Jach, S. Kellomäki, and et al. 1999. Effects of Elevated [CO2] on Photosynthesis in European Forest Species: a Meta-Analysis of Model Parameters. Plant, Cell \& Environment 22:1475–1495.}

\hypertarget{citeproc_bib_item_71}{Meir, P., B. Kruijt, M. Broadmeadow, E. Barbosa, O. Kull, F. Carswell, A. Nobre, and P. G. Jarvis. 2002. Acclimation of Photosynthetic Capacity To Irradiance in Tree Canopies in Relation To Leaf Nitrogen Concentration and Leaf Mass Per Unit Area. Plant, Cell and Environment 25:343–357.}

\hypertarget{citeproc_bib_item_72}{Meir, P., P. E. Levy, J. Grace, and P. G. Jarvis. 2007. Photosynthetic Parameters From Two Contrasting Woody Vegetation Types in West Africa. Plant Ecology 192:277–287.}

\hypertarget{citeproc_bib_item_73}{Messier, J., B. J. McGill, and M. J. Lechowicz. 2010. How Do Traits Vary Across Ecological Scales? a Case for Trait-Based Ecology. Ecology Letters 13:838–848.}

\hypertarget{citeproc_bib_item_74}{Meziane, D., and B. Shipley. 1999. Interacting Determinants of Specific Leaf Area in 22 Herbaceous Species: Effects of Irradiance and Nutrient Availability. Plant, Cell \& Environment 22:447–459.}

\hypertarget{citeproc_bib_item_75}{Milla, R., and P. B. Reich. 2011. Multi-Trait Interactions, Not Phylogeny, Fine-Tune Leaf Size Reduction With Increasing Altitude. Annals of Botany 107:455–465.}

\hypertarget{citeproc_bib_item_76}{Minden, V., S. Andratschke, J. Spalke, H. Timmermann, and M. Kleyer. 2012. Plant Trait-Environment Relationships in Salt Marshes: Deviations From Predictions By Ecological Concepts. Perspectives in Plant Ecology, Evolution and Systematics 14:183–192.}

\hypertarget{citeproc_bib_item_77}{Minden, V., and M. Kleyer. 2011. Testing the Effect-Response Framework: Key Response and Effect Traits Determining Above-Ground Biomass of Salt Marshes. Journal of Vegetation Science 22:387–401.}

\hypertarget{citeproc_bib_item_78}{Müller, S. C., G. E. Overbeck, J. Pfadenhauer, and V. D. Pillar. 2006. Plant Functional Types of Woody Species Related To Fire Disturbance in Forest-Grassland Ecotones. Plant Ecology 189:1–14.}

\hypertarget{citeproc_bib_item_79}{Niinemets, Ü. 2001. Global-Scale Climatic Controls of Leaf Dry Mass Per Area, Density, and Thickness in Trees and Shrubs. Ecology 82:453–469.}

\hypertarget{citeproc_bib_item_80}{Ogaya, R., and J. Peñuelas. 2003. Comparative Field Study of Quercus Ilex and Phillyrea Latifolia: Photosynthetic Response To Experimental Drought Conditions. Environmental and Experimental Botany 50:137–148.}

\hypertarget{citeproc_bib_item_81}{Onoda, Y., M. Westoby, P. B. Adler, A. M. F. Choong, F. J. Clissold, J. H. C. Cornelissen, S. Díaz, N. J. Dominy, A. Elgart, L. Enrico, and et al. 2011. Global Patterns of Leaf Mechanical Properties. Ecology Letters 14:301–312.}

\hypertarget{citeproc_bib_item_82}{Ordoñez, J. C., P. M. van Bodegom, Witte Jan‐Philip M., R. P. Bartholomeus, J. R. van Hal, and R. Aerts. 2010. Plant Strategies in Relation To Resource Supply in Mesic To Wet Environments: Does Theory Mirror Nature? The American Naturalist 175:225–239.}

\hypertarget{citeproc_bib_item_83}{Pahl, A. T., J. Kollmann, A. Mayer, and S. Haider. 2013. No Evidence for Local Adaptation in an Invasive Alien Plant: Field and Greenhouse Experiments Tracing a Colonization Sequence. Annals of Botany 112:1921–1930.}

\hypertarget{citeproc_bib_item_84}{Peco, B., I. de Pablos, J. Traba, and C. Levassor. 2005. The Effect of Grazing Abandonment on Species Composition and Functional Traits: the Case of Dehesa Grasslands. Basic and Applied Ecology 6:175–183.}

\hypertarget{citeproc_bib_item_85}{Penuelas, J., J. Sardans, J. Llusià, S. M. Owen, J. Carnicer, T. W. Giambelluca, E. L. Rezende, M. Waite, and Ü. Niinemets. 2009. Faster Returns on “Leaf Economics” and Different Biogeochemical Niche in Invasive Compared With Native Plant Species. Global Change Biology 16:2171–2185.}

\hypertarget{citeproc_bib_item_86}{Pierce, S., G. Brusa, M. Sartori, and B. E. L. Cerabolini. 2012. Combined Use of Leaf Size and Economics Traits Allows Direct Comparison of Hydrophyte and Terrestrial Herbaceous Adaptive Strategies. Annals of Botany 109:1047–1053.}

\hypertarget{citeproc_bib_item_87}{Pierce, S., G. Brusa, I. Vagge, and B. E. L. Cerabolini. 2013. Allocating Csr Plant Functional Types: the Use of Leaf Economics and Size Traits To Classify Woody and Herbaceous Vascular Plants. Functional Ecology 27:1002–1010.}

\hypertarget{citeproc_bib_item_88}{Pierce, S., A. Luzzaro, M. Caccianiga, R. M. Ceriani, and B. Cerabolini. 2007. Disturbance Is the Principal $alpha$-scale Filter Determining Niche Differentiation, Coexistence and Biodiversity in an Alpine Community. Journal of Ecology 95:698–706.}

\hypertarget{citeproc_bib_item_89}{Pillar, V. D., and E. E. Sosinski. 2003. An Improved Method for Searching Plant Functional Types By Numerical Analysis. Journal of Vegetation Science 14:323–332.}

\hypertarget{citeproc_bib_item_90}{Poorter, H., Ü. Niinemets, L. Poorter, I. J. Wright, and R. Villar. 2009. Causes and Consequences of Variation in Leaf Mass Per Area (LMA): a Meta-Analysis. New Phytologist 182:565–588.}

\hypertarget{citeproc_bib_item_91}{Powers, J. S., and P. Tiffin. 2010. Plant Functional Type Classifications in Tropical Dry Forests in Costa Rica: Leaf Habit Versus Taxonomic Approaches. Functional Ecology 24:927–936.}

\hypertarget{citeproc_bib_item_92}{Prentice, I. C., T. Meng, H. Wang, S. P. Harrison, J. Ni, and G. Wang. 2010. Evidence of a Universal Scaling Relationship for Leaf CO2 Drawdown Along an Aridity Gradient. New Phytologist 190:169–180.}

\hypertarget{citeproc_bib_item_93}{Preston, K. A., W. K. Cornwell, and J. L. DeNoyer. 2006. Wood Density and Vessel Traits As Distinct Correlates of Ecological Strategy in 51 California Coast Range Angiosperms. New Phytologist 170:807–818.}

\hypertarget{citeproc_bib_item_94}{Price, C. A., and B. J. Enquist. 2007. Scaling Mass and Morphology in Leaves: an Extension of the Wbe Model. Ecology 88:1132–1141.}

\hypertarget{citeproc_bib_item_95}{Pyankov, V. I., A. V. Kondratchuk, and B. Shipley. 1999. Leaf Structure and Specific Leaf Mass: the Alpine Desert Plants of the Eastern Pamirs, Tadjikistan. New Phytologist 143:131–142.}

\hypertarget{citeproc_bib_item_96}{Quested, H. M., J. H. C. Cornelissen, M. C. Press, T. V. Callaghan, R. Aerts, F. Trosien, P. Riemann, D. Gwynn-Jones, A. Kondratchuk, and S. E. Jonasson. 2003. Decomposition of Sub-Arctic Plants With Differing Nitrogen Economies: a Functional Role for Hemiparasites. Ecology 84:3209–3221.}

\hypertarget{citeproc_bib_item_97}{Reich, P. B., M. G. Tjoelker, K. S. Pregitzer, I. J. Wright, J. Oleksyn, and J.-L. Machado. 2008. Scaling of Respiration To Nitrogen in Leaves, Stems and Roots of Higher Land Plants. Ecology Letters 11:793–801.}

\hypertarget{citeproc_bib_item_98}{Rüger, N., A. Huth, S. P. Hubbell, and R. Condit. 2009. Response of Recruitment To Light Availability Across a Tropical Lowland Rain Forest Community. Journal of Ecology 97:1360–1368.}

\hypertarget{citeproc_bib_item_99}{Rüger, N., A. Huth, S. P. Hubbell, and R. Condit. 2011. Determinants of Mortality Across a Tropical Lowland Rainforest Community. Oikos 120:1047–1056.}

\hypertarget{citeproc_bib_item_100}{Sandel, B., J. D. Corbin, and M. Krupa. 2011. Using Plant Functional Traits To Guide Restoration: a Case Study in California Coastal Grassland. Ecosphere 2:art23.}

\hypertarget{citeproc_bib_item_101}{Schererlorenzen, M., E. Schulze, A. Don, J. Schumacher, and E. Weller. 2007. Exploring the Functional Significance of Forest Diversity: a New Long-Term Experiment With Temperate Tree Species (BIOTREE). Perspectives in Plant Ecology, Evolution and Systematics 9:53–70.}

\hypertarget{citeproc_bib_item_102}{Schweingruber, F., and W. Landolt. 2005. The xylem database.}

\hypertarget{citeproc_bib_item_103}{Shiodera, S., J. S. Rahajoe, and T. Kohyama. 2008. Variation in Longevity and Traits of Leaves Among Co-Occurring Understorey Plants in a Tropical Montane Forest. Journal of Tropical Ecology 24:121–133.}

\hypertarget{citeproc_bib_item_104}{Shipley, B. 1995. Structured Interspecific Determinants of Specific Leaf Area in 34 Species of Herbaceous Angiosperms. Functional Ecology 9:312.}

\hypertarget{citeproc_bib_item_105}{Shipley, B. 2002. Trade-Offs Between Net Assimilation Rate and Specific Leaf Area in Determining Relative Growth Rate: Relationship With Daily Irradiance. Functional Ecology 16:682–689.}

\hypertarget{citeproc_bib_item_106}{Shipley, B., and M. J. Lechowicz. 2000. The Functional Co-Ordination of Leaf Morphology, Nitrogen Concentration, and Gas Exchange In40 Wetland Species. Écoscience 7:183–194.}

\hypertarget{citeproc_bib_item_107}{Shipley, B., and T.-T. Vu. 2002. Dry Matter Content As a Measure of Dry Matter Concentration in Plants and Their Parts. New Phytologist 153:359–364.}

\hypertarget{citeproc_bib_item_108}{Spasojevic, M. J., and K. N. Suding. 2012. Inferring Community Assembly Mechanisms From Functional Diversity Patterns: the Importance of Multiple Assembly Processes. Journal of Ecology 100:652–661.}

\hypertarget{citeproc_bib_item_109}{Swaine, E. 2007. Ecological and evolutionary drivers of plant community assembly in a Bornean rain forest. University of Aberdeen, Aberdeen, Scotland, UK.}

\hypertarget{citeproc_bib_item_110}{Tucker, S. S., J. M. Craine, and J. B. Nippert. 2011. Physiological Drought Tolerance and the Structuring of Tallgrass Prairie Assemblages. Ecosphere 2:art48.}

\hypertarget{citeproc_bib_item_111}{Veihmeyer, F. J. 1956. Soil Moisture. Pflanze und Wasser / Water Relations of Plants:64–123.}

\hypertarget{citeproc_bib_item_112}{Vergutz, L., S. Manzoni, A. Porporato, R. Novais, and R. Jackson. 2012. A Global Database of Carbon and Nutrient Concentrations of Green and Senesced Leaves.}

\hypertarget{citeproc_bib_item_113}{Vile, D. 2005. Significations fonctionnelle et ecologique des traits des especes vegetales: Exemple dans une succession post-cultural Mediterraneenne et generalisations.}

\hypertarget{citeproc_bib_item_114}{Von Holle, B., and D. Simberloff. 2004. Testing Fox’s Assembly Rule: Does Plant Invasion Depend on Recipient Community Structure? Oikos 105:551–563.}

\hypertarget{citeproc_bib_item_115}{Weedon, J. T., W. K. Cornwell, J. H. Cornelissen, A. E. Zanne, C. Wirth, and D. A. Coomes. 2009. Global Meta-Analysis of Wood Decomposition Rates: a Role for Trait Variation Among Tree Species? Ecology Letters 12:45–56.}

\hypertarget{citeproc_bib_item_116}{Williams, M., Y. Shimabokuro, and E. Rastetter. 2012. LBA-ECO CD-09 Soil and Vegetation Characteristics, Tapajos National Forest, Brazil.}

\hypertarget{citeproc_bib_item_117}{Willis, C. G., M. Halina, C. Lehman, P. B. Reich, A. Keen, S. McCarthy, and J. Cavender-Bares. 2009. Phylogenetic Community Structure in Minnesota Oak Savanna Is Influenced By Spatial Extent and Environmental Variation. Ecography:no-no.}

\hypertarget{citeproc_bib_item_118}{Wilson, K. B., D. D. Baldocchi, and P. J. Hanson. 2000. Spatial and Seasonal Variability of Photosynthetic Parameters and Their Relationship To Leaf Nitrogen in a Deciduous Forest. Tree Physiology 20:565–578.}

\hypertarget{citeproc_bib_item_119}{Wirth, C., and J. W. Lichstein. 2009. The Imprint of Species Turnover on Old-Growth Forest Carbon Balances - Insights From a Trait-Based Model of Forest Dynamics. Ecological Studies:81–113.}

\hypertarget{citeproc_bib_item_120}{Wohlfahrt, G., M. Bahn, E. Haubner, I. Horak, W. Michaeler, K. Rottmar, U. Tappeiner, and A. Cernusca. 1999. Inter-Specific Variation of the Biochemical Limitation To Photosynthesis and Related Leaf Traits of 30 Species From Mountain Grassland Ecosystems Under Different Land Use. Plant, Cell and Environment 22:1281–1296.}

\hypertarget{citeproc_bib_item_121}{Wright, I. J., D. D. Ackerly, F. Bongers, K. E. Harms, G. Ibarra-Manriquez, M. Martinez-Ramos, S. J. Mazer, H. C. Muller-Landau, H. Paz, N. C. A. Pitman, and et al. 2006. Relationships Among Ecologically Important Dimensions of Plant Trait Variation in Seven Neotropical Forests. Annals of Botany 99:1003–1015.}

\hypertarget{citeproc_bib_item_122}{Wright, I. J., P. B. Reich, M. Westoby, D. D. Ackerly, Z. Baruch, F. Bongers, J. Cavender-Bares, T. Chapin, J. H. C. Cornelissen, M. Diemer, and et al. 2004. The Worldwide Leaf Economics Spectrum. Nature 428:821–827.}

\hypertarget{citeproc_bib_item_123}{Wright, J. P., and A. Sutton-Grier. 2012. Does the Leaf Economic Spectrum Hold Within Local Species Pools Across Varying Environmental Conditions? Functional Ecology 26:1390–1398.}

\hypertarget{citeproc_bib_item_124}{Wright, S. J., K. Kitajima, N. J. B. Kraft, P. B. Reich, I. J. Wright, D. E. Bunker, R. Condit, J. W. Dalling, S. J. Davies, S. Díaz, and et al. 2010. Functional Traits and the Growth-Mortality Trade-Off in Tropical Trees. Ecology 91:3664–3674.}

\hypertarget{citeproc_bib_item_125}{Xu, L., and D. D. Baldocchi. 2003. Seasonal Trends in Photosynthetic Parameters and Stomatal Conductance of Blue Oak (Quercus douglasii) Under Prolonged Summer Drought and High Temperature. Tree Physiology 23:865–877.}

\hypertarget{citeproc_bib_item_126}{Yguel, B., R. Bailey, N. D. Tosh, A. Vialatte, C. Vasseur, X. Vitrac, F. Jean, and A. Prinzing. 2011. Phytophagy on Phylogenetically Isolated Trees: Why Hosts Should Escape Their Relatives. Ecology Letters 14:1117–1124.}

\hypertarget{citeproc_bib_item_127}{Zapata-Cuartas, M., C. A. Sierra, and L. Alleman. 2012. Probability Distribution of Allometric Coefficients and Bayesian Estimation of Aboveground Tree Biomass. Forest Ecology and Management 277:173–179.}
\end{hangparas}
\end{document}

### 2019-09-09 authors.tex
Fix Appendix 1. Refactor authors file.

\usepackage[noblocks]{authblk}

\author[1,*]{Alexey N. Shiklomanov <alexey.shiklomanov@pnnl.gov>}
\affil[1]{Joint Global Change Research Institute, Pacific Northwest National Laboratory, College Park, MD}

\author[2]{Elizabeth M. Cowdery <ecowdery@bu.edu>}
\affil[2]{Department of Earth \& Environment, Boston University, Boston, MA}

\author[3]{Michael Bahn <Michael.Bahn@uibk.ac.at>}
\affil[3]{Institute of Ecology, University of Innsbruck, 6020 Innsbruck, Austria}

\author[4]{Chaeho Byun <chaeho.byun@mail.mcgill.ca>}
\affil[4]{School of Civil and Environmental Engineering, Yonsei University, Seoul 03722, Korea}

\author[5]{Steven Jansen <steven.jansen@uni-ulm.de>}
\affil[5]{Institute of Systematic Botany and Ecology, Ulm University, Albert-Einstein-Allee 11, 89081, Ulm, Germany}

\author[6]{Koen Kramer <koen.kramer@wur.nl>}
\affil[6]{Department of Vegetation, Forest, and Landscape Ecology, Wageningen Environmental Research and Wageningen University, Wageningen, Gelderland, The Netherlands}

\author[7,8]{Vanessa Minden <vanessa.minden@uni-oldenburg.de>}
\affil[7]{Institute for Biology and Environmental Sciences, Carl von Ossietzky-University of Oldenburg, Carl von Ossietzky Str. 9-11, 26129 Oldenburg, Germany}
\affil[8]{Department of Biology, Ecology and Evolution, Vrije Universiteit Brussel, Pleinlaan 2, 1050 Brussels}

\author[9]{\"Ulo Niinemets <ylo.niinemets@emu.ee>}
\affil[9]{Institute of Agricultural and Environmental Sciences, Estonian University of Life Sciences, Kreutzwaldi 1, 51014 Tartu, Estonia}

\author[10]{Yusuke Onoda <yusuke.onoda@gmail.com>}
\affil[10]{Graduate School of Agriculture, Kyoto University, Kyoto, 605-8503, Japan}

\author[11]{Nadejda A. Soudzilovskaia <n.a.soudzilovskaia@cml.leidenuniv.nl>}
\affil[11]{Conservation Biology Department, Institute of Environmental Sciences, Leiden University, Leiden, The Netherlands}

\author[2]{Michael C. Dietze <dietze@bu.edu>}

\affil[*]{Corresponding author; Phone: (301) 314-6713; Fax: (301) 314-6719; Email: alexey.shiklomanov@pnnl.gov; Mail: 5825 University Research Ct., Office 3533, College Park, MD 20740}

\renewcommand\Authfont{\footnotesize}
\renewcommand\Affilfont{\scriptsize}

### 2019-09-09 AppendixS1.tex
Fix Appendix S1 formatting.

@@
-% Created 2019-09-09 Mon 08:53
+% Created 2019-09-09 Mon 09:26
 % Intended LaTeX compiler: pdflatex
@@
 \input{authors.tex}
-\usepackage{hanging}
 \date{}
@@
 
-\section*{Appendix S1: References for the original sources of the TRY trait data used in our analysis.}
+\noindent
+\textbf{Appendix S1:} References for the original sources of the TRY trait data used in our analysis.
 
-\begin{hangparas}{1.5em}{1}\hypertarget{citeproc_bib_item_1}{Adler, P.
-B., D.
-G.
-Milchunas, W.
-K.
-Lauenroth, O.
-E.
-Sala, and I.
-C.
-Burke.
-2004.
-Functional Traits of Graminoids in Semi-Arid Steppes: a Test of Grazing Histories.
-Journal of Applied Ecology 41:653–663.}
+\vspace{\baselineskip}
 
-\hypertarget{citeproc_bib_item_2}{Adriaenssens, S. 2012. Dry deposition and canopy exchange for temperate tree species under high nitrogen deposition.}
+\nocite{kleyer_2008_leda_traitbase}
 
-\hypertarget{citeproc_bib_item_3}{Akhmetzhanova, A.
-A., N.
-A.
-Soudzilovskaia, V.
-G.
-Onipchenko, W.
-K.
-Cornwell, V.
-A.
-Agafonov, I.
-A.
-Selivanov, and J.
-H.
-C.
-Cornelissen.
-2012.
-A Rediscovered Treasure: Mycorrhizal Intensity Database for 3000 Vascular Plant Species Across the Former Soviet Union.
-Ecology 93:689–690.}
+\nocite{wirth_2009_imprint}
 
-\hypertarget{citeproc_bib_item_4}{Araujo, A.
-de, J.
-Ometto, A.
-Dolman, B.
-Kruijt, M.
-Waterloo, and J.
-Ehleringer.
-2012.
-LBA-ECO CD-02 C and N Isotopes in Leaves and Atmospheric CO2, Amazonas, Brazil.
-ORNL Distributed Active Archive Center.}
+\nocite{baraloto_2010_decoupled}
 
-\hypertarget{citeproc_bib_item_5}{Atkin, O.
-K., M.
-Schortemeyer, N.
-McFarlane, and J.
-R.
-Evans.
-1999.
-The Response of Fast- and Slow-Growing Acacia Species To Elevated Atmospheric CO2: An Analysis of the Underlying Components of Relative Growth Rate.
-Oecologia 120:544–554.}
+\nocite{craine_2009_global}
 
-\hypertarget{citeproc_bib_item_6}{Auger, S., and B.
-Shipley.
-2012.
-Inter-Specific and Intra-Specific Trait Variation Along Short Environmental Gradients in an Old-Growth Temperate Forest.
-Journal of Vegetation Science 24:419–428.}
+\nocite{veihmeyer_1956_soil_moisture}
 
-\hypertarget{citeproc_bib_item_7}{Bakker, C., J.
-Rodenburg, and P.
-M.
-van Bodegom.
-2005.
-Effects of Ca- and Fe-Rich Seepage on P Availability and Plant Performance in Calcareous Dune Soils.
-Plant and Soil 275:111–122.}
+\nocite{cerabolini_2010_can}
 
-\hypertarget{citeproc_bib_item_8}{Bakker, C., P.
-M.
-Van Bodegom, H.
-J.
-M.
-Nelissen, W.
-H.
-O.
-Ernst, and R.
-Aerts.
-2006.
-Plant Responses To Rising Water Tables and Nutrient Management in Calcareous Dune Slacks.
-Plant Ecology 185:19–28.}
+\nocite{wright_2010_functional}
 
-\hypertarget{citeproc_bib_item_9}{Baraloto, C., C.
-E.
-T.
-Paine, L.
-Poorter, J.
-Beauchene, D.
-Bonal, A.-M.
-Domenach, B.
-Hérault, S.
-Patiño, J.-C.
-Roggy, and J.
-Chave.
-2010.
-Decoupled Leaf and Stem Economics in Rain Forest Trees.
-Ecology Letters 13:1338–1347.}
+\nocite{wright_2004_worldwide}
 
-\hypertarget{citeproc_bib_item_10}{Beckmann, M., M.
-Hock, H.
-Bruelheide, and A.
-Erfmeier.
-2012.
-The Role of UV-B Radiation in the Invasion of Hieracium pilosella-A Comparison of German and New Zealand Plants.
-Environmental and Experimental Botany 75:173–180.}
+\nocite{ogaya_2003_comparative}
 
-\hypertarget{citeproc_bib_item_11}{Blonder, B., V.
-Buzzard, I.
-Simova, L.
-Sloat, B.
-Boyle, R.
-Lipson, B.
-Aguilar-Beaucage, A.
-Andrade, B.
-Barber, C.
-Barnes, and et al.
-2012.
-The Leaf-Area Shrinkage Effect Can Bias Paleoclimate and Ecology Research.
-American Journal of Botany 99:1756–1763.}
+\nocite{akhmetzhanova_2012_rediscovered_treasure}
 
-\hypertarget{citeproc_bib_item_12}{Blonder, B., C. Violle, L. P. Bentley, and B. J. Enquist. 2010. Venation Networks and the Origin of the Leaf Economics Spectrum. Ecology Letters 14:91–100.}
+\nocite{frenette-dussault_2011_functional}
 
-\hypertarget{citeproc_bib_item_13}{Blonder, B., C.
-Violle, and B.
-J.
-Enquist.
-2013.
-Assessing the Causes and Scales of the Leaf Economics Spectrum Using Venation Networks Inpopulus Tremuloides.
-Journal of Ecology 101:981–989.}
+\nocite{medlyn_1999_effects}
 
-\hypertarget{citeproc_bib_item_14}{Bocanegra, K., F.
-Fernández, and J.
-Galvis.
-2015.
-Grupos Funcionales De Árboles En Bosques Secundarios De La Región Bajo Calima (Buenaventura, Colombia).
-Boletín Científico.
-Centro de Museos.
-Museo de Historia Natural 19:17–40.}
+\nocite{kattge_2009_quantifying}
 
-\hypertarget{citeproc_bib_item_15}{Bodegom, P.
-M.
-van, B.
-K.
-Sorrell, A.
-Oosthoek, C.
-Bakker, and R.
-Aerts.
-2008.
-Separating the Effects of Partial Submergence and Soil Oxygen Demand on Plant Physiology.
-Ecology 89:193–204.}
+\nocite{laughlin_2011_climatic}
 
-\hypertarget{citeproc_bib_item_16}{Bond-Lamberty, B., S. T. Gower, C. Wang, P. Cyr, and H. Veldhuis. 2006. Nitrogen Dynamics of a Boreal Black Spruce Wildfire Chronosequence. Biogeochemistry 81:1–16.}
+\nocite{campetella_2011_patterns}
 
-\hypertarget{citeproc_bib_item_17}{Bond-Lamberty, B., C. Wang, S. T. Gower, and J. Norman. 2002a. Leaf Area Dynamics of a Boreal Black Spruce Fire Chronosequence. Tree Physiology 22:993–1001.}
+\nocite{auger_2012_inter}
 
-\hypertarget{citeproc_bib_item_18}{Bond-Lamberty, B., C.
-Wang, and S.
-T.
-Gower.
-2002b.
-Aboveground and Belowground Biomass and Sapwood Area Allometric Equations for Six Boreal Tree Species of Northern Manitoba.
-Canadian Journal of Forest Research 32:1441–1450.}
+\nocite{fyllas_2009_basin}
 
-\hypertarget{citeproc_bib_item_19}{Bond-Lamberty, B., C.
-Wang, and S.
-T.
-Gower.
-2004.
-Net Primary Production and Net Ecosystem Production of a Boreal Black Spruce Wildfire Chronosequence.
-Global Change Biology 10:473–487.}
+\nocite{blonder_2012_leaf}
 
-\hypertarget{citeproc_bib_item_20}{Brown, K.
-A., D.
-F.
-B.
-Flynn, N.
-K.
-Abram, J.
-C.
-Ingram, S.
-E.
-Johnson, and P.
-Wright.
-2011.
-Assessing Natural Resource Use By Forest-Reliant Communities in Madagascar Using Functional Diversity and Functional Redundancy Metrics.
-PLoS ONE 6:e24107.}
+\nocite{cornelissen_2003_functional}
 
-\hypertarget{citeproc_bib_item_21}{Burrascano, S., R.
-Copiz, E.
-Del Vico, S.
-Fagiani, E.
-Giarrizzo, M.
-Mei, A.
-Mortelliti, F.
-M.
-Sabatini, and C.
-Blasi.
-2015.
-Wild Boar Rooting Intensity Determines Shifts in Understorey Composition and Functional Traits.
-Community Ecology 16:244–253.}
+\nocite{cavender-bares_2006_phylogenetic}
 
-\hypertarget{citeproc_bib_item_22}{Butterfield, B. J., and J. M. Briggs. 2010. Regeneration Niche Differentiates Functional Strategies of Desert Woody Plant Species. Oecologia 165:477–487.}
+\nocite{garnier_2007_assessing}
 
-\hypertarget{citeproc_bib_item_23}{Byun, C., S.
-de Blois, and J.
-Brisson.
-2012.
-Plant Functional Group Identity and Diversity Determine Biotic Resistance To Invasion By an Exotic Grass.
-Journal of Ecology 101:128–139.}
+\nocite{prentice_2010_evidence}
 
-\hypertarget{citeproc_bib_item_24}{Campbell, C., L.
-Atkinson, J.
-Zaragoza-Castells, M.
-Lundmark, O.
-Atkin, and V.
-Hurry.
-2007.
-Acclimation of Photosynthesis and Respiration Is Asynchronous in Response To Changes in Temperature Regardless of Plant Functional Group.
-New Phytologist 176:375–389.}
+\nocite{brown_2011_assessing}
 
-\hypertarget{citeproc_bib_item_25}{Campetella, G., Z.
-Botta-Dukát, C.
-Wellstein, R.
-Canullo, S.
-Gatto, S.
-Chelli, L.
-Mucina, and S.
-Bartha.
-2011.
-Patterns of Plant Trait-Environment Relationships Along a Forest Succession Chronosequence.
-Agriculture, Ecosystems \& Environment 145:38–48.}
+\nocite{schweingruber_2005}
 
-\hypertarget{citeproc_bib_item_26}{Carswell, F.
-E., P.
-Meir, E.
-V.
-Wandelli, L.
-C.
-M.
-Bonates, B.
-Kruijt, E.
-M.
-Barbosa, A.
-D.
-Nobre, J.
-Grace, and P.
-G.
-Jarvis.
-2000.
-Photosynthetic Capacity in a Central Amazonian Rain Forest.
-Tree Physiology 20:179–186.}
+\nocite{onoda_2011_global}
 
-\hypertarget{citeproc_bib_item_27}{Cavender-Bares, J., A. Keen, and B. Miles. 2006. Phylogenetic Structure of Floridian Plant Communities Depends on Taxonomic and Spatial Scale. Ecology 87:S109–S122.}
+\nocite{han_2005_leaf}
 
-\hypertarget{citeproc_bib_item_28}{Cerabolini, B.
-E.
-L., G.
-Brusa, R.
-M.
-Ceriani, R.
-De Andreis, A.
-Luzzaro, and S.
-Pierce.
-2010.
-Can CSR Classification Be Generally Applied Outside Britain?
-Plant Ecology 210:253–261.}
+\nocite{domingues_2010_co}
 
-\hypertarget{citeproc_bib_item_29}{Chambers, J.
-Q., E.
-S.
-Tribuzy, L.
-C.
-Toledo, B.
-F.
-Crispim, N.
-Higuchi, J.
-d.
-Santos, A.
-C.
-Araújo, B.
-Kruijt, A.
-D.
-Nobre, and S.
-E.
-Trumbore.
-2004.
-Respiration From a Tropical Forest Ecosystem: Partitioning of Sources and Low Carbon Use Efficiency.
-Ecological Applications 14:72–88.}
+\nocite{han_2005_leaf}
 
-\hypertarget{citeproc_bib_item_30}{Chen, Y., W.
-Han, L.
-Tang, Z.
-Tang, and J.
-Fang.
-2011.
-Leaf Nitrogen and Phosphorus Concentrations of Woody Plants Differ in Responses To Climate, Soil and Plant Growth Form.
-Ecography 36:178–184.}
+\nocite{messier_2010_how}
 
-\hypertarget{citeproc_bib_item_31}{Choat, B., S.
-Jansen, T.
-J.
-Brodribb, H.
-Cochard, S.
-Delzon, R.
-Bhaskar, S.
-J.
-Bucci, T.
-S.
-Feild, S.
-M.
-Gleason, U.
-G.
-Hacke, and et al.
-2012.
-Global Convergence in the Vulnerability of Forests To Drought.
-Nature 491:752–755.}
+\nocite{price_2007_scaling}
 
-\hypertarget{citeproc_bib_item_32}{Cornelissen, J.
-H.
-C.
-1996.
-An Experimental Comparison of Leaf Decomposition Rates in a Wide Range of Temperate Plant Species and Types.
-The Journal of Ecology 84:573.}
+\nocite{kraft_2008_functional}
 
-\hypertarget{citeproc_bib_item_33}{Cornelissen, J.
-H.
-C., P.
-C.
-Diez, and R.
-Hunt.
-1996.
-Seedling Growth, Allocation and Leaf Attributes in a Wide Range of Woody Plant Species and Types.
-The Journal of Ecology 84:755.}
+\nocite{pierce_2013_allocating}
 
-\hypertarget{citeproc_bib_item_34}{Cornelissen, J.
-H.
-C., H.
-M.
-Quested, D.
-Gwynn-jones, R.
-S.
-P.
-van Logtestijn, M.
-A.
-H.
-De Beus, A.
-Kondratchuk, T.
-V.
-Callaghan, and R.
-Aerts.
-2004.
-Leaf Digestibility and Litter Decomposability Are Related in a Wide Range of Subarctic Plant Species and Types.
-Functional Ecology 18:779–786.}
+\nocite{wright_2006_relationships}
 
-\hypertarget{citeproc_bib_item_35}{Cornelissen, J., B.
-Cerabolini, P.
-Castro-Díez, P.
-Villar-Salvador, G.
-Montserrat-Martí, J.
-Puyravaud, M.
-Maestro, M.
-Werger, and R.
-Aerts.
-2003.
-Functional Traits of Woody Plants: Correspondence of Species Rankings Between Field Adults and Laboratory-Grown Seedlings?
-Journal of Vegetation Science 14:311–322.}
+\nocite{reich_2008_scaling}
 
-\hypertarget{citeproc_bib_item_36}{Cornwell, W.
-K., J.
-H.
-C.
-Cornelissen, K.
-Amatangelo, E.
-Dorrepaal, V.
-T.
-Eviner, O.
-Godoy, S.
-E.
-Hobbie, B.
-Hoorens, H.
-Kurokawa, N.
-Pérez-Harguindeguy, and et al.
-2008.
-Plant Species Traits Are the Predominant Control on Litter Decomposition Rates Within Biomes Worldwide.
-Ecology Letters 11:1065–1071.}
+\nocite{domingues_2013_seasonal}
 
-\hypertarget{citeproc_bib_item_37}{Craine, J.
-M., A.
-J.
-Elmore, M.
-P.
-M.
-Aidar, M.
-Bustamante, T.
-E.
-Dawson, E.
-A.
-Hobbie, A.
-Kahmen, M.
-C.
-Mack, K.
-K.
-McLauchlan, A.
-Michelsen, and et al.
-2009.
-Global Patterns of Foliar Nitrogen Isotopes and Their Relationships With Climate, Mycorrhizal Fungi, Foliar Nutrient Concentrations, and Nitrogen Availability.
-New Phytologist 183:980–992.}
+\nocite{niinemets_2001_global}
 
-\hypertarget{citeproc_bib_item_38}{Craine, J.
-M., W.
-G.
-Lee, W.
-J.
-Bond, R.
-J.
-Williams, and L.
-C.
-Johnson.
-2005.
-Environmental Constraints on a Global Relationship Among Leaf and Root Traits of Grasses.
-Ecology 86:12–19.}
+\nocite{cornwell_2008_plant}
 
-\hypertarget{citeproc_bib_item_39}{Craine, J.
-M., J.
-B.
-Nippert, E.
-G.
-Towne, S.
-Tucker, S.
-W.
-Kembel, A.
-Skibbe, and K.
-K.
-McLauchlan.
-2011.
-Functional Consequences of Climate Change-Induced Plant Species Loss in a Tallgrass Prairie.
-Oecologia 165:1109–1117.}
+\nocite{vergutz_2012_global}
 
-\hypertarget{citeproc_bib_item_40}{Craine, J.
-M., E.
-G.
-Towne, T.
-W.
-Ocheltree, and J.
-B.
-Nippert.
-2012.
-Community Traitscape of Foliar Nitrogen Isotopes Reveals N Availability Patterns in a Tallgrass Prairie.
-Plant and Soil 356:395–403.}
+\nocite{sandel_2011_using}
 
-\hypertarget{citeproc_bib_item_41}{Craven, D., D.
-Braden, M.
-Ashton, G.
-Berlyn, M.
-Wishnie, and D.
-Dent.
-2007.
-Between and Within-Site Comparisons of Structural and Physiological Characteristics and Foliar Nutrient Content of 14 Tree Species At a Wet, Fertile Site and a Dry, Infertile Site in Panama.
-Forest Ecology and Management 238:335–346.}
+\nocite{poorter_2009_causes}
 
-\hypertarget{citeproc_bib_item_42}{Demey, A., J.
-Staelens, L.
-Baeten, P.
-Boeckx, M.
-Hermy, J.
-Kattge, and K.
-Verheyen.
-2013.
-Nutrient Input From Hemiparasitic Litter Favors Plant Species With a Fast-Growth Strategy.
-Plant and Soil 371:53–66.}
+\nocite{gallagher_2012_global}
 
-\hypertarget{citeproc_bib_item_43}{Diaz, S., J.
-Hodgson, K.
-Thompson, M.
-Cabido, J.
-Cornelissen, A.
-Jalili, G.
-Montserrat-Martí, J.
-Grime, F.
-Zarrinkamar, Y.
-Asri, and et al.
-2004.
-The Plant Traits That Drive Ecosystems: Evidence From Three Continents.
-Journal of Vegetation Science 15:295–304.}
+\nocite{williams_2012_lba}
 
-\hypertarget{citeproc_bib_item_44}{Domingues, T.
-F., L.
-A.
-Martinelli, and J.
-R.
-Ehleringer.
-2013.
-Seasonal Patterns of Leaf-Level Photosynthetic Gas Exchange in an Eastern Amazonian Rain Forest.
-Plant Ecology \& Diversity 7:189–203.}
+\nocite{choat_2012_global}
 
-\hypertarget{citeproc_bib_item_45}{Domingues, T.
-F., P.
-Meir, T.
-R.
-Feldpausch, G.
-Saiz, E.
-M.
-Veenendaal, F.
-Schrodt, M.
-Bird, G.
-Djagbletey, F.
-Hien, H.
-Compaore, and et al.
-2010.
-Co-Limitation of Photosynthetic Capacity By Nitrogen and Phosphorus in West Africa Woodlands.
-Plant, Cell \& Environment 33:959–980.}
+\nocite{fonseca_2000_shifts}
 
-\hypertarget{citeproc_bib_item_46}{Fitter, A. H., and H. J. Peat. 1994. The Ecological Flora Database. The Journal of Ecology 82:415.}
+\nocite{burrascano_2015_wild}
 
-\hypertarget{citeproc_bib_item_47}{Fonseca, C.
-R., J.
-M.
-Overton, B.
-Collins, and M.
-Westoby.
-2000.
-Shifts in Trait-Combinations Along Rainfall and Phosphorus Gradients.
-Journal of Ecology 88:964–977.}
+\nocite{milla_2011_multi}
 
-\hypertarget{citeproc_bib_item_48}{Frenette-Dussault, C., B.
-Shipley, J.-F.
-Léger, D.
-Meziane, and Y.
-Hingrat.
-2011.
-Functional Structure of an Arid Steppe Plant Community Reveals Similarities With Grime’s C-S-R Theory.
-Journal of Vegetation Science 23:208–222.}
+\nocite{adriaenssens_2012_dry}
 
-\hypertarget{citeproc_bib_item_49}{Fyllas, N.
-M., S.
-Patiño, T.
-R.
-Baker, G.
-Bielefeld Nardoto, L.
-A.
-Martinelli, C.
-A.
-Quesada, R.
-Paiva, M.
-Schwarz, V.
-Horna, L.
-M.
-Mercado, and et al.
-2009.
-Basin-Wide Variations in Foliar Properties of Amazonian Forest: Phylogeny, Soils and Climate.
-Biogeosciences 6:2677–2708.}
+\nocite{willis_2009_phylogenetic}
 
-\hypertarget{citeproc_bib_item_50}{Gallagher, R. V., and M. R. Leishman. 2012. A Global Analysis of Trait Variation and Evolution in Climbing Plants. Journal of Biogeography 39:1757–1771.}
+\nocite{kisel_2012_testing}
 
-\hypertarget{citeproc_bib_item_51}{Garnier, E., S.
-Lavorel, P.
-Ansquer, H.
-Castro, P.
-Cruz, J.
-Dolezal, O.
-Eriksson, C.
-Fortunel, H.
-Freitas, C.
-Golodets, and et al.
-2007.
-Assessing the Effects of Land-Use Change on Plant Traits, Communities and Ecosystem Functioning in Grasslands: a Standardized Methodology and Lessons From an Application To 11 European Sites.
-Annals of Botany 99:967–985.}
+\nocite{pierce_2012_combined}
 
-\hypertarget{citeproc_bib_item_52}{Givnish, T.
-J., R.
-A.
-Montgomery, and G.
-Goldstein.
-2004.
-Adaptive Radiation of Photosynthetic Physiology in the Hawaiian Lobeliads: Light Regimes, Static Light Responses, and Whole-Plant Compensation Points.
-American Journal of Botany 91:228–246.}
+\nocite{cornelissen_1996_experimental}
 
-\hypertarget{citeproc_bib_item_53}{Guerin, G. R., H. Wen, and A. J. Lowe. 2012. Leaf Morphology Shift Linked To Climate Change. Biology Letters 8:882–886.}
+\nocite{yguel_2011_phytophagy}
 
-\hypertarget{citeproc_bib_item_54}{Gutiérrez, A.
-G., and A.
-Huth.
-2012.
-Successional Stages of Primary Temperate Rainforests of Chiloé Island, Chile.
-Perspectives in Plant Ecology, Evolution and Systematics 14:243–256.}
+\nocite{xu_2003_seasonal}
 
-\hypertarget{citeproc_bib_item_55}{Guy, A.
-L., J.
-M.
-Mischkolz, and E.
-G.
-Lamb.
-2013.
-Limited Effects of Simulated Acidic Deposition on Seedling Survivorship and Root Morphology of Endemic Plant Taxa of the Athabasca Sand Dunes in Well-Watered Greenhouse Trials.
-Botany 91:176–181.}
+\nocite{bond-lamberty_2004_net}
 
-\hypertarget{citeproc_bib_item_56}{Han, W., J. Fang, D. Guo, and Y. Zhang. 2005. Leaf Nitrogen and Phosphorus Stoichiometry Across 753 Terrestrial Plant Species in China. New Phytologist 168:377–385.}
+\nocite{von_2004_testing}
 
-\hypertarget{citeproc_bib_item_57}{Hickler, T. 1999. Plant Functional Types and community characteristics along environmental gradients on Öland’s Great Alvar. University of Lund, Lund, Sweden.}
+\nocite{chen_2011_leaf}
 
-\hypertarget{citeproc_bib_item_58}{Kattge, J., W.
-Knorr, T.
-Raddatz, and C.
-Wirth.
-2009.
-Quantifying Photosynthetic Capacity and Its Relationship To Leaf Nitrogen Content for Global-Scale Terrestrial Biosphere Models.
-Global Change Biology 15:976–991.}
+\nocite{byun_2012_plant}
 
-\hypertarget{citeproc_bib_item_59}{Kazakou, E., D.
-Vile, B.
-Shipley, C.
-Gallet, and E.
-Garnier.
-2006.
-Co-Variations in Litter Decomposition, Leaf Traits and Plant Growth in Species From a Mediterranean Old-Field Succession.
-Functional Ecology 20:21–30.}
+\nocite{laughlin_2011_climatic}
 
-\hypertarget{citeproc_bib_item_60}{Kerkhoff, A.
-J., W.
-F.
-Fagan, J.
-J.
-Elser, and B.
-J.
-Enquist.
-2006.
-Phylogenetic and Growth Form Variation in the Scaling of Nitrogen and Phosphorus in the Seed Plants.
-The American Naturalist 168:E103–E122.}
+\nocite{ordonez_2010_plant}
 
-\hypertarget{citeproc_bib_item_61}{Kichenin, E., D.
-A.
-Wardle, D.
-A.
-Peltzer, C.
-W.
-Morse, and G.
-T.
-Freschet.
-2013.
-Contrasting Effects of Plant Inter- and Intraspecific Variation on Community-Level Trait Measures Along an Environmental Gradient.
-Functional Ecology 27:1254–1261.}
+\nocite{kerkhoff_2006_phylogenetic}
 
-\hypertarget{citeproc_bib_item_62}{Kisel, Y., A.
-C.
-Moreno-Letelier, D.
-Bogarín, M.
-P.
-Powell, M.
-W.
-Chase, and T.
-G.
-Barraclough.
-2012.
-Testing the Link Between Population Genetic Differentiation and Clade Diversification in Costa Rican Orchids.
-Evolution 66:3035–3052.}
+\nocite{butterfield_2010_regeneration}
 
-\hypertarget{citeproc_bib_item_63}{Kleyer, M., R.
-Bekker, I.
-Knevel, J.
-Bakker, K.
-Thompson, M.
-Sonnenschein, P.
-Poschlod, J.
-van Groenendael, L.
-Klimevs, J.
-Klimevsová, and et al.
-2008.
-The LEDA Traitbase: a Database of Life-History Traits of the Northwest European Flora.
-Journal of Ecology 96:1266–1274.}
+\nocite{bond-lamberty_2002_aboveground}
 
-\hypertarget{citeproc_bib_item_64}{Kraft, N. J. B., R. Valencia, and D. D. Ackerly. 2008. Functional Traits and Niche-Based Tree Community Assembly in an Amazonian Forest. Science 322:580–582.}
+\nocite{pierce_2007_disturbance}
 
-\hypertarget{citeproc_bib_item_65}{Laughlin, D. C., P. Z. Fulé, D. W. Huffman, J. Crouse, and E. Laliberté. 2011. Climatic Constraints on Trait-Based Forest Assembly. Journal of Ecology 99:1489–1499.}
+\nocite{gutierrez_2012_successional}
 
-\hypertarget{citeproc_bib_item_66}{Loranger, J., S.
-T.
-Meyer, B.
-Shipley, J.
-Kattge, H.
-Loranger, C.
-Roscher, and W.
-W.
-Weisser.
-2012.
-Predicting Invertebrate Herbivory From Plant Traits: Evidence From 51 Grassland Species in Experimental Monocultures.
-Ecology 93:2674–2682.}
+\nocite{manzoni_2013_optimization}
 
-\hypertarget{citeproc_bib_item_67}{Louault, F., V.
-Pillar, J.
-Aufrère, E.
-Garnier, and J.-F.
-Soussana.
-2005.
-Plant Traits and Functional Types in Response To Reduced Disturbance in a Semi-Natural Grassland.
-Journal of Vegetation Science 16:151–160.}
+\nocite{wright_2012_does}
 
-\hypertarget{citeproc_bib_item_68}{Loveys, B.
-R., L.
-J.
-Atkinson, D.
-J.
-Sherlock, R.
-L.
-Roberts, A.
-H.
-Fitter, and O.
-K.
-Atkin.
-2003.
-Thermal Acclimation of Leaf and Root Respiration: an Investigation Comparing Inherently Fast- and Slow-Growing Plant Species.
-Global Change Biology 9:895–910.}
+\nocite{chambers_2004_respiration}
 
-\hypertarget{citeproc_bib_item_69}{Manzoni, S., G.
-Vico, S.
-Palmroth, A.
-Porporato, and G.
-Katul.
-2013.
-Optimization of Stomatal Conductance for Maximum Carbon Gain Under Dynamic Soil Moisture.
-Advances in Water Resources 62:90–105.}
+\nocite{guerin_2012_leaf}
 
-\hypertarget{citeproc_bib_item_70}{Medlyn, B.
-E., F.
-.-W.
-Badeck, D.
-G.
-G.
-De Pury, C.
-V.
-M.
-Barton, M.
-Broadmeadow, R.
-Ceulemans, P.
-De Angelis, M.
-Forstreuter, M.
-E.
-Jach, S.
-Kellomäki, and et al.
-1999.
-Effects of Elevated [CO2] on Photosynthesis in European Forest Species: a Meta-Analysis of Model Parameters.
-Plant, Cell \& Environment 22:1475–1495.}
+\nocite{bond-lamberty_2002_leaf}
 
-\hypertarget{citeproc_bib_item_71}{Meir, P., B.
-Kruijt, M.
-Broadmeadow, E.
-Barbosa, O.
-Kull, F.
-Carswell, A.
-Nobre, and P.
-G.
-Jarvis.
-2002.
-Acclimation of Photosynthetic Capacity To Irradiance in Tree Canopies in Relation To Leaf Nitrogen Concentration and Leaf Mass Per Unit Area.
-Plant, Cell and Environment 25:343–357.}
+\nocite{mueller_2006_plant}
 
-\hypertarget{citeproc_bib_item_72}{Meir, P., P.
-E.
-Levy, J.
-Grace, and P.
-G.
-Jarvis.
-2007.
-Photosynthetic Parameters From Two Contrasting Woody Vegetation Types in West Africa.
-Plant Ecology 192:277–287.}
+\nocite{cerabolini_2010_can}
 
-\hypertarget{citeproc_bib_item_73}{Messier, J., B. J. McGill, and M. J. Lechowicz. 2010. How Do Traits Vary Across Ecological Scales? a Case for Trait-Based Ecology. Ecology Letters 13:838–848.}
+\nocite{craine_2011_functional}
 
-\hypertarget{citeproc_bib_item_74}{Meziane, D., and B.
-Shipley.
-1999.
-Interacting Determinants of Specific Leaf Area in 22 Herbaceous Species: Effects of Irradiance and Nutrient Availability.
-Plant, Cell \& Environment 22:447–459.}
+\nocite{craine_2012_community}
 
-\hypertarget{citeproc_bib_item_75}{Milla, R., and P. B. Reich. 2011. Multi-Trait Interactions, Not Phylogeny, Fine-Tune Leaf Size Reduction With Increasing Altitude. Annals of Botany 107:455–465.}
+\nocite{shipley_2002_trade}
 
-\hypertarget{citeproc_bib_item_76}{Minden, V., S.
-Andratschke, J.
-Spalke, H.
-Timmermann, and M.
-Kleyer.
-2012.
-Plant Trait-Environment Relationships in Salt Marshes: Deviations From Predictions By Ecological Concepts.
-Perspectives in Plant Ecology, Evolution and Systematics 14:183–192.}
+\nocite{wohlfahrt_1999_inter}
 
-\hypertarget{citeproc_bib_item_77}{Minden, V., and M.
-Kleyer.
-2011.
-Testing the Effect-Response Framework: Key Response and Effect Traits Determining Above-Ground Biomass of Salt Marshes.
-Journal of Vegetation Science 22:387–401.}
+\nocite{rueger_2011_determinants}
 
-\hypertarget{citeproc_bib_item_78}{Müller, S.
-C., G.
-E.
-Overbeck, J.
-Pfadenhauer, and V.
-D.
-Pillar.
-2006.
-Plant Functional Types of Woody Species Related To Fire Disturbance in Forest-Grassland Ecotones.
-Plant Ecology 189:1–14.}
+\nocite{rueger_2009_response}
 
-\hypertarget{citeproc_bib_item_79}{Niinemets, Ü. 2001. Global-Scale Climatic Controls of Leaf Dry Mass Per Area, Density, and Thickness in Trees and Shrubs. Ecology 82:453–469.}
+\nocite{quested_2003_decomposition}
 
-\hypertarget{citeproc_bib_item_80}{Ogaya, R., and J.
-Peñuelas.
-2003.
-Comparative Field Study of Quercus Ilex and Phillyrea Latifolia: Photosynthetic Response To Experimental Drought Conditions.
-Environmental and Experimental Botany 50:137–148.}
+\nocite{choat_2012_global}
 
-\hypertarget{citeproc_bib_item_81}{Onoda, Y., M.
-Westoby, P.
-B.
-Adler, A.
-M.
-F.
-Choong, F.
-J.
-Clissold, J.
-H.
-C.
-Cornelissen, S.
-Díaz, N.
-J.
-Dominy, A.
-Elgart, L.
-Enrico, and et al.
-2011.
-Global Patterns of Leaf Mechanical Properties.
-Ecology Letters 14:301–312.}
+\nocite{pahl_2013_no}
 
-\hypertarget{citeproc_bib_item_82}{Ordoñez, J.
-C., P.
-M.
-van Bodegom, Witte Jan‐Philip M., R.
-P.
-Bartholomeus, J.
-R.
-van Hal, and R.
-Aerts.
-2010.
-Plant Strategies in Relation To Resource Supply in Mesic To Wet Environments: Does Theory Mirror Nature?
-The American Naturalist 175:225–239.}
+\nocite{shipley_2002_dry}
 
-\hypertarget{citeproc_bib_item_83}{Pahl, A.
-T., J.
-Kollmann, A.
-Mayer, and S.
-Haider.
-2013.
-No Evidence for Local Adaptation in an Invasive Alien Plant: Field and Greenhouse Experiments Tracing a Colonization Sequence.
-Annals of Botany 112:1921–1930.}
+\nocite{minden_2011_testing}
 
-\hypertarget{citeproc_bib_item_84}{Peco, B., I.
-de Pablos, J.
-Traba, and C.
-Levassor.
-2005.
-The Effect of Grazing Abandonment on Species Composition and Functional Traits: the Case of Dehesa Grasslands.
-Basic and Applied Ecology 6:175–183.}
+\nocite{meir_2002_acclimation}
 
-\hypertarget{citeproc_bib_item_85}{Penuelas, J., J.
-Sardans, J.
-Llusià, S.
-M.
-Owen, J.
-Carnicer, T.
-W.
-Giambelluca, E.
-L.
-Rezende, M.
-Waite, and Ü.
-Niinemets.
-2009.
-Faster Returns on “Leaf Economics” and Different Biogeochemical Niche in Invasive Compared With Native Plant Species.
-Global Change Biology 16:2171–2185.}
+\nocite{schererlorenzen_2007_exploring}
 
-\hypertarget{citeproc_bib_item_86}{Pierce, S., G.
-Brusa, M.
-Sartori, and B.
-E.
-L.
-Cerabolini.
-2012.
-Combined Use of Leaf Size and Economics Traits Allows Direct Comparison of Hydrophyte and Terrestrial Herbaceous Adaptive Strategies.
-Annals of Botany 109:1047–1053.}
+\nocite{bakker_2006_plant}
 
-\hypertarget{citeproc_bib_item_87}{Pierce, S., G.
-Brusa, I.
-Vagge, and B.
-E.
-L.
-Cerabolini.
-2013.
-Allocating Csr Plant Functional Types: the Use of Leaf Economics and Size Traits To Classify Woody and Herbaceous Vascular Plants.
-Functional Ecology 27:1002–1010.}
+\nocite{penuelas_2009_faster}
 
-\hypertarget{citeproc_bib_item_88}{Pierce, S., A.
-Luzzaro, M.
-Caccianiga, R.
-M.
-Ceriani, and B.
-Cerabolini.
-2007.
-Disturbance Is the Principal $alpha$-scale Filter Determining Niche Differentiation, Coexistence and Biodiversity in an Alpine Community.
-Journal of Ecology 95:698–706.}
+\nocite{powers_2010_plant}
 
-\hypertarget{citeproc_bib_item_89}{Pillar, V. D., and E. E. Sosinski. 2003. An Improved Method for Searching Plant Functional Types By Numerical Analysis. Journal of Vegetation Science 14:323–332.}
+\nocite{shiodera_2008_variation}
 
-\hypertarget{citeproc_bib_item_90}{Poorter, H., Ü.
-Niinemets, L.
-Poorter, I.
-J.
-Wright, and R.
-Villar.
-2009.
-Causes and Consequences of Variation in Leaf Mass Per Area (LMA): a Meta-Analysis.
-New Phytologist 182:565–588.}
+\nocite{fitter_1994_ecological_flora_database}
 
-\hypertarget{citeproc_bib_item_91}{Powers, J.
-S., and P.
-Tiffin.
-2010.
-Plant Functional Type Classifications in Tropical Dry Forests in Costa Rica: Leaf Habit Versus Taxonomic Approaches.
-Functional Ecology 24:927–936.}
+\nocite{minden_2012_plant}
 
-\hypertarget{citeproc_bib_item_92}{Prentice, I.
-C., T.
-Meng, H.
-Wang, S.
-P.
-Harrison, J.
-Ni, and G.
-Wang.
-2010.
-Evidence of a Universal Scaling Relationship for Leaf CO2 Drawdown Along an Aridity Gradient.
-New Phytologist 190:169–180.}
+\nocite{bond-lamberty_2006_nitrogen}
 
-\hypertarget{citeproc_bib_item_93}{Preston, K.
-A., W.
-K.
-Cornwell, and J.
-L.
-DeNoyer.
-2006.
-Wood Density and Vessel Traits As Distinct Correlates of Ecological Strategy in 51 California Coast Range Angiosperms.
-New Phytologist 170:807–818.}
+\nocite{rueger_2011_determinants}
 
-\hypertarget{citeproc_bib_item_94}{Price, C. A., and B. J. Enquist. 2007. Scaling Mass and Morphology in Leaves: an Extension of the Wbe Model. Ecology 88:1132–1141.}
+\nocite{peco_2005_effect}
 
-\hypertarget{citeproc_bib_item_95}{Pyankov, V.
-I., A.
-V.
-Kondratchuk, and B.
-Shipley.
-1999.
-Leaf Structure and Specific Leaf Mass: the Alpine Desert Plants of the Eastern Pamirs, Tadjikistan.
-New Phytologist 143:131–142.}
+\nocite{tucker_2011_physiological}
 
-\hypertarget{citeproc_bib_item_96}{Quested, H.
-M., J.
-H.
-C.
-Cornelissen, M.
-C.
-Press, T.
-V.
-Callaghan, R.
-Aerts, F.
-Trosien, P.
-Riemann, D.
-Gwynn-Jones, A.
-Kondratchuk, and S.
-E.
-Jonasson.
-2003.
-Decomposition of Sub-Arctic Plants With Differing Nitrogen Economies: a Functional Role for Hemiparasites.
-Ecology 84:3209–3221.}
+\nocite{weedon_2009_global}
 
-\hypertarget{citeproc_bib_item_97}{Reich, P.
-B., M.
-G.
-Tjoelker, K.
-S.
-Pregitzer, I.
-J.
-Wright, J.
-Oleksyn, and J.-L.
-Machado.
-2008.
-Scaling of Respiration To Nitrogen in Leaves, Stems and Roots of Higher Land Plants.
-Ecology Letters 11:793–801.}
+\nocite{blonder_2010_venation}
 
-\hypertarget{citeproc_bib_item_98}{Rüger, N., A.
-Huth, S.
-P.
-Hubbell, and R.
-Condit.
-2009.
-Response of Recruitment To Light Availability Across a Tropical Lowland Rain Forest Community.
-Journal of Ecology 97:1360–1368.}
+\nocite{spasojevic_2012_inferring}
 
-\hypertarget{citeproc_bib_item_99}{Rüger, N., A. Huth, S. P. Hubbell, and R. Condit. 2011. Determinants of Mortality Across a Tropical Lowland Rainforest Community. Oikos 120:1047–1056.}
+\nocite{campbell_2007_acclimation}
 
-\hypertarget{citeproc_bib_item_100}{Sandel, B., J. D. Corbin, and M. Krupa. 2011. Using Plant Functional Traits To Guide Restoration: a Case Study in California Coastal Grassland. Ecosphere 2:art23.}
+\nocite{craven_2007_between}
 
-\hypertarget{citeproc_bib_item_101}{Schererlorenzen, M., E.
-Schulze, A.
-Don, J.
-Schumacher, and E.
-Weller.
-2007.
-Exploring the Functional Significance of Forest Diversity: a New Long-Term Experiment With Temperate Tree Species (BIOTREE).
-Perspectives in Plant Ecology, Evolution and Systematics 9:53–70.}
+\nocite{cornelissen_2004_leaf}
 
-\hypertarget{citeproc_bib_item_102}{Schweingruber, F., and W. Landolt. 2005. The xylem database.}
+\nocite{pillar_2003_improved}
 
-\hypertarget{citeproc_bib_item_103}{Shiodera, S., J.
-S.
-Rahajoe, and T.
-Kohyama.
-2008.
-Variation in Longevity and Traits of Leaves Among Co-Occurring Understorey Plants in a Tropical Montane Forest.
-Journal of Tropical Ecology 24:121–133.}
+\nocite{hickler_1999_plant}
 
-\hypertarget{citeproc_bib_item_104}{Shipley, B. 1995. Structured Interspecific Determinants of Specific Leaf Area in 34 Species of Herbaceous Angiosperms. Functional Ecology 9:312.}
+\nocite{bakker_2005_effects}
 
-\hypertarget{citeproc_bib_item_105}{Shipley, B.
-2002.
-Trade-Offs Between Net Assimilation Rate and Specific Leaf Area in Determining Relative Growth Rate: Relationship With Daily Irradiance.
-Functional Ecology 16:682–689.}
+\nocite{vile_2005_significations}
 
-\hypertarget{citeproc_bib_item_106}{Shipley, B., and M.
-J.
-Lechowicz.
-2000.
-The Functional Co-Ordination of Leaf Morphology, Nitrogen Concentration, and Gas Exchange In40 Wetland Species.
-Écoscience 7:183–194.}
+\nocite{craine_2005_environmental}
 
-\hypertarget{citeproc_bib_item_107}{Shipley, B., and T.-T. Vu. 2002. Dry Matter Content As a Measure of Dry Matter Concentration in Plants and Their Parts. New Phytologist 153:359–364.}
+\nocite{swaine_2007_ecological_bornean}
 
-\hypertarget{citeproc_bib_item_108}{Spasojevic, M.
-J., and K.
-N.
-Suding.
-2012.
-Inferring Community Assembly Mechanisms From Functional Diversity Patterns: the Importance of Multiple Assembly Processes.
-Journal of Ecology 100:652–661.}
+\nocite{wilson_2000_spatial}
 
-\hypertarget{citeproc_bib_item_109}{Swaine, E. 2007. Ecological and evolutionary drivers of plant community assembly in a Bornean rain forest. University of Aberdeen, Aberdeen, Scotland, UK.}
+\nocite{preston_2006_wood}
 
-\hypertarget{citeproc_bib_item_110}{Tucker, S. S., J. M. Craine, and J. B. Nippert. 2011. Physiological Drought Tolerance and the Structuring of Tallgrass Prairie Assemblages. Ecosphere 2:art48.}
+\nocite{zapata-cuartas_2012_probability}
 
-\hypertarget{citeproc_bib_item_111}{Veihmeyer, F. J. 1956. Soil Moisture. Pflanze und Wasser / Water Relations of Plants:64–123.}
+\nocite{blonder_2013_assessing}
 
-\hypertarget{citeproc_bib_item_112}{Vergutz, L., S. Manzoni, A. Porporato, R. Novais, and R. Jackson. 2012. A Global Database of Carbon and Nutrient Concentrations of Green and Senesced Leaves.}
+\nocite{blonder_2010_venation}
 
-\hypertarget{citeproc_bib_item_113}{Vile, D.
-2005.
-Significations fonctionnelle et ecologique des traits des especes vegetales: Exemple dans une succession post-cultural Mediterraneenne et generalisations.}
+\nocite{guy_2013_limited}
 
-\hypertarget{citeproc_bib_item_114}{Von Holle, B., and D. Simberloff. 2004. Testing Fox’s Assembly Rule: Does Plant Invasion Depend on Recipient Community Structure? Oikos 105:551–563.}
+\nocite{kichenin_2013_contrasting}
 
-\hypertarget{citeproc_bib_item_115}{Weedon, J.
-T., W.
-K.
-Cornwell, J.
-H.
-Cornelissen, A.
-E.
-Zanne, C.
-Wirth, and D.
-A.
-Coomes.
-2009.
-Global Meta-Analysis of Wood Decomposition Rates: a Role for Trait Variation Among Tree Species?
-Ecology Letters 12:45–56.}
+\nocite{louault_2005_plant}
 
-\hypertarget{citeproc_bib_item_116}{Williams, M., Y. Shimabokuro, and E. Rastetter. 2012. LBA-ECO CD-09 Soil and Vegetation Characteristics, Tapajos National Forest, Brazil.}
+\nocite{araujo_2012_lba}
 
-\hypertarget{citeproc_bib_item_117}{Willis, C.
-G., M.
-Halina, C.
-Lehman, P.
-B.
-Reich, A.
-Keen, S.
-McCarthy, and J.
-Cavender-Bares.
-2009.
-Phylogenetic Community Structure in Minnesota Oak Savanna Is Influenced By Spatial Extent and Environmental Variation.
-Ecography:no-no.}
+\nocite{bocanegra_2015_grupos}
 
-\hypertarget{citeproc_bib_item_118}{Wilson, K.
-B., D.
-D.
-Baldocchi, and P.
-J.
-Hanson.
-2000.
-Spatial and Seasonal Variability of Photosynthetic Parameters and Their Relationship To Leaf Nitrogen in a Deciduous Forest.
-Tree Physiology 20:565–578.}
+\nocite{beckmann_2012_role}
 
-\hypertarget{citeproc_bib_item_119}{Wirth, C., and J.
-W.
-Lichstein.
-2009.
-The Imprint of Species Turnover on Old-Growth Forest Carbon Balances - Insights From a Trait-Based Model of Forest Dynamics.
-Ecological Studies:81–113.}
+\nocite{adler_2004_functional}
 
-\hypertarget{citeproc_bib_item_120}{Wohlfahrt, G., M.
-Bahn, E.
-Haubner, I.
-Horak, W.
-Michaeler, K.
-Rottmar, U.
-Tappeiner, and A.
-Cernusca.
-1999.
-Inter-Specific Variation of the Biochemical Limitation To Photosynthesis and Related Leaf Traits of 30 Species From Mountain Grassland Ecosystems Under Different Land Use.
-Plant, Cell and Environment 22:1281–1296.}
+\nocite{loveys_2003_thermal}
 
-\hypertarget{citeproc_bib_item_121}{Wright, I.
-J., D.
-D.
-Ackerly, F.
-Bongers, K.
-E.
-Harms, G.
-Ibarra-Manriquez, M.
-Martinez-Ramos, S.
-J.
-Mazer, H.
-C.
-Muller-Landau, H.
-Paz, N.
-C.
-A.
-Pitman, and et al.
-2006.
-Relationships Among Ecologically Important Dimensions of Plant Trait Variation in Seven Neotropical Forests.
-Annals of Botany 99:1003–1015.}
+\nocite{bodegom_2008_separating}
 
-\hypertarget{citeproc_bib_item_122}{Wright, I.
-J., P.
-B.
-Reich, M.
-Westoby, D.
-D.
-Ackerly, Z.
-Baruch, F.
-Bongers, J.
-Cavender-Bares, T.
-Chapin, J.
-H.
-C.
-Cornelissen, M.
-Diemer, and et al.
-2004.
-The Worldwide Leaf Economics Spectrum.
-Nature 428:821–827.}
+\nocite{domingues_2013_seasonal}
 
-\hypertarget{citeproc_bib_item_123}{Wright, J.
-P., and A.
-Sutton-Grier.
-2012.
-Does the Leaf Economic Spectrum Hold Within Local Species Pools Across Varying Environmental Conditions?
-Functional Ecology 26:1390–1398.}
+\nocite{shipley_2000_functional}
 
-\hypertarget{citeproc_bib_item_124}{Wright, S.
-J., K.
-Kitajima, N.
-J.
-B.
-Kraft, P.
-B.
-Reich, I.
-J.
-Wright, D.
-E.
-Bunker, R.
-Condit, J.
-W.
-Dalling, S.
-J.
-Davies, S.
-Díaz, and et al.
-2010.
-Functional Traits and the Growth-Mortality Trade-Off in Tropical Trees.
-Ecology 91:3664–3674.}
+\nocite{loranger_2012_predicting}
 
-\hypertarget{citeproc_bib_item_125}{Xu, L., and D.
-D.
-Baldocchi.
-2003.
-Seasonal Trends in Photosynthetic Parameters and Stomatal Conductance of Blue Oak (Quercus douglasii) Under Prolonged Summer Drought and High Temperature.
-Tree Physiology 23:865–877.}
+\nocite{pyankov_1999_leaf}
 
-\hypertarget{citeproc_bib_item_126}{Yguel, B., R.
-Bailey, N.
-D.
-Tosh, A.
-Vialatte, C.
-Vasseur, X.
-Vitrac, F.
-Jean, and A.
-Prinzing.
-2011.
-Phytophagy on Phylogenetically Isolated Trees: Why Hosts Should Escape Their Relatives.
-Ecology Letters 14:1117–1124.}
+\nocite{shipley_1995_structured}
 
-\hypertarget{citeproc_bib_item_127}{Zapata-Cuartas, M., C.
-A.
-Sierra, and L.
-Alleman.
-2012.
-Probability Distribution of Allometric Coefficients and Bayesian Estimation of Aboveground Tree Biomass.
-Forest Ecology and Management 277:173–179.}
-\end{hangparas}
+\nocite{cornelissen_1996_seedling}
+
+\nocite{carswell_2000_photosynthetic}
+
+\nocite{givnish_2004_adaptive}
+
+\nocite{meziane_1999_interacting}
+
+\nocite{meir_2007_photosynthetic}
+
+\nocite{atkin_1999_response}
+
+\nocite{diaz_2004_plant}
+
+\nocite{kazakou_2006_co}
+
+\nocite{demey_2013_nutrient}
+
+\printbibliography[heading=none]
 \end{document}

### 2019-09-09 AppendixS1.tex
Fix other appendix formatting/label issues

@@
-% Created 2019-09-09 Mon 09:26
+% Created 2019-09-09 Mon 09:38
 % Intended LaTeX compiler: pdflatex
@@
 \addbibresource{/Users/shik544/Dropbox/references/library.bib}
-\input{authors.tex}
 \date{}
-\title{Does the leaf economic spectrum hold within plant functional types? A Bayesian multivariate trait meta-analysis}
+\title{}
 \hypersetup{
  pdfauthor={},
- pdftitle={Does the leaf economic spectrum hold within plant functional types? A Bayesian multivariate trait meta-analysis},
+ pdftitle={},
  pdfkeywords={},
@@
 
-\maketitle
+\noindent \textbf{Title:}
+Does the leaf economic spectrum hold within plant functional types?
+A Bayesian multivariate trait meta-analysis
+
+\noindent \textbf{Authors:}
+Alexey N. Shiklomanov,
+Elizabeth M. Cowdery,
+Michael Bahn,
+Chaeho Byun,
+Steven Jansen,
+Koen Kramer,
+Vanessa Minden,
+Ülo Niinemets,
+Yusuke Onoda,
+Nadejda A. Soudzilovskaia,
+Michael C. Dietze
+
 \noindent \textbf{Journal:} Ecological Applications
 
-\clearpage
+\vspace{\baselineskip}
 

### 2019-09-09 AppendixS2.Rmd
Fix other appendix formatting/label issues

--- 2019-09-04 EAshiklomanovSA2.Rmd
@@
 ---
-title: Supporting information
-author:
-  - >
-    Alexey N. Shiklomanov,
-    Elizabeth M. Cowdery,
-    Michael Bahn,
-    Chaeho Byun,
-    Steven Jansen,
-    Koen Kramer,
-    Vanessa Minden,
-    Ülo Niinemets,
-    Yusuke Onoda,
-    Nadejda A. Soudzilovskaia,
-    Michael C. Dietze
 output: pdf_document
@@
 
+**Title:** 
+Does the leaf economic spectrum hold within plant functional types?
+A Bayesian multivariate trait meta-analysis
+
+**Authors:** 
+Alexey N. Shiklomanov,
+Elizabeth M. Cowdery,
+Michael Bahn,
+Chaeho Byun,
+Steven Jansen,
+Koen Kramer,
+Vanessa Minden,
+Ülo Niinemets,
+Yusuke Onoda,
+Nadejda A. Soudzilovskaia,
+Michael C. Dietze
+
+**Journal:** Ecological Applications
+
+**Appendix S2: Supporting information**
+
 ```{r libraries, include = FALSE}
@@
 ```
-
-(Page intentionally left blank)
 
@@
 
-# Method S1: Estimating trait means and covariances through multiple imputation {-}
+# Section S1: Estimating trait means and covariances through multiple imputation {-}
 
@@
 
-# Method S2: Biased mean estimates in non-representatively sampled bivariate data
+# Section S2: Biased mean estimates in non-representatively sampled bivariate data
 

### 2019-11-01 track-changes.sh
Old-new in track-changes

@@ -1,5 +1,5 @@
 #!/usr/bin/env bash
 
 rm -rf shiklomanov_ea_diff.{aux,log,out,tex,pdf}
-latexdiff shiklomanov_ea_old.tex shiklomanov_ea.tex > shiklomanov_ea_diff.tex
+latexdiff shiklomanov_ea_old.tex shiklomanov_ea_new.tex > shiklomanov_ea_diff.tex
 pdflatex shiklomanov_ea_diff.tex

### 2019-11-22 shiklomanov_formatted.tex
Add formatted LaTeX file

\documentclass{article}
\title{Does the leaf economic spectrum hold within plant functional types? A Bayesian multivariate trait meta-analysis}

\usepackage[left=1in,right=1in,top=1in,bottom=1in]{geometry}
\usepackage{lineno}

\usepackage{setspace}
\usepackage{fontspec}
\usepackage{graphicx}

\usepackage{booktabs}

\usepackage{caption}
\captionsetup{font={stretch=2}}

\usepackage{natbib}
\bibliographystyle{humannat}

\input{authors.tex}
\date{}

\begin{document}
\maketitle

\begin{footnotesize}
\noindent
\textbf{Submission type:} Articles

\noindent
\textbf{Running head:} Trait covariance within vs.\ across PFTs (39 characters)

\noindent
\textbf{Keywords:} Functional trade-off; hierarchical modeling; trait variation; ecological modeling; leaf morphology; leaf biochemistry

\end{footnotesize}

\begin{center}
\footnotesize
\begin{tabular}{lrllr}
Section & Word count &  &  & \\
\midrule
Abstract & 322 &  & Figures & 4\\
Introduction & 1035 &  & Tables & 1\\
Materials and methods & 2019 &  &  & \\
Results & 1300 &  &  & \\
Discussion & 1464 &  &  & \\
\midrule
Total & 6140 &  &  & \\
\end{tabular}
\end{center}

\linenumbers
\doublespacing
\pagebreak

\section{Abstract}
The leaf economic spectrum is a widely-studied axis of plant trait variability that defines a trade-off between leaf longevity and productivity.
While this has been investigated at the global scale, where it is robust, and at local scales, where deviations from it are common, it has received less attention at the intermediate scale of PFTs.
We investigated whether global leaf economic relationships are also present within the scale of plant functional types (PFTs) commonly used by Earth System models, and the extent to which this global-PFT hierarchy can be used to constrain trait estimates.
We developed a hierarchical multivariate Bayesian model that assumes separate means and covariance structures within and across PFTs and fit this model to seven leaf traits from the TRY database related to leaf longevity, morphology, biochemistry, and photosynthetic metabolism.
Although patterns of trait covariation were generally consistent with the leaf economic spectrum, we found three approximate tiers to this consistency.
Relationships among morphological and biochemical traits (SLA, N, P) were the most robust within and across PFTs, suggesting that covariation in these traits is driven by universal leaf construction trade-offs and stoichiometry.
Relationships among metabolic traits ($R_d$, $V_{c,\max}$, $J_{\max}$) were slightly less consistent, reflecting in part their much sparser sampling (especially for high-latitude PFTs), but also pointing to more flexible plasticity in plant metabolistm.
Finally, relationships involving leaf lifespan were the least consistent, indicating that leaf economic relationships related to leaf lifespan are dominated by across-PFT differences and that within-PFT variation in leaf lifespan is more complex and idiosyncratic.
Across all traits, these covariance were an important source of information, as evidenced by the improved imputation accuracy and reduced predictive uncertainty in multivariate models compared to univariate models.
Ultimately, our study reaffirms the value of studying not just individual traits but the multivariate trait space and the utility of hierarchical modeling for studying the scale dependence of trait relationships.

\section{Introduction}

Plant functional traits link directly measurable features of individuals to their fitness within an ecosystem, and are often related to various aspects of whole-ecosystem function \citep{violle_2007_let,cardinale_2012_biodiversity}.
Although global trait databases are larger and more open now than ever before, large gaps and sampling biases in these databases continue to pose a challenge to trait ecology \citep{cornwell_2019_what}.
If all traits and plant species were completely independent from each other, the only way forward would be to collect more trait data, which is expensive and time-consuming \citep{cornwell_2019_what}.
Fortunately, recent trait syntheses have revealed that variability in plant functional traits is constrained by biophysical limitations and trade-offs between ecological strategies \citep{kattge_2011_try,wright_2004_worldwide,kleyer_2015_why,diaz_2015_global}.

One such constraint is the "leaf economic spectrum", which defines a negative relationship between SLA and leaf lifespan, and a positive relationship of SLA with $N_{mass}$, $P_{mass}$, and photosynthesis and respiration rates \citep{wright_2004_worldwide,shipley_2006_fundamental,reich_2014_world,diaz_2015_global}.
Leaf economic traits are correlated with
plant productivity \citep{shipley_2005_functional,niinemets_2016_within,wu_2016_convergence},
litter decomposition rates \citep{bakker_2010_leaf,hobbie_2015_plant},
community composition \citep{burns_2004_patterns,cavender-bares_2004_multiple},
and ecosystem function \citep{diaz_2004_plant,musavi_2015_imprint}.
The position of species along the leaf economic spectrum is related to climate and soil conditions
\citep{wright_2004_worldwide,wright_2005_modulation,cornwell_2009_community,ordonez_2009_global,wigley_2016_leaf}.
As a result, relationships between leaf economic traits and climate have been used in ecosystem models to more finely resolve variation in plant function \citep{sakschewski_2015_leaf,verheijen_2015_inclusion,scheiter_2013_next}.

The global, interspecific trait space in which the classic leaf economic spectrum was defined is the end result of a multitude of different processes operating at different spatial, temporal, and phylogenetic scales.
The subset of these processes operating on time-scales of centuries to millennia, such as evolution or turnover in soil carbon and nutrients,
may not be relevant for predicting how individual plants and ecosystems will respond to changes on policy-relevant timescales of months to decades \citep{shaw_2012_rapid}.
The extent to which processes operating on shorter timescales result in the same trait trade-offs is an open question in trait ecology.
Observational studies of trait correlations at smaller spatial scales, such as sites, species, and individuals, produce inconsistent results,
with some studies finding consistent correlations across scales \citep{wright_2004_worldwide,albert_2010_multi,asner_2014_amazonian}
and others that correlation strength and direction are scale-dependent  \citep{albert_2010_intraspecific_functional_variability,messier_2010_how,wright_2012_does,feng_2013_scale,grubb_2015_relationships,wigley_2016_leaf,messier_2017_interspecific,kichenin_2013_contrasting}.

Many mechanisms have been suggested for scale-dependence of trait relationships.
Trade-offs may only apply when multiple competing strategies co-occur, and alternative processes can drive community assembly where strong environmental filters severely limit the range of feasible strategies \citep{rosado_2017_relative,grime_2012_evolutionary}.
Different selective pressures dominate at different scales, particularly within versus across species \citep{albert_2010_intraspecific_functional_variability,messier_2010_how,kichenin_2013_contrasting},
and different traits have different sensitivities to such pressures \citep{messier_2016_trait}.
Experimental evidence shows that species can alter different aspects of their leaf economy independently \citep{wright_2012_does}.
Global analyses show that allocation patterns (and therefore investment strategies and trait relationships) vary across plant functional types \citep{ghimire_2017_global}.
Moreover, plants maintain their fitness through multiple strategies, not just leaf economics, which can lead to multiple mutually orthogonal axes of trait variability.
As a result, changes in leaf economic traits often fail to predict changes in other aspects of plant function, such as
hydraulics \citep{li_2015_leaf},
dispersal \citep{westoby_2002_plant_ecological_strategies},
and overall plant carbon budget \citep{edwards_2014_leaf}.

For these reasons, observed global trait relationships may have limited predictive power at finer scales.
On the other hand, trying to understand ecosystem through bottom-up approaches starting with individual species is also challenging.
For one, the required species-specific trait observations do not exist for a very large number of species \citep{cornwell_2019_what}.
Even where sufficient trait data are available, scaling functional traits to ecosystem-scale processes also requires data on species relative abundance \citep{grime_1998_benefits},
which can be even more uncertain than the trait data \citep{clark_2016_why}.
Finally, plant interactions can result in community-level responses to environmental change that are distinct from the sum of species-specific changes \citep{poorter_2003_plant}.

An intermediate strategy is to aggregate species with similar structural and functional characteristics into plant functional types (PFTs).
Although PFTs are most widely used as the unit of plant functional differentiation in dynamic vegetation and earth system models
\citep{lavorel_1997_plant_functional_classifications,wullschleger_2014_plant,prentice_1992_special_paper},
the underlying concept of plant assemblages has been an important part of ecological discourse for over a century \citep{cowles_1899_ecological,clements_1936_nature,naeem_2003_disentangling}.
Patterns of trait variation within and across PFTs are relevant for several reasons.
First, trait covariance can be leveraged to impute missing trait values \citep{swenson_2013_phylogenetic};
the extent to which leveraging trait covariance reduces the size (i.e. variability and/or uncertainty) of the trait space has important implications for
quantifying the parametric uncertainty in vegetation model projections \citep{dietze_2013_improving,lebauer_2013_facilitating,dietze_2014_quantitative}.
Second, the large uncertainty in model projections of future global carbon budgets \citep{friedlingstein_2006_climate,friedlingstein_2014_uncertainties}
has been at least partially attributed to the models' failure to account for plant adaptation to changing environments \citep{sitch_2008_evaluation},
which has led to an increased interest in adding within-PFT trait plasticity to models \citep{van_2011_going,verheijen_2015_inclusion}.

While the leaf economic spectrum has been investigated at the global scale, where it is robust, and at local scales, where deviations from it are common, it has received less attention at the intermediate scale of PFTs.
Thus, this paper seeks to answer the following questions:
First, does the leaf economic spectrum hold within vs. across PFTs?
Second, can the leaf economic spectrum and similar covariance patterns be leveraged to reduce uncertainties in trait estimates, particularly under data limitation?
The answers to these question have implications for both functional ecology and ecosystem modeling.
To address these questions, we developed a hierarchical multivariate Bayesian model that explicitly accounts for across- and within-PFT variability in trait correlations.
We then fit this model to a global trait database to estimate mean trait values and variance-covariance matrices for PFTs as defined in a major earth system model (Community Land Model, CLM, \citealt{clm45_note}).
We evaluate the ability of this model to reduce uncertainties in trait estimates and reproduce observed patterns of global trait variation compared to univariate models.
Finally, we assess the scale dependence and generality of estimated trait covariances.

\section{Materials and methods}

\subsection{Trait data}

We focused on seven leaf traits obtained from the TRY global database \citep{kattge_2011_try} (see Appendix S1):
longevity (months),
specific leaf area (SLA, m$^2$ kg$^{-1}$),
nitrogen content ($N_{mass}$, mg N g$^{-1}$ or $N_{area}$, g m$^{-2}$),
phosphorus content ($P_{mass}$, mg P g$^{-1}$ or $P_{area}$, g m$^{-2}$),
dark respiration at 25°C ($R_{d,mass}$, µmol g$^{-1}$ s$^{-1}$, or $R_{d,area}$, µmol m$^{-2}$ s$^{-1}$),
maximum RuBisCO carboxylation rate at 25°C ($V_{c,\max,mass}$, µmol g$^{-1}$ s$^{-1}$, or $V_{c,\max,area}$, µmol m$^{-2}$ s$^{-1}$),
and maximum electron transport rate at 25°C ($J_{\max,mass}$, µmol g$^{-1}$ s$^{-1}$, or $J_{\max,area}$, µmol m$^{-2}$ s$^{-1}$).
For $V_{c,\max}$, we only used values reported at 25°C.
For $R_{d}$ and $J_{\max}$, we normalized the values to 25°C using reported leaf temperature values following \citet{atkin_2015_global} and \citet{kattge_2007_temperature} (equation 1 therein), respectively.
To avoid issues with trait normalization, we performed analyses separately for both mass- and area-normalized traits \citep{osnas_2013_global,lloyd_2013_photosynthetically}.
We restricted our analysis to quality-controlled values from species with sufficient information for functional type classification \cite[see][]{kattge_2011_try}.
Following past studies \cite[e.g.][]{wright_2004_worldwide,onoda_2011_global,diaz_2015_global}, we log-transformed all trait values to correct for their strong right-skewness.

\subsection{Plant functional types}

We assigned each species a PFT following the scheme in the Community Land Model (CLM4.5, \citealt{clm45_note}) (Tab. 1, Fig. 1).
We obtained categorical data on growth form, leaf type, phenology, and photosynthetic pathway from TRY.
Where species attributes disagreed between datasets, we assigned the most frequently observed attribute (e.g., if five datasets say "shrub" but only one says "tree", we would use "shrub").
Where species attributes were missing, we assigned attributes based on higher order phylogeny if possible (e.g., /Poaceae/ family are grasses, /Larix/ genus are deciduous needleleaved trees) or omitted the species if not.
For biome specification, we matched geographic coordinates for each species to annual mean temperature ($AMT$, averaged 1970-2000) data from WorldClim-2 \citep{fick_2017_worldclim},
calculated the mean $AMT$ for all sites where each species was observed,
and then binned these species based on the following cutoffs: boreal/arctic ($AMT \leq 5^\circ C$), temperate ($AMT \leq 20^\circ C$), and tropical ($AMT > 20^\circ C$).

\subsection{Multivariate analysis}

\subsubsection{Basic model description}

We compared three models with different levels of complexity.
The simplest was the "univariate" model, in which each trait is independent.
For an observation $x_{i,t}$ of trait $t$ and sample $i$:

\begin{equation}
x_{i,t} \sim N(\mu_t, \sigma_t)
\end{equation}

where $N$ is the univariate Gaussian distribution with mean $\mu_t$ and standard deviation $\sigma_t$ for trait $t$.

The second-simplest model was the "multivariate" model, in which traits are drawn from a single multivariate distribution.
For observed trait vector ${\mathbf{x_i}}$ for sample $i$:

\begin{equation}
\mathbf{x_i} \sim mvN(\mathbf{\mu}, \mathbf{\Sigma})
\end{equation}

where $mvN$ is the multivariate Gaussian distribution with mean vector $\mathbf{\mu}$ and covariance matrix $\mathbf{\Sigma}$.
We fit both of these models independently for each PFT and once for the entire dataset (i.e. one global PFT).

The most complex model was the ``hierarchical multivariate'' model (henceforth, just ``hierarchical model''),
where traits are drawn from a PFT-specific multivariate distribution describing within-PFT variation,
and whose mean vector is itself sampled from a global multivariate distribution describing variation across PFTs.
For observed trait vector $\mathbf{x}_{i,p}$ for sample $i$ belonging to PFT $p$:

\begin{equation}
\mathbf{x}_{i,p} \sim mvN(\mathbf{\mu}_p, \mathbf{\Sigma}_p)
\end{equation}

\begin{equation}
\mathbf{\mu}_p \sim mvN(\mathbf{\mu}_g, \mathbf{\Sigma}_g)
\end{equation}

where $\mathbf{\mu}_p$ and $\mathbf{\Sigma}_p$ are the mean vector and covariance matrix describing variation within PFT $p$, and $\mathbf{\mu}_g$ and $\mathbf{\Sigma}_g$ are the mean vector and covariance matrix describing across-PFT (global) variation.

\subsection{Model implementation}

We fit the above models using Gibbs sampling, which leverages conjugate prior relationships for efficient exploration of the sampling space.
The main advantages of Gibbs sampling over distribution-agnostic Bayesian algorithms such as Metropolis Hastings \citep{haario_2001_adaptive}, Differential Evolution \citep{terbraak_2008_differential}, and Hamiltonian Monte-Carlo \citep{neal_2011_hmc} is that Gibbs sampling has a 100\% proposal acceptance rate (compared to 10--65\% for these algorithms), meaning that it requires roughly 2--10 times fewer MCMC iterations.

For priors on all multivariate mean vectors ($\mathbf{\mu}$), we used multivariate normal distributions.
For priors on all multivariate variance-covariance matrices, we used the Wishart distribution ($W$), which leads to the following posterior distribution:

\begin{equation}
P(\mathbf{\Sigma} \mid
\mathbf{x}, \mathbf{\mu},
\nu_0, \mathbf{\Sigma}_0)
\sim
(W(\nu^*, S^*))^{-1}
\end{equation}

\begin{equation}
\nu^* = 1 + \nu_0 + n + m
\end{equation}

\begin{equation}
\mathbf{S^*} = (\mathbf{S}_0 + (\bar{\mathbf{x}} - \mu)^T (\bar{\mathbf{x}} - \mu))^{-1}
\end{equation}

where $n$ is the number of observations, $m$ is the number of traits in data matrix $\mathbf{x}$, and $\bar{\mathbf{x}}$ is the column means of $\mathbf{x}$.
For further details on the derivation of the conjugate relationship, see \citet[Section 3.6, "Multivariate normal with unknown mean and variance", pg. 72]{gelman_2003_bayesian}.

We used weakly-informative priors for trait means and variances (diagonals of the multivariate normal covariance matrix), the values of which are shown in Appendix S2: Table S1.
All of the covariance (off-diagonal) terms in the prior variance matrix were set to zero.
We used uninformative priors for the Wishart distribution ($\nu_0 = 0$, $\mathbf{S}_0 = \mathrm{diag}(1, m)$).

The above equations defining the conjugacy relationship do not work if the data matrix $x$ has any missing values.
Therefore, we modeled the partially missing observations as latent variables conditioned on the present observations and estimated mean vector and covariance matrix.
This approach is conceptually similar to multiple imputation \citep{white_2010_multiple,graham_2009_missing_data_analysis},
and is quite distinct from single imputation, where data are imputed once in a separate step prior to parameter estimation \citep{white_2010_multiple,graham_2009_missing_data_analysis}.
For a block of data $\mathbf{x\prime}$ containing missing observations in columns $\mathbf{m}$ and present observations in columns $\mathbf{p}$,
missing values $\mathbf{x\prime}[m]$ are drawn randomly from a conditional multivariate normal distribution at each iteration of the sampling algorithm:

\begin{equation}
\mathbf{x^\prime}[m|p] \sim mvN(\mathbf{\mu}^\prime, \mathbf{\Sigma}^\prime)
\end{equation}

\begin{equation}
\mathbf{\mu\prime} =
(\mathbf{x\prime}[p] - \mathbf{\mu^\prime}[p])
(\mathbf{\Sigma}[p,p]^{-1} \mathbf{\Sigma}[p,m])
\end{equation}

\begin{equation}
\mathbf{\Sigma\prime} = \mathbf{\Sigma}[m,m] -
\mathbf{\Sigma}[m,p]
(\mathbf{\Sigma}[p,p]^{-1} \mathbf{\Sigma}[p,m])
\end{equation}

Sampling proceeds according to the following algorithm:
Let $\mu_i$ and $\mathbf{\Sigma}_i$ be the estimates of the mean vector and covariance matrix, respectively, at MCMC iteration $i$.
Similarly, let $x_i$ be the realization of the data $\mathbf{x}\prime$ with missing (latent) values imputed at MCMC iteration $i$.

1. Initialize $\mu_1$ and $\mathbf{\Sigma}_1$ as a random draw from their respective priors.
2. Generate $x_1$ as a function of $\mu_1$ and $\mathbf{\Sigma}_1$.
3. Draw $\mu_2$ and $\mathbf{Sigma}_2$ as a function of $x_1$, according to the corresponding Gibbs sampling step.
4. Generate $x_2$ as function of $\mu_2$ and $\mathbf{\Sigma}_2$.
5. Draw $\mu_3$ and $\mathbf{Sigma}_3$ as a function of $x_2$.
6. Continue alternating these steps until a stable distribution of $\mu$ and $\mathbf{Sigma}$ is reached.

A detailed demonstration of this approach is shown in Appendix S2: Section S1.
By performing imputation at every MCMC iteration, we integrate over the uncertainty in the missing data.
Combined with uninformative priors on the covariance centered on zero (as described above),
this means our approach provides an inherently conservative estimate of both trait covariances and imputed missing values.
Where data are limited, our approach will tend towards covariance estimates of zero with wide credible intervals,
and the resulting weak and uninformative covariance estimates will lead to larger uncertainties in the imputed values.

For each model fit, we ran independent five chains, continuing sampling until the final result achieved convergence as determined by a univariate Gelman-Rubin potential scale reduction statistic less than 1.1 for all parameters \citep{gelman_1992_inference}.
We implemented this sampling algorithm in a publicly available R \citep{team_2018_r} package (<http://github.com/ashiklom/mvtraits>).

\subsection{Analysis of results}

To assess the consistency of within- and across-PFT trait trade-offs,
we calculated the mean and 95\% credible interval of the pairwise reduced major axis slope ($M$) for each trait pair ($i$, $j$)
from posterior samples of their variance-covariance matrices ($\mathbf{\Sigma}$) using the following equation:

\begin{equation}
M_{i,j} = \frac{\mathbf{\Sigma}_{j,j}}{\mathbf{\Sigma}_{i,i}} \textrm{sign}(\mathbf{\Sigma}_{i,j})
\end{equation}

Although this is a Bayesian analysis and therefore has no formal tests of statistical significance,
we approximated the statistical significance of slope estimates as those whose 95\% credible interval did not overlap zero.
We calculated reduced major axis slopes both within and across PFTs.

To explore patterns of trait variation across PFTs,
and to provide updated parameter values for earth system models,
we calculated the mean and 95\% credible intervals of PFT-level trait estimates from our hierarchical model.
We also compare these values to to the default parameter values of CLM 4.5 (Table 8.1 in \citealt{clm45_note}) for SLA, $N_{mass}$, $N_{area}$, $V_{c,\max,mass}$ and $V_{c,\max,area}$.
To convert CLM's reported C:N ratio to $N_{mass}$, we assumed a uniform leaf C fraction of 0.46.
We then divided this calculated $N_{mass}$ by the reported SLA to obtain $N_{area}$.
We calculated $V_{c,\max,mass}$ by multiplying the reported $V_{c,\max,area}$ by the reported SLA.

To compare the ability of the different models to predict missing trait observations,
we performed a cross-validation where we randomly removed 1000 observations from the data
and evaluated the ability of the fitted models to impute these missing observations.
We report the results of the normalized mean root mean square error (RMSE) of these predicted observations.

To test whether multivariate and hierarchical models offer relatively more utility at smaller sample sizes,
we calculated the relative uncertainty ($\alpha$) as a function of the mean ($\mu$) and upper ($q_{0.975}$) and lower ($q_{0.025}$) confidence limits of trait estimates.

\begin{equation}
\alpha = \frac{q_{0.975} - q_{0.025}}{\mu}
\end{equation}

We then fit a log-linear least-squares regression relating relative uncertainty to sample size ($n$) for each model (univariate, multivariate, and hierarchical; Fig. 5).

\begin{equation}
\log{\alpha} = b_0 + b_1 \log{n}
\end{equation}

If all three models performed equally well at all sample sizes, their respective slope and intercept coefficients would be statistically indistinguishable.
Meanwhile, models that perform better should have
a lower intercept ($b_0$), indicating lower overall uncertainty,
and
a lower slope ($b_1$), indicating reduced sensitivity of uncertainty ($\alpha$) to sample size ($n$).

\subsection{Data and code availability}

All R analyses were run using R version 3.6.1 \citep{r_361}.
The R code and data for running these analyses is publicly available online via the Open Science Framework at https://osf.io/w8y73/.
To comply with TRY intellectual property guidelines, the trait data used in this study have been "anonymized" such that they can only be identified to the PFT level (not the species level) as required to reproduce this analysis.
The complete TRY data request used for this analysis has been archived at http://try-db.org, and can be retrieved by providing the TRY data request ID (1584).

\section{Results}

\subsection{Trait covariance patterns within- and across-PFTs}

The direction and magnitude of pairwise trait relationships was quite variable within- and across-PFTs (Fig. 1).
Broadly, this variability can be captured by breaking up the seven leaf traits considered in this analysis into three groups:
morphology and biochemistry (SLA, N, P),
metabolism ($R_d$, $V_{c,\max}$, $J_{\max}$),
and leaf lifespan.

Morphological and biochemical traits (SLA, N, P) showed the most robust and consistent mutual covariance of these three groups.
SLA was positively related to $N_{mass}$ and $P_{mass},$ and negatively related to $N_{area}$ and $P_{area}$, both across PFTs and within all PFTs.
The magnitude of the slopes between N and P (regardless of normalization), and of SLA with $N_{area}$ and $P_{area}$, were relatively constant within all PFTs,
but the magnitude of the slopes of SLA with $N_{mass}$ and $P_{mass}$ were more variable.
In particular, temperate tree species (BlETe, BlDTe, NlETe) showed steeper SLA-$N_{mass}$ slopes (more variation in SLA relative to $N_{mass}$) than most other PFTs.

Covariance among metabolic traits ($R_d$, $V_{c,\max}$, $J_{\max}$) was slightly less robust.
Pairwise relationships among metabolic traits were weaker across-PFTs than within-PFTs.
Across-PFT relationships among metabolic traits were also weaker than across-PFT relationships among SLA, N, and P.
Within PFTs, the relationship between $V_{c,\max}$ and $J_{\max}$ (regardless of normalization) was largely consistent in magnitude and direction,
while the relationship of $R_d$ with both of these traits was more variable.
Within-PFT relationships of metabolic traits with N and P were usually positive,
and relationships with SLA were usually positive under mass normalization and negative under area normalization.
Two PFTs had notable deviations from these patterns under area normalization:
Broadleaved deciduous temperate (BlDTe) trees had opposite slopes for the SLA-$R_{d,area}$, SLA-$J_{\max,area}$, and $N_{area}$-$R_{d,area}$,
while needleleaved evergreen temperate trees (NlETe) had opposite slopes for $R_{d,area}$-$V_{c,\max,area}$ and SLA-$V_{c,\max,area}$.
Finally, an important feature of metabolic traits is the much larger number of near-zero pairwise slope estimates,
which is driven by the relative paucity of observations (especially pairwise observations) of these traits for many PFTs.

Slopes of all of the above traits with leaf lifespan showed the most variability.
Across-PFT relationships of leaf lifespan with other traits were, on average, stronger than across-PFT relationships among the other traits, especially for mass-normalized traits.
Within-PFT relationships of leaf lifespan with mass normalized traits were most often positive, but varied systematically with leaf habit and biome.
Namely, among deciduous PFTs, leaf lifespan-SLA and leaf lifespan-$N_{mass}$ slopes were less positive or more negative in colder biomes than warmer ones
(BlETr > BlETe, BlDTr > BlDTe > BlDBo, ShDTe > ShDBo, C3GTe > C3GAr).
Meanwhile, slopes of leaf lifespan with area-normalized traits were generally weaker and idiosyncratic.

An important caveat to these results is that many slopes, including all of the across-PFT slopes, had 95\% credible intervals that intersected zero---i.e. we are less than 95\% confident in the direction of these slopes.
This is primarily due to variations in the effective number of pairwise observations used to estimate the covariance matrix ---
the more pairwise observations are available, the smaller the minimum covariance that can be estimated with the same level of statistical power and confidence.
For example, a power analysis of correlation coefficients (`pwr::pwr.r.test` in R; \citealt{r_pwr_package}) showed that with 14 plant functional types ($n = 14$),
the smallest across-PFT correlation we would be able to estimate with 95\% power ($\alpha = 0.95$) and confidence ($p = 0.05$) is 0.74,
so we can confidently say that all PFT correlation coefficients (different from, but closely related to slope) were smaller than that value.
That being said, because all across-PFT slopes have the same sample size,
we can reasonably expect differences in the mean strength of pairwise across-PFT trait relationships to be ecologically meaningful.
The situation is more complex for PFT-level estimates, where sample size varies by multiple orders of magnitude by PFT and trait pair (Tab. 1, Appendix S2: Tab. S4).
In particular, high-latitude PFTs (BlDBo, NlEBo, NlD, ShDBo, and C3GAr) and metabolic traits ($R_d$, $V_{c,\max}$, $J_{\max}$) stand out as having particularly low sample sizes.

\subsection{Estimates of PFT-level means}

Across-PFT patterns in SLA, $N_{mass}$, $P_{mass}$, and $R_{d,mass}$ were similar,
with the highest values in temperate broadleaved deciduous PFTs and the lowest values in evergreen PFTs (Fig. 2).
However, none of these patterns was universal to all four traits.
For example, tropical evergreen trees had relatively high $N_{mass}$ and average SLA and $R_{d,mass}$, but among the lowest $P_{mass}$.
Similarly, compared to grass PFTs, temperate and boreal shrubs had similar SLA but higher $N_{mass}$ and $P_{mass}$.
Patterns were different when these traits were normalized by area instead of mass.
For example, needleleaf evergreen trees had relatively low $N_{mass}$ and $P_{mass}$ but relatively high $N_{area}$ and $P_{area}$, while the opposite was true of deciduous temperate trees and shrubs.

A key application of this study was to provide data-driven parameter estimates for Earth System models.
To this end, we compared our mean parameter estimates with corresponding default parameters in CLM 4.5 \citep{clm45_note} (Fig. 2).
Our SLA estimates were lower (non-overlapping 95\% credible interval) than CLM parameters for all PFTs except tropical broadleaved evergreen trees.
Our $N_{mass}$ estimates showed more across-PFT variability than CLM parameters, and only agreed with CLM for evergreen temperate trees, needleleaved trees, and C3 arctic grasses.
Similarly to \citet{kattge_2009_quantifying}, we found that CLM overestimates $V_{c,\max}$, both by mass and area.

\subsection{Comparing different models}

Both our multivariate and hierarchical models consistently outperformed the univariate approach in terms of their ability to impute missing trait values (Fig. 3).
The relative amount of improvement from the univariate to the multivariate or hierarchical model was roughly proportional to the sample size of the underlying trait.
For instance, for SLA---the best-sampled trait in our analysis---the hierarchical model's RMSE improved on the univariate model by only 4--6\%,
while the improvement for the much more sparsely observed $V_{c,\max}$ and $J_{\max}$ was 30--40\%.
The differences between the grouped multivariate model and the hierarchical model were negligible,
indicating that the additional information content of the across-PFT covariance is limited.

In general, leaf trait estimates from the univariate, multivariate, and hierarchical models were similar (Appendix S2: Fig. S1).
Where estimates differed between models, the largest differences were between the univariate and multivariate models, and additional constraint from the hierarchical model relative to PFT-specific multivariate models had a minimal effect on trait estimates.
Significant differences in trait estimates between univariate and multivariate models occurred even for well-sampled traits, such as leaf nitrogen content.
We also observed differences in posterior predictive uncertainties of mean estimates with respect to sample size.
High-latitude PFTs had large uncertainties relative to other PFTs, and the traits with the largest uncertainties were dark respiration, $V_{c,\max}$, and $J_{\max}$.
For many of these trait-PFT combinations, the additional constraint from trait covariance provided by the multivariate and hierarchical models reduced error bars, making it possible to compare estimates against those of other PFTs.
Our analysis of the relationship between sample size and trait uncertainty found that, compared to the univariate model, the multivariate model both reduced uncertainty overall (lower intercept) and reduced the sensitivity of uncertainty to sample size (lower slope) (Fig. 4).
The hierarchical model further reduced both sensitivity to sample size and overall uncertainty, but this benefit was primarily detectable only at very small sample sizes.

\section{Discussion}

\subsection{Scale dependence of the leaf economic spectrum}

Our first objective was to investigate the extent to which the global relationships defined by the leaf economic spectrum---namely, positive relationships among SLA, $N_{mass}$, $P_{mass}$, and $R_{d,mass}$ and negative relationships of all these traits with leaf lifespan \citep{wright_2004_worldwide,shipley_2006_fundamental,reich_2014_world,diaz_2015_global} ---hold within and across PFTs.
Our results suggest that, among the seven traits we investigated, there are three levels of ``robustness'' for leaf economic relationships.
The top tier of leaf economic relationships involves morphological and biochemical traits---SLA, N, and P---which had covariance patterns consistent with the leaf economic spectrum both across PFTs and within all PFTs.
The second tier involves metabolic traits---$R_d$, $V_{c,\max}$, and $J_{\max}$---which were generally consistent with the leaf economic spectrum, but with a weaker relationship across PFTs and with notable deviations within specific PFTs.
The third tier involves leaf lifespan, which had a relatively strong leaf economic spectrum signal across PFTs and within a majority of PFTs,
but which showed systematic deviations from the leaf economic spectrum within many PFTs.

The consistent direction of relationships among SLA, N, and P (by mass and area) across and within all PFTs suggests that they are driven by processes that are more-or-less universal (Fig. 1).
The consistent positive relationship between N and P (by mass or area) reflects the tight stoichiometric link between these two nutrients, and suggests that the variations in nutrient supply that would drive changes in the N:P ratio are larger within-PFTs than across \citep{elser_2010_biological}.
Meanwhile, the consistent positive relationships of SLA with mass-normalized N and P reflects the fact that increases in leaf mass per area (i.e.\ decreases in SLA) are driven primarily by increases in structural carbohydrates, which inevitably leads to a decline in nutrient mass fractions \citep{poorter_2009_causes}.
At the same time, the consistent negative relationships of SLA with area-normalized N and P reflect the role of these nutrients in structural proteins \citep{onoda_2017_physiological}.
It should be noted that although the direction of SLA-$N_{mass}$ and SLA-$P_{mass}$ relationships was consistent, the magnitude of their slopes showed non-trivial variation, particularly on a mass basis.

The less robust leaf economic spectrum signal in metabolic traits (Fig. 1) is likely a combination of two factors:
more plasticity in plant metabolism relative to morphological and biochemical traits,
and much smaller sample sizes for confidently estimating relationships.
Plasticity in plant metabolic traits independent of the leaf economic spectrum is well documented.
For example, \citealt{kattge_2009_quantifying} showed that across-PFT variation in $V_{c,\max,area}$ was driven by differences in photosynthetic N use efficiency while variation within PFTs was driven by differences in N content, and that $N_{area}$-$V_{c,\max,area}$ relationships within PFTs were variable.
More generally, there is substantial variability across PFTs in how leaf N is allocated to photosynthesis \citep{ghimire_2017_global} and across leaf biochemical consituents more generally \citep{onoda_2011_global}.
The scale dependence we observed in $V_{c,\max}$-$J_{\max}$ relationship---namely, that its slope was consistent within PFTs, but very weak across PFTs---may be a reflection of strong variation in growth irradiance and temperature across biomes, which have been shown to alter the $J_{\max}$/$V_{c,\max}$ ratio \citep{hikosaka_2005_nitrogen,hikosaka_2005_temperature,xiang_2013_contrasting}.
An important limitation to these results is the relative scarcity of metabolic trait measurements, especially for high-latitude PFTs (Table 1, Appendix 2: Tab. S4).
More simultaenous observations of metabolic traits and other leaf economic traits on the same leaf samples are needed to better understand how much these are actual ecological patterns versus just artifacts of sampling bias.

The fact that trait relationships involving leaf lifespan showed the most scale dependence and within-PFT variability (Fig. 1) is not particularly surprising considering that leaf habit (deciduous vs.\ evergreen)---the largest driver of global variability in leaf lifespan---is a part of the PFT definition.
As noted by \citealt{wright_2004_worldwide} in their original presentation of the leaf economic spectrum, specific leaf area and leaf lifespan were decoupled in deciduous species, largely because of these specues' relatively small variation in leaf lifespan.
The very inconsistent direction of relationships of area-normalized traits with leaf lifespan is also consistent with the results of \citealt{wright_2004_worldwide}.
The systematic differences in the leaf lifespan-SLA relationship with biome we observed among deciduous PFTs can be interpreted in terms of within-PFT climate variability.
Specifically, for deciduous species, leaf lifespan is primarily driven by the length of the local growing season, which generally decreases with annual mean temperature, whereas the larger variability in leaf lifespan of evergreen species is less sensitive (or even inversely related) to changes in climate (Appendix S2: Fig. S2).
Ultimately, this suggests that leaf economic relationships related to leaf lifespan are dominated by across-PFT differences, particularly those between deciduous and evergreen PFTs, while factors driving variability in leaf lifespan within PFTs are more complex and idiosyncratic \citep{reich_2014_biogeographic,wu_2016_leaf}.

\subsection{Covariance as constraint}

The second objective of this paper was to investigate the information content of trait covariance;
i.e.\ how much more can we learn about specific traits based on their relationships with other traits?
We show that accounting for covariance both improved the accuracy of trait imputation (Fig. 3) and reduced posterior predictive uncertainty around PFT-level trait means, particularly for undersampled trait-PFT combinations (Fig. 4, Appendix S2: Fig. S1).
Moreover, accounting for covariance occasionally resulted in small but statistically significant differences in the \textit{position} of trait mean estimates even for well-sampled PFT-trait combinations (e.g. $N_{mass}$ for temperate broadleaved deciduous trees, Appendix S2: Fig. S1).
This result echoes \citet{diaz_2015_global} in demonstrating the importance of studying the multivariate trait space rather than individual traits.
Significant differences between univariate and multivariate estimates of trait means suggest that sampling of these traits in TRY is not representative (Tab. 1; Appendix S2: Section S2; see also \citealt{kattge_2011_try}).
These differences also indicate that parameter estimates based on univariate trait data~\cite[e.g.,][]{lebauer_2013_facilitating,dietze_2014_quantitative,butler_2017_mapping} may not only overestimate uncertainty, but may also be systematically biased.
Although some traits in our analysis ($R_{d}$, $V_{c,\max}$, and $J_{\max}$) had too few observations to estimate covariance patterns for some PFTs with much statistical power,
we show that leveraging covariance increases the effective sample size of all traits.
This means that field and remote sensing studies that estimate only certain traits (like SLA and $N_{mass}$) may be able to use trait correlations to provide constraint on traits they do not directly observe (such as $P_{mass}$ and $R_{d,mass}$) \citep{singh_2015_imaging,musavi_2015_imprint,lepine_2016_examining,serbin_2014_spectroscopic}.
As such, future observational campaigns should consider trait covariance when deciding which traits to measure.

The additional benefit of hierarchical multivariate modeling in our study was limited, due to a combination of the low number of points used to estimate across-PFT covariance, the weak slopes of those relationships, and the usually consistent direction of pairwise slopes within and across PFTs.
Therefore, for parameterizing the current generation of ecosystem models using well-sampled traits, simple multivariate models fit independently to each PFT may be sufficient and the additional conceptual challenges and computational overhead of hierarchical modeling are not required.
However, for modeling larger numbers of PFTs, the benefits of hierarchical modeling may accumulate \citep{dietze_2008_capturing,cressie_2009_accounting,webb_2010_structured,clark_2004_why}, particularly in situations where within- and across-group covariance patterns differ.
Future work should use similar methods --- potentially in combination with additional information from phylogenetic or taxonomic similarity \citep{symonds_2014_primer} --- to explore the extent to which leaf economic relationships hold within vs.\ across other groups, such as taxonomic levels (species, genus, family, clade), successional stages, or spatial domains.

This raises the question: What is the “best” way to represent plant functional diversity in the next generation of terrestrial ecosystem models?
The current PFTs are products of an era in which computational power was more limited and data on functional diversity were relatively scarce \citep{box_1995_factors,woodward_1996_pftintro,prentice_1992_special_paper}; this study, among others, points to their limitations.
This PFT structure is, however, not immutable.
One alternative would be to explicitly account for systematic differences in trait values between regions with similar climates \citep{butler_2017_mapping}.
A second alternative is to further disaggregate PFTs based on successional stage, shade tolerance, or similar ecological characteristic \citep{longo_2019_ed1,hickler_2011_projecting}.
A third is to allow PFTs to emerge from the data by applying classification and clustering techniques to functional trait observations \citep{boulangeat_2012_improving}.
Finally, one could eschew PFTs in favor of modeling individual species \citep{post_1996_linkages,weng_2015_scaling}, or even abandon discrete categories altogether and model the continuous trait space \citep{scheiter_2013_next}.
Our methods for quantifying trait covariance would benefit any or all of these approaches.

More generally, we foresee tremendous potential for multivariate and hierarchical modeling to elucidate the relationship between traits and organismal and ecosystem function.
A natural next step would be to apply the same approach to traits whose relationship to the leaf economic spectrum is less clear.
One example is hydraulic traits:
While stem and leaf hydraulic traits are correlated \citep{bartlett_2016_correlations}, the scaling between hydraulic and leaf economic traits is poorly understood \citep{reich_2014_world,li_2015_leaf}.
Similarly, reexamining the relationships defining wood \citep{chave_2009_towards,fortunel_2012_leaf,baraloto_2010_decoupled} and root \citep{kramer-walter_2016_root,valverde-barrantes_2016_root} economic spectra, as well as their relationship to the foliar traits, would provide useful information on scale-dependence of plant growth and allocation strategies.
The difficulty of measuring hydraulic and other non-foliar traits~\cite[e.g.][]{jansen_2015_current} further increases the value of any technique that can fully leverage the information they provide.
Ultimately, multivariate and hierarchical modeling may reveal functional trade-offs that are mutually confounding at different scales, thereby enhancing our understanding of processes driving functional diversity.

\section{Acknowledgments}

This project was supported by NASA grant NNX14AH65G and NSF grants 1261582, 1458021, and 1655095, as well as the TRY initiative on plant traits (http://www.try-db.org).
The TRY initiative and database is hosted, developed, and maintained by J. Kattge and G. Boenisch (Max Planck Institute for Biogeochemistry, Jena, Germany).
TRY is currently supported by DIVERSITAS/Future Earth and the German Centre for Integrative Biodiversity Research (iDiv) Halle-Jena-Leipzig.
The authors would also like to thank Ben Bond-Lamberty, Christine Rollinson, Istem Fer, and Colin Averill for their valuable feedback on early drafts of this manuscript.

\bibliography{library}

\section{Tables}

\pagebreak

\begin{table}[!h]

\caption{\label{tab:}\label{tab:pfts}Names, labels, species counts, and number of non-missing observations of each trait for plant functional types (PFTs) used in this analysis.}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}{ccccccccccccccc}
\toprule
\multicolumn{5}{c}{ } & \multicolumn{5}{c}{Mass} & \multicolumn{5}{c}{Area} \\
\cmidrule(l{3pt}r{3pt}){6-10} \cmidrule(l{3pt}r{3pt}){11-15}
Label & PFT & Species & Leaf lifespan & SLA & N & P & Rd & Vcmax & Jmax & N & P & Rd & Vcmax & Jmax\\
\midrule
BlETr & Broadleaf Evergreen Tropical Tree & 1229 & 153 & 11710 & 7547 & 2912 & 237 & 205 & 58 & 4023 & 1684 & 326 & 225 & 152\\
BlETe & Broadleaf Evergreen Temperate Tree & 363 & 135 & 2210 & 1811 & 1194 & 121 & 36 & 16 & 928 & 339 & 196 & 106 & 87\\
BlDTr & Broadleaf Deciduous Tropical Tree & 286 & 82 & 2166 & 1545 & 812 & 98 & 54 & 30 & 813 & 500 & 113 & 56 & 53\\
BlDTe & Broadleaf Deciduous Temperate Tree & 345 & 181 & 9536 & 5982 & 2163 & 942 & 245 & 576 & 2163 & 398 & 866 & 697 & 849\\
BlDBo & Broadleaf Deciduous Boreal Tree & 62 & 58 & 908 & 898 & 340 & 142 & 0 & 0 & 141 & 60 & 11 & 5 & 5\\
\addlinespace
NlETe & Needleleaf Evergreen Temperate Tree & 130 & 66 & 2958 & 4940 & 3729 & 262 & 92 & 91 & 1227 & 462 & 84 & 274 & 106\\
NlEBo & Needleleaf Evergreen Boreal Tree & 30 & 24 & 530 & 1457 & 393 & 493 & 0 & 0 & 101 & 14 & 16 & 3 & 3\\
NlD & Needleleaf Deciduous Tree & 19 & 16 & 195 & 328 & 179 & 34 & 1 & 0 & 48 & 10 & 3 & 4 & 0\\
ShE & Shrub Evergreen & 1120 & 298 & 5018 & 3555 & 2404 & 207 & 22 & 13 & 1376 & 747 & 205 & 41 & 32\\
ShDTe & Shrub Deciduous Temperate & 330 & 100 & 3026 & 1525 & 1227 & 10 & 9 & 1 & 576 & 281 & 13 & 33 & 19\\
\addlinespace
ShDBo & Shrub Deciduous Boreal & 94 & 80 & 482 & 552 & 313 & 0 & 1 & 1 & 133 & 51 & 0 & 1 & 1\\
C3GAr & C3 Grass Arctic & 157 & 65 & 989 & 996 & 573 & 11 & 1 & 2 & 219 & 85 & 7 & 1 & 2\\
C3GTe & C3 Grass Temperate & 624 & 76 & 6322 & 3802 & 1541 & 103 & 21 & 27 & 1257 & 382 & 93 & 52 & 47\\
C4G & C4 Grass & 255 & 31 & 1312 & 1461 & 335 & 44 & 0 & 0 & 410 & 56 & 28 & 0 & 0\\
\bottomrule
\end{tabular}}
\end{table}

\section{Figures}

\pagebreak

\noindent Figure 1:
Mean pairwise reduced major axis (RMA) slope estimates calculated from within- and across-PFT covariance matrix estimates from the hierarchical model.
The slope numerator ($y$) is the outer trait and the denominator ($x$) is the inner trait (so in the top row, the slope is $\Delta(\textrm{leaf lifespan}) / \Delta(\textrm{SLA})$).
Blue colors indicate positive slopes and red colors indicate negative slopes, with darker shades indicating steeper slopes.
Asterisks (``*'') indicate slopes whose 95\% credible intervals do not overlap zero.

\vspace{\baselineskip}

\noindent Figure 2:
Mean and 95\% credible interval on best estimates of traits for each plant functional type from the hierarchical model.
For leaf lifespan and SLA, results were similar whether the other traits were normalized by mass- or area-, so only results from the mass-based fit are shown.
Values and uncertainties for estimates from the hierarchical model are reported in Appendix S2, tables S1 and S2.

\vspace{\baselineskip}

\noindent Figure 3:
Normalized mean root mean square error (RMSE) estimates from 20-fold cross-validation, by model and trait.
Normalization is such that the highest RMSE for a given trait-model combination is 1.
Model abbreviations are as follows:
`uni' is the univariate model fit separately to each PFT;\@
`multi' is the multivariate model fit separately to each PFT;\@
and `hier' is the hierarchical model.

\vspace{\baselineskip}

\noindent Figure 4:
Relative uncertainty in PFT-level trait estimates as a function of sample size for each model type.
Lines represent linear models ($\log(y) = b_0 + b_1 \log(x)$) fit independently for each model type.
In general, differences in estimate uncertainty between the univariate and multivariate models were minimal at large sample sizes but increasingly important at low sample sizes.
However, differences in estimate uncertainty between the multivariate and hierarchical models were consistently negligible.

\pagebreak

\noindent
\textbf{Author contributions:}
ANS wrote the manuscript and implemented the analysis.
ANS and EMC designed the analysis and figures.
MCD conceived the original idea for the manuscript, guided its development, and provided financial support.
MB, SJ, KK, ÜN, and NAS provided extensive feedback on multiple drafts of the manuscript, and contributed data.
CB and YO contributed data.

\noindent
\textbf{Data accessibility:}
The R code and ancillary data for running these analyses is publicly available online via the Open Science Framework at https://osf.io/w8y73/.
The TRY data request used for this analysis has been archived at http://try-db.org,
and can be retrieved by providing the TRY data request ID (\#1584).
Alternatively, the exact pre-formatted data used in this analysis are available on request to Alexey Shiklomanov (alexey.shiklomanov@pnnl.gov).

\end{document}

### 2019-11-22 shiklomanov_formatted.tex
Formatting revisions to formatted LaTeX

@@
 \usepackage{graphicx}
+\usepackage{bm}
 
@@
 
-One such constraint is the "leaf economic spectrum", which defines a negative relationship between SLA and leaf lifespan, and a positive relationship of SLA with $N_{mass}$, $P_{mass}$, and photosynthesis and respiration rates \citep{wright_2004_worldwide,shipley_2006_fundamental,reich_2014_world,diaz_2015_global}.
+One such constraint is the ``leaf economic spectrum'', which defines a negative relationship between SLA and leaf lifespan, and a positive relationship of SLA with $N_{mass}$, $P_{mass}$, and photosynthesis and respiration rates~\citep{wright_2004_worldwide,shipley_2006_fundamental,reich_2014_world,diaz_2015_global}.
 Leaf economic traits are correlated with
@@
 First, trait covariance can be leveraged to impute missing trait values \citep{swenson_2013_phylogenetic};
-the extent to which leveraging trait covariance reduces the size (i.e. variability and/or uncertainty) of the trait space has important implications for
+the extent to which leveraging trait covariance reduces the size (i.e.\ variability and/or uncertainty) of the trait space has important implications for
 quantifying the parametric uncertainty in vegetation model projections \citep{dietze_2013_improving,lebauer_2013_facilitating,dietze_2014_quantitative}.
@@
 Thus, this paper seeks to answer the following questions:
-First, does the leaf economic spectrum hold within vs. across PFTs?
+First, does the leaf economic spectrum hold within vs.\ across PFTs?
 Second, can the leaf economic spectrum and similar covariance patterns be leveraged to reduce uncertainties in trait estimates, particularly under data limitation?
@@
 To avoid issues with trait normalization, we performed analyses separately for both mass- and area-normalized traits \citep{osnas_2013_global,lloyd_2013_photosynthetically}.
-We restricted our analysis to quality-controlled values from species with sufficient information for functional type classification \cite[see][]{kattge_2011_try}.
-Following past studies \cite[e.g.][]{wright_2004_worldwide,onoda_2011_global,diaz_2015_global}, we log-transformed all trait values to correct for their strong right-skewness.
+We restricted our analysis to quality-controlled values from species with sufficient information for functional type classification~\cite[see][]{kattge_2011_try}.
+Following past studies~\cite[e.g.][]{wright_2004_worldwide,onoda_2011_global,diaz_2015_global}, we log-transformed all trait values to correct for their strong right-skewness.
 
@@
 We obtained categorical data on growth form, leaf type, phenology, and photosynthetic pathway from TRY.
-Where species attributes disagreed between datasets, we assigned the most frequently observed attribute (e.g., if five datasets say "shrub" but only one says "tree", we would use "shrub").
-Where species attributes were missing, we assigned attributes based on higher order phylogeny if possible (e.g., /Poaceae/ family are grasses, /Larix/ genus are deciduous needleleaved trees) or omitted the species if not.
-For biome specification, we matched geographic coordinates for each species to annual mean temperature ($AMT$, averaged 1970-2000) data from WorldClim-2 \citep{fick_2017_worldclim},
+Where species attributes disagreed between datasets, we assigned the most frequently observed attribute (e.g., if five datasets say ``shrub'' but only one says ``tree'', we would use ``shrub'').
+Where species attributes were missing, we assigned attributes based on higher order phylogeny if possible (e.g., \textit{Poaceae} family are grasses, \textit{Larix} genus are deciduous needleleaved trees) or omitted the species if not.
+For biome specification, we matched geographic coordinates for each species to annual mean temperature ($AMT$, averaged 1970--2000) data from WorldClim-2 \citep{fick_2017_worldclim},
 calculated the mean $AMT$ for all sites where each species was observed,
@@
 We compared three models with different levels of complexity.
-The simplest was the "univariate" model, in which each trait is independent.
+The simplest was the ``univariate'' model, in which each trait is independent.
 For an observation $x_{i,t}$ of trait $t$ and sample $i$:
@@
 
-The second-simplest model was the "multivariate" model, in which traits are drawn from a single multivariate distribution.
-For observed trait vector ${\mathbf{x_i}}$ for sample $i$:
+The second-simplest model was the ``multivariate'' model, in which traits are drawn from a single multivariate distribution.
+For observed trait vector ${\bm{x_i}}$ for sample $i$:
 
 \begin{equation}
-\mathbf{x_i} \sim mvN(\mathbf{\mu}, \mathbf{\Sigma})
+\bm{x_i} \sim mvN(\bm{\mu}, \bm{\Sigma})
 \end{equation}
 
-where $mvN$ is the multivariate Gaussian distribution with mean vector $\mathbf{\mu}$ and covariance matrix $\mathbf{\Sigma}$.
-We fit both of these models independently for each PFT and once for the entire dataset (i.e. one global PFT).
+where $mvN$ is the multivariate Gaussian distribution with mean vector $\bm{\mu}$ and covariance matrix $\bm{\Sigma}$.
+We fit both of these models independently for each PFT and once for the entire dataset (i.e.\ one global PFT).
 
@@
 and whose mean vector is itself sampled from a global multivariate distribution describing variation across PFTs.
-For observed trait vector $\mathbf{x}_{i,p}$ for sample $i$ belonging to PFT $p$:
+For observed trait vector $\bm{x}_{i,p}$ for sample $i$ belonging to PFT $p$:
 
 \begin{equation}
-\mathbf{x}_{i,p} \sim mvN(\mathbf{\mu}_p, \mathbf{\Sigma}_p)
+\bm{x}_{i,p} \sim mvN(\bm{\mu}_p, \bm{\Sigma}_p)
 \end{equation}
@@
 \begin{equation}
-\mathbf{\mu}_p \sim mvN(\mathbf{\mu}_g, \mathbf{\Sigma}_g)
+\bm{\mu}_p \sim mvN(\bm{\mu}_g, \bm{\Sigma}_g)
 \end{equation}
 
-where $\mathbf{\mu}_p$ and $\mathbf{\Sigma}_p$ are the mean vector and covariance matrix describing variation within PFT $p$, and $\mathbf{\mu}_g$ and $\mathbf{\Sigma}_g$ are the mean vector and covariance matrix describing across-PFT (global) variation.
+where $\bm{\mu}_p$ and $\bm{\Sigma}_p$ are the mean vector and covariance matrix describing variation within PFT $p$, and $\bm{\mu}_g$ and $\bm{\Sigma}_g$ are the mean vector and covariance matrix describing across-PFT (global) variation.
 
@@
 
-For priors on all multivariate mean vectors ($\mathbf{\mu}$), we used multivariate normal distributions.
+For priors on all multivariate mean vectors ($\bm{\mu}$), we used multivariate normal distributions.
 For priors on all multivariate variance-covariance matrices, we used the Wishart distribution ($W$), which leads to the following posterior distribution:
@@
 \begin{equation}
-P(\mathbf{\Sigma} \mid
-\mathbf{x}, \mathbf{\mu},
-\nu_0, \mathbf{\Sigma}_0)
+P(\bm{\Sigma} \mid
+\bm{x}, \bm{\mu},
+\nu_0, \bm{\Sigma}_0)
 \sim
@@
 \begin{equation}
-\mathbf{S^*} = (\mathbf{S}_0 + (\bar{\mathbf{x}} - \mu)^T (\bar{\mathbf{x}} - \mu))^{-1}
+\bm{S^*} = (\bm{S}_0 + (\bar{\bm{x}} - \mu)^T (\bar{\bm{x}} - \mu))^{-1}
 \end{equation}
 
-where $n$ is the number of observations, $m$ is the number of traits in data matrix $\mathbf{x}$, and $\bar{\mathbf{x}}$ is the column means of $\mathbf{x}$.
+where $n$ is the number of observations, $m$ is the number of traits in data matrix $\bm{x}$, and $\bar{\bm{x}}$ is the column means of $\bm{x}$.
 For further details on the derivation of the conjugate relationship, see \citet[Section 3.6, "Multivariate normal with unknown mean and variance", pg. 72]{gelman_2003_bayesian}.
@@
 All of the covariance (off-diagonal) terms in the prior variance matrix were set to zero.
-We used uninformative priors for the Wishart distribution ($\nu_0 = 0$, $\mathbf{S}_0 = \mathrm{diag}(1, m)$).
+We used uninformative priors for the Wishart distribution ($\nu_0 = 0$, $\bm{S}_0 = \mathrm{diag}(1, m)$).
 
@@
 and is quite distinct from single imputation, where data are imputed once in a separate step prior to parameter estimation \citep{white_2010_multiple,graham_2009_missing_data_analysis}.
-For a block of data $\mathbf{x\prime}$ containing missing observations in columns $\mathbf{m}$ and present observations in columns $\mathbf{p}$,
-missing values $\mathbf{x\prime}[m]$ are drawn randomly from a conditional multivariate normal distribution at each iteration of the sampling algorithm:
+For a block of data $\bm{x\prime}$ containing missing observations in columns $\bm{m}$ and present observations in columns $\bm{p}$,
+missing values $\bm{x\prime}[m]$ are drawn randomly from a conditional multivariate normal distribution at each iteration of the sampling algorithm:
 
 \begin{equation}
-\mathbf{x^\prime}[m|p] \sim mvN(\mathbf{\mu}^\prime, \mathbf{\Sigma}^\prime)
+\bm{x^\prime}[m|p] \sim mvN(\bm{\mu}^\prime, \bm{\Sigma}^\prime)
 \end{equation}
@@
 \begin{equation}
-\mathbf{\mu\prime} =
-(\mathbf{x\prime}[p] - \mathbf{\mu^\prime}[p])
-(\mathbf{\Sigma}[p,p]^{-1} \mathbf{\Sigma}[p,m])
+\bm{\mu\prime} =
+(\bm{x\prime}[p] - \bm{\mu^\prime}[p])
+(\bm{\Sigma}[p,p]^{-1} \bm{\Sigma}[p,m])
 \end{equation}
@@
 \begin{equation}
-\mathbf{\Sigma\prime} = \mathbf{\Sigma}[m,m] -
-\mathbf{\Sigma}[m,p]
-(\mathbf{\Sigma}[p,p]^{-1} \mathbf{\Sigma}[p,m])
+\bm{\Sigma\prime} = \bm{\Sigma}[m,m] -
+\bm{\Sigma}[m,p]
+(\bm{\Sigma}[p,p]^{-1} \bm{\Sigma}[p,m])
 \end{equation}
@@
 Sampling proceeds according to the following algorithm:
-Let $\mu_i$ and $\mathbf{\Sigma}_i$ be the estimates of the mean vector and covariance matrix, respectively, at MCMC iteration $i$.
-Similarly, let $x_i$ be the realization of the data $\mathbf{x}\prime$ with missing (latent) values imputed at MCMC iteration $i$.
+Let $\mu_i$ and $\bm{\Sigma}_i$ be the estimates of the mean vector and covariance matrix, respectively, at MCMC iteration $i$.
+Similarly, let $x_i$ be the realization of the data $\bm{x}\prime$ with missing (latent) values imputed at MCMC iteration $i$.
 
-1. Initialize $\mu_1$ and $\mathbf{\Sigma}_1$ as a random draw from their respective priors.
-2. Generate $x_1$ as a function of $\mu_1$ and $\mathbf{\Sigma}_1$.
-3. Draw $\mu_2$ and $\mathbf{Sigma}_2$ as a function of $x_1$, according to the corresponding Gibbs sampling step.
-4. Generate $x_2$ as function of $\mu_2$ and $\mathbf{\Sigma}_2$.
-5. Draw $\mu_3$ and $\mathbf{Sigma}_3$ as a function of $x_2$.
-6. Continue alternating these steps until a stable distribution of $\mu$ and $\mathbf{Sigma}$ is reached.
+\begin{enumerate}
+\item Initialize $\mu_1$ and $\bm{\Sigma}_1$ as a random draw from their respective priors.
+\item Generate $x_1$ as a function of $\mu_1$ and $\bm{\Sigma}_1$.
+\item Draw $\mu_2$ and $\bm{\Sigma}_2$ as a function of $x_1$, according to the corresponding Gibbs sampling step.
+\item Generate $x_2$ as function of $\mu_2$ and $\bm{\Sigma}_2$.
+\item Draw $\mu_3$ and $\bm{\Sigma}_3$ as a function of $x_2$.
+\item Continue alternating these steps until a stable distribution of $\mu$ and $\bm{\Sigma}$ is reached.
+\end{enumerate}
 
@@
 we calculated the mean and 95\% credible interval of the pairwise reduced major axis slope ($M$) for each trait pair ($i$, $j$)
-from posterior samples of their variance-covariance matrices ($\mathbf{\Sigma}$) using the following equation:
+from posterior samples of their variance-covariance matrices ($\bm{\Sigma}$) using the following equation:
 
 \begin{equation}
-M_{i,j} = \frac{\mathbf{\Sigma}_{j,j}}{\mathbf{\Sigma}_{i,i}} \textrm{sign}(\mathbf{\Sigma}_{i,j})
+M_{i,j} = \frac{\bm{\Sigma}_{j,j}}{\bm{\Sigma}_{i,i}} \textrm{sign}(\bm{\Sigma}_{i,j})
 \end{equation}
@@
 The R code and data for running these analyses is publicly available online via the Open Science Framework at https://osf.io/w8y73/.
-To comply with TRY intellectual property guidelines, the trait data used in this study have been "anonymized" such that they can only be identified to the PFT level (not the species level) as required to reproduce this analysis.
+To comply with TRY intellectual property guidelines, the trait data used in this study have been ``anonymized'' such that they can only be identified to the PFT level (not the species level) as required to reproduce this analysis.
 The complete TRY data request used for this analysis has been archived at http://try-db.org, and can be retrieved by providing the TRY data request ID (1584).
@@
 Morphological and biochemical traits (SLA, N, P) showed the most robust and consistent mutual covariance of these three groups.
-SLA was positively related to $N_{mass}$ and $P_{mass},$ and negatively related to $N_{area}$ and $P_{area}$, both across PFTs and within all PFTs.
+SLA was positively related to $N_{mass}$ and $P_{mass}$, and negatively related to $N_{area}$ and $P_{area}$, both across PFTs and within all PFTs.
 The magnitude of the slopes between N and P (regardless of normalization), and of SLA with $N_{area}$ and $P_{area}$, were relatively constant within all PFTs,
@@
 
-An important caveat to these results is that many slopes, including all of the across-PFT slopes, had 95\% credible intervals that intersected zero---i.e.
-we are less than 95\% confident in the direction of these slopes.
+An important caveat to these results is that many slopes, including all of the across-PFT slopes, had 95\% credible intervals that intersected zero---i.e.\ we are less than 95\% confident in the direction of these slopes.
 This is primarily due to variations in the effective number of pairwise observations used to estimate the covariance matrix ---
 the more pairwise observations are available, the smaller the minimum covariance that can be estimated with the same level of statistical power and confidence.
-For example, a power analysis of correlation coefficients (`pwr::pwr.r.test` in R; \citealt{r_pwr_package}) showed that with 14 plant functional types ($n = 14$),
+For example, a power analysis of correlation coefficients (`pwr::pwr.r.test' in R;\@\citealt{r_pwr_package}) showed that with 14 plant functional types ($n = 14$),
 the smallest across-PFT correlation we would be able to estimate with 95\% power ($\alpha = 0.95$) and confidence ($p = 0.05$) is 0.74,

### 2019-11-26 shiklomanov_formatted.tex
Final submission revisions

@@
 
-\input{authors.tex}
+\usepackage[noblocks]{authblk}
+
+\author[1,*]{Alexey N. Shiklomanov <alexey.shiklomanov@pnnl.gov>}
+\affil[1]{Joint Global Change Research Institute, Pacific Northwest National Laboratory, College Park, MD}
+
+\author[2]{Elizabeth M. Cowdery <ecowdery@bu.edu>}
+\affil[2]{Department of Earth \& Environment, Boston University, Boston, MA}
+
+\author[3]{Michael Bahn <Michael.Bahn@uibk.ac.at>}
+\affil[3]{Institute of Ecology, University of Innsbruck, 6020 Innsbruck, Austria}
+
+\author[4]{Chaeho Byun <chaeho.byun@mail.mcgill.ca>}
+\affil[4]{School of Civil and Environmental Engineering, Yonsei University, Seoul 03722, Korea}
+
+\author[5]{Steven Jansen <steven.jansen@uni-ulm.de>}
+\affil[5]{Institute of Systematic Botany and Ecology, Ulm University, Albert-Einstein-Allee 11, 89081, Ulm, Germany}
+
+\author[6]{Koen Kramer <koen.kramer@wur.nl>}
+\affil[6]{Department of Vegetation, Forest, and Landscape Ecology, Wageningen Environmental Research and Wageningen University, Wageningen, Gelderland, The Netherlands}
+
+\author[7,8]{Vanessa Minden <vanessa.minden@uni-oldenburg.de>}
+\affil[7]{Institute for Biology and Environmental Sciences, Carl von Ossietzky-University of Oldenburg, Carl von Ossietzky Str. 9-11, 26129 Oldenburg, Germany}
+\affil[8]{Department of Biology, Ecology and Evolution, Vrije Universiteit Brussel, Pleinlaan 2, 1050 Brussels}
+
+\author[9]{\"Ulo Niinemets <ylo.niinemets@emu.ee>}
+\affil[9]{Institute of Agricultural and Environmental Sciences, Estonian University of Life Sciences, Kreutzwaldi 1, 51014 Tartu, Estonia}
+
+\author[10]{Yusuke Onoda <yusuke.onoda@gmail.com>}
+\affil[10]{Graduate School of Agriculture, Kyoto University, Kyoto, 605-8503, Japan}
+
+\author[11]{Nadejda A. Soudzilovskaia <n.a.soudzilovskaia@cml.leidenuniv.nl>}
+\affil[11]{Conservation Biology Department, Institute of Environmental Sciences, Leiden University, Leiden, The Netherlands}
+
+\author[2]{Michael C. Dietze <dietze@bu.edu>}
+
+\affil[*]{Corresponding author; Phone: (301) 314-6713; Fax: (301) 314-6719; Email: alexey.shiklomanov@pnnl.gov; Mail: 5825 University Research Ct., Office 3533, College Park, MD 20740}
+
+\renewcommand\Authfont{\footnotesize}
+\renewcommand\Affilfont{\scriptsize}
 \date{}
@@
 \textbf{Running head:} Trait covariance within vs.\ across PFTs (39 characters)
-
-\noindent
-\textbf{Keywords:} Functional trade-off; hierarchical modeling; trait variation; ecological modeling; leaf morphology; leaf biochemistry
 
@@
 Ultimately, our study reaffirms the value of studying not just individual traits but the multivariate trait space and the utility of hierarchical modeling for studying the scale dependence of trait relationships.
+
+\noindent
+\textbf{Keywords:} Functional trade-off; hierarchical modeling; trait variation; ecological modeling; leaf morphology; leaf biochemistry
 
@@
 For each model fit, we ran independent five chains, continuing sampling until the final result achieved convergence as determined by a univariate Gelman-Rubin potential scale reduction statistic less than 1.1 for all parameters \citep{gelman_1992_inference}.
-We implemented this sampling algorithm in a publicly available R \citep{team_2018_r} package (<http://github.com/ashiklom/mvtraits>).
+We implemented this sampling algorithm in a publicly available R \citep{r_361} package (http://github.com/ashiklom/mvtraits).
 
@@
 
-\bibliography{library}
+\bibliography{bibliography}
 
-\section{Tables}
+\vspace{5mm}
+
+\noindent
+\textbf{Data accessibility:}
+R code and data are available from the Open Science Framework: https://osf.io/w8y73/.
+Trait data are archived in the TRY Plant Trait Database at http://try-db.org and can be retrieved by providing TRY data request ID \#1584.
+
+\vspace{5mm}
+
+\noindent
+\textbf{Author contributions:}
+ANS wrote the manuscript and implemented the analysis.
+ANS and EMC designed the analysis and figures.
+MCD conceived the original idea for the manuscript, guided its development, and provided financial support.
+MB, SJ, KK, ÜN, and NAS provided extensive feedback on multiple drafts of the manuscript, and contributed data.
+CB and YO contributed data.
 
@@
 
+\section*{Tables}
+
 \begin{table}[!h]
 
-\caption{\label{tab:}\label{tab:pfts}Names, labels, species counts, and number of non-missing observations of each trait for plant functional types (PFTs) used in this analysis.}
+\caption{\label{tab:pfts} Names, labels, species counts, and number of non-missing observations of each trait for plant functional types (PFTs) used in this analysis.}
 \centering
@@
 
-\section{Figures}
+\pagebreak
 
-\pagebreak
+\section*{Figures}
 
@@
 
-\pagebreak
-
-\noindent
-\textbf{Author contributions:}
-ANS wrote the manuscript and implemented the analysis.
-ANS and EMC designed the analysis and figures.
-MCD conceived the original idea for the manuscript, guided its development, and provided financial support.
-MB, SJ, KK, ÜN, and NAS provided extensive feedback on multiple drafts of the manuscript, and contributed data.
-CB and YO contributed data.
-
-\noindent
-\textbf{Data accessibility:}
-The R code and ancillary data for running these analyses is publicly available online via the Open Science Framework at https://osf.io/w8y73/.
-The TRY data request used for this analysis has been archived at http://try-db.org,
-and can be retrieved by providing the TRY data request ID (\#1584).
-Alternatively, the exact pre-formatted data used in this analysis are available on request to Alexey Shiklomanov (alexey.shiklomanov@pnnl.gov).
-
 \end{document}
